---
title: "Zachary's Karate Club"
author: "Team Blue"
date: "3/23/2022"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```

Loading packages.

```{r}
library(readr)
library(tidyverse)
library(tidygraph)
library(igraph)
library(ggraph)
library(ggforce)
```

# 1. Import the data and create an undirected tbl_graph called karate_netw.

```{r}
karate_edges <- read_table(
  "soc-karate.zip",
  col_names = FALSE,
  skip = 24
)

# Adding node argument to make the name "node" to make things easier
# Adding sort argument to ensure values are calibrated
# Adding unique argument to make sure there's no duplicates
# data frame type is required for network to function

karate_nodes <- tibble(
  node = sort(unique(c(
    karate_edges$X1,
    karate_edges$X2
  )))
)
karate_netw <- tbl_graph(
  nodes = karate_nodes,
  edges = karate_edges,
  directed = FALSE
)
```

# 2. How many nodes and edges are in the network?

```{r}
# Number of nodes
karate_netw |>
    activate(nodes) |>
    as.tibble() |>
    nrow()

# Number of edges
karate_netw |>
    activate(edges) |>
    as.tibble() |>
    nrow()

```

# 3. Are there parallel edges or loops in the network?

```{r}
is_simple(karate_netw)
```
No, the graph is simple so there are no parallel edges or loops.

# 4. Node 1 is the karate instructor. Node 34 is the club president. Add a node attribute called role to the tbl_graph with the values Instructor, President and Other.

```{r}
instructor_id <- 1
president_id <- 34
karate_netw <- 
    karate_netw |>
    activate(nodes) |> # advantage of activation is that if you insert other code in between the code will always work
    mutate(role = case_when(
        node == instructor_id ~ "Instructor", 
        node == president_id ~ "President",
        TRUE ~ "Other"
        ))
```

# 5. Show that the two nodes with the highest degree are the president and the instructor.

```{r}
karate_netw <- 
    karate_netw |>
    activate(nodes) |>
    mutate(degree = centrality_degree())
karate_netw |>
    activate(nodes) |>
    as.tibble() |>
    slice_max(degree, n = 2)
```

# 6. Find all shortest paths between the instructor and the president.

```{r}
instructor_to_president <- all_shortest_paths(karate_netw, 1, 34)
instructor_to_president$res
```

# 7. What is the local clustering coefficient of the instructor? What is the local clustering coefficient of the president? How do you interpret these numbers?

```{r}
karate_netw <- 
    karate_netw |>
    activate(nodes) |>
    mutate(cluster_coef = local_transitivity())
karate_netw |>
    activate(nodes) |>
    as.tibble() |>
    filter(role %in% c("Instructor", "President"))
```

The local clustering coefficient of a node quatifies how close the neigbouring nodes are to being a clique. 

# 8. Which are the largest cliques in the network? Which nodes are in these cliques?

```{r}
large_cliques <- largest_cliques(karate_netw)
large_cliques

unique(large_cliques[[1]])
```

# 9. Confirm that the network is connected (i.e. there is only one component).

```{r}
karate_netw |>
  activate(nodes) |>
  mutate(component_id = group_components()) |>
  as_tibble() |>
  select(component_id) |>
  unique()
```
Since there is only one connected component, the network is connected.

# 10. Which communities does the Louvain algorithm detect? Use tidygraph to add information about the communities to karate_netw.

```{r}
karate_netw <-
  karate_netw |>
  activate(nodes) |>
  mutate(community_id = group_louvain())
karate_netw |>
  as_tibble() |> # Extract node information as tibble
  group_by(community_id) |>
  group_split() |>
  map(pluck("node"))
```

# 11. Make a plot similar to figure 1.

```{r}
karate_layout <- create_layout(karate_netw, layout = "stress")
layout_attributes <- attributes(karate_layout)
node_positions <- read_csv("karate_node_positions.csv")
karate_layout <-
  karate_layout |>
  select(-x, -y) |>
  left_join(node_positions, by = "node") |>
  select(x, y, everything())
attributes(karate_layout) <- layout_attributes
###NOT COMPLETE###
ggraph(karate_layout) +
  geom_edge_link() +
  geom_mark_hull(
    aes(
      x,
      y,
      fill = as.factor(community_id),
      label = str_c("community ", community_id)
    ),
    colour = NA,
    con.colour = "grey",
    show.legend = FALSE
  ) +
  labs(
    title = "Zachary's Karate Club",
    subtitle = "Communities assigned by Louvain algorithm",
    caption = "Source: https://networkrepository.com" 
  ) +
  geom_node_label(aes(label = node, color = role)) +
  scale_color_brewer(name = "Role", palette = "Dark2") + 
  geom_curve(aes(x= 0.35, y= 1.25, xend = 0.45, yend = -0.6), curvature = 0.9) +
  annotate("text", x= 0.3, y= 1.4, label = "Observed split")
```

# 12. Briefly explain to a reader what the figure reveals about the karate-club network.

The figure reveals that the groups that formed after the observed split was for the most part "predicted" by the network analysis on communties. It shows that Community 1 and 4 split with the Instructor, while Commuinty 2 and 3 went with the President. 