---
title: "Solar Cyles"
author: "Team Blue"
date: "3/30/2022"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, fig.align = "center")
```

Loading libraries
```{r}
library(tidyverse)
library(lubridate)
library(timetk)
```

# 1

```{r}
sunspots <- read_delim(
  "SN_d_tot_V2.0.csv",
  delim = ";",
  escape_double = FALSE,
  col_names = c(
    "year",
    "month",
    "day",
    "year_fraction",
    "ss_number",
    "sd",
    "n_obs",
    "definitive"
  ),
  col_types = "iiininii",
  na = c("-1", "-1.0"),
  trim_ws = TRUE
)
```

The variables are year, month, day, date in fraction of year, daily total sunspot number, daily standard deviation of the sunspot numbers, number of observations, and definitive/provisional indicator, respectively. 

Missing values are "-1" or "-1.0".

# 2

```{r}
sunspots <- mutate(sunspots, date = make_date(year, month, day))
```

# 3

```{r}
my_stamp <- function(x) {
  stamp("Tuesday, 3 December, 2019", quiet = TRUE)(x)
}

sunspots |>
  filter(is.na(ss_number)) |>
  slice_max(n = 1, date) |>
  pull(date) |>
  my_stamp()
```

# 4

```{r}
sunspots <-
  sunspots |>
  mutate(diff_to_prev_data = date - lag(date))

# Checking if all data points have a difference of 1 day from the previous data.
all(sunspots$diff_to_prev_data == 1, na.rm = TRUE)
```
Yes, they are all consecutive.

# 5

```{r}
spots <-
  sunspots |>
  summarise_by_time(
    date,
    .by = "month",
    mean_daily_ss_number = mean(ss_number, na.rm = TRUE)
  ) |>
  filter(date >= ymd("1975-01-01"))
```

# 6

```{r}
solar_cycles <-
  read_csv("solar_cycles.csv") |>
  mutate(
    `Start (Minimum)` = ym(`Start (Minimum)`),
    Maximum = ym(Maximum)
  )
```

# 7

```{r}
### SOLVE###
solar_cycles <-
  solar_cycles |>
  mutate(
    isequal = round(
      interval(
        `Start (Minimum)`, Maximum
      ) / years(1),
      digits = 1
    ) == `Time of Rise (years)`
  )
table(solar_cycles$isequal)
```

All are equal except two entries.

# 8

```{r}
solar_cycles <- filter(solar_cycles, `Start (Minimum)` >= ymd("1975-01-01"))
```

# 9

```{r}
solar_cycles <-
  solar_cycles |>
  mutate(
    next_start = lead(`Start (Minimum)`),
    mid_date = `Start (Minimum)` + ((next_start - `Start (Minimum)`) / 2),
    next_start = NULL
  )
```

# 10

```{r, fig.height = 4, fig,width = 4, out.width = '60%'}
solar_cycles <-
  solar_cycles |>
  mutate(`Solar Cycle Number` = 21:25)

ggplot(data = spots, aes(date, mean_daily_ss_number)) +
  geom_point(alpha = 0.5, color = "turquoise4") +
  labs(
    title = "Solar Activity",
    subtitle = "Sunspot number averaged by month",
    caption = "Source: Royal Observatory of Belgium",
    x = "Year",
    y = "Mean daily sunspot number",
    linetype = "LOESS fit",
    label = "Solar cycle",
    fill = "Solar cycle"
  ) +
  geom_smooth(se = FALSE, span = 0.1, color = "black", aes(linetype = "")) +
  geom_label(
    data = solar_cycles,
    aes(mid_date, 1, label = `Solar Cycle Number`, fill = "")
  ) +
  theme(legend.position = c(0.9, 0.70))
```

# 11

The plot reveals that sunspot density generally happens in cycles. We see that there is a trough, whereby there is generally lower sunspot density around the years of 1976, 1986, 1997, 2009, 2019. We also see there is a peak, whereby there is generally higher sunspot density around the years of 1980, 1990, 2002, 2014. This is plausible, as solar eclipses and other solar occurrences also happen on a cyclical basis.
What is interesting is that the peaks are seen to have a decreasing trend over the past 40 years and I speculate that this may be due to climate change. Increased CO2 in the atmosphere may reduce the number of sunspots visible on the sun.