---
title: "Week08_davis_fri"
author: "shuna"
date: "3/11/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE) 
```

```{r}
library(countrycode)
library(gapminder)
library(readxl)
library(tidyverse)
```

# 1. load data 

```{r}
gdp <- readxl::read_excel("gdp.xls", skip = 3, sheet = "Data") |>
  rename(country = "Country Name", country_code = "Country Code")
life <- readxl::read_excel("life.xls", skip = 3, sheet = "Data") |>
  rename(country = "Country Name", country_code = "Country Code")
pop <- readxl::read_excel("pop.xls", skip = 3, sheet = "Data") |>
  rename(country = "Country Name", country_code = "Country Code")
```

# 2. check country code 

```{r}
all(gdp$country_code == life$country_code)
all(gdp$country_code == pop$country_code)
```
# 3. merge the three speadsheets into a single tibble
```{r}
gdp_pivoted <- pivot_longer(
  data = gdp, 
  cols = c('1960':'2020'),
  names_to = "year",
  values_to = "gdp"
  ) |>
  select(country, country_code, year, gdp)

life_pivoted <- pivot_longer(
  data = life, 
  cols = c('1960':'2020'),
  names_to = "year",
  values_to = "life"
  )|>
  select(country, country_code, year, life)

pop_pivoted <- pivot_longer(
  data = pop, 
  cols = c('1960':'2020'),
  names_to = "year",
  values_to = "pop"
  ) |>
  select(country, country_code, year, pop)


wb <- gdp_pivoted |>
  left_join(life_pivoted) |>
  left_join(pop_pivoted)
```

# 4. What "countries" do not have the matching codes? Do the result make sense? 
```{r}
wb |>
  anti_join(codelist, by = c("country_code" = "iso3c")) |>
  select("Country", "Country code") |>
  unique()
```


# 5. Use dplyr to remove rows that do not match any country code in codelist. 
```{r}
wb_removed <- wb |>
  semi_join(codelist, by = c("country_code" = "iso3c"))
```

# 6. 
```{r}
wb <- wb_removed |>
  filter(!(is.na(gdp) | is.na(life) | is.na(pop)))

missing_values <- wb_removed |>
  filter(is.na(gdp) | is.na(life) | is.na(pop)) |>
  group_by(year) |>
  summarise(na_countries = n())

head(missing_values, n = 6)
```


# 7. Plot the number of missing countries per year 
```{r}
ggplot(data = missing_values, aes(as.integer(year), na_countries)) + 
  geom_line() +
  labs(
    x = "Year",
    y = "Number of Countries with NA",
    title = "Countries with NA by year",
    caption = "Source: World Bank"
  ) + 
  theme(
    plot.title = element_text(hjust = 0.5)
  ) +
  ylim(0,225) +
  scale_x_continuous(breaks= c(seq(1960, 2020, 10)))
```

# 8
```{r}
gap <-
  gapminder_unfiltered |>
  filter(country != "Netherlands Antilles") |>
  mutate(country_code = countrycode(country, "country.name", "iso3c"))
```

## Are there countries in gap without a country code?
```{r}
sum(is.na(gap$country_code))
```
## Are there countries that share the same country code? 
```{r}
gap |>
  select(country, country_code) |>
  distinct() |>
  group_by(country_code) |>
  summarise(overlaps = n()) |>
  filter(overlaps != 1) 
```

```{r}
all(length(unique(gap$country_code)) == length(unique(gap$country)))
```

# 9. Which countries are in gap but not in wb 

```{r}
gap |>
  select(country) |>
  filter(!(country %in% wb$country)) |>
  distinct() 
```

# 10. 

```{r}
wb <- wb |>
  filter(year == 2007) |>
  select(country, gdp_wb = gdp)
gap <- gap |>
  filter(year == 2007) |>
  select(country, gdp_gap = gdpPercap)

wb
gap
wb_gap <- wb |>
  inner_join(gap)
wb_gap
```

# 11.
```{r}
Q11 <- wb_gap |>
  mutate(perc = 100*(gdp_gap - gdp_wb)/gdp_wb)
```


# 12. 
```{r}
Q11 |>
  arrange(perc) |>
  head(5)

Q11 |>
  arrange(-perc) |>
  head(5)
```


