---
title: "R Notebook"
output: html_notebook
---
# Import the data
```{r}
oph <- read.csv("HornbillID.csv")
```

# Packages 
```{r}
#load the packages 
library(tidyverse)
library(tibble)
library(ggplot2)
library(vegan)
library(dplyr)
library(RVAideMemoire)
library(skimr)
library(MASS)
library(moments)
library(mda)
library(caret)
library(gg3D)
library(plotly)
library("writexl")
```

# Separate data 
```{r}
male <- oph %>%
  filter(sex == "m")%>%
  dplyr::select(-id,-hornbill, -ID, -side, -sex, -ratio_colour_ub,
         -eccen_ub_mark, -convex_ub_mark, -sd_grey_ub,-mean_grey_ub,
         -solidity_ub_mark)

female <- oph %>%
  filter(sex == "f")%>%
  dplyr::select(-id,-ID, -hornbill, -side, -sex)
```

# Normality
## females
### Check normality for female
```{r}
#Cheers. Hair et al. (2010) and Bryne (2010) argued that data is considered to be normal if Skewness is between ‐2 to +2 and Kurtosis is between ‐7 to +7. 

female_skewness <- female %>%
  dplyr::select(-name)%>%
  skewness()

female_kurtosis <- female %>%
  dplyr::select(-name)%>%
  kurtosis()

female_skewness
female_kurtosis

#ratio_maj_min_lb -> negatively skewed 
#ratio_colour_c -> positively skewed 
#eccen_c_mark -> negatively skewed 
#mean_grey_lb -> positively skewed 
```

### Fix skewness 
```{r}
#https://www.datanovia.com/en/lessons/transform-data-to-normal-distribution-in-r/

female_normal<- female %>%
  #moderately negatively skewed ratio_maj_min_lb
  mutate(ratio_maj_min_lb = log10(max(ratio_maj_min_lb+1) - ratio_maj_min_lb))%>%
  #severely positively skewed ratio_colour_c
  mutate(ratio_colour_c =  1/(ratio_colour_c))%>%
  #severely  negatively skewed eccen_c_mark
  mutate(eccen_c_mark = log10(max(eccen_c_mark+1) - eccen_c_mark))%>%
  #moderately positively skewed mean_grey_lb
  mutate(mean_grey_lb =  log10(mean_grey_lb))

female_skewness_new <- female_normal %>%
  dplyr::select(-name)%>%
  skewness()

female_kurtosis_new <- female_normal %>%
  dplyr::select(-name)%>%
  kurtosis()

female_skewness_new
female_kurtosis_new

```

## Males 

### Visualization
```{r}
ggplot(data = male)+
  geom_histogram(aes(x = ratio_maj_min_c))

ggplot(data = male)+
  geom_histogram(aes(x = sqrt(ratio_maj_min_lb)))

ggplot(data = male)+
  geom_histogram(aes(x = ratio_colour_lb))

```

### Check normality 
```{r}
male_skewness <- male %>%
  dplyr::select(-name)%>%
  skewness()

male_kurtosis <- male %>%
  dplyr::select(-name)%>%
  kurtosis()

male_skewness
male_kurtosis

#ratio_maj_min_c -> positively skewed 
#ratio_maj_min_lb -> positively skewed
#ratio_colour_lb -> positively skewed 
#eccen_c_mark -> positively skewed 
#sd_grey_c -> positively skewed 
#sd_grey_lb -> positively skewed 
```

### Fix skewness 
```{r}

male_normal <- male %>%
  mutate(ratio_maj_min_lb = sqrt(ratio_maj_min_lb))%>%
  mutate(eccen_c_mark = 1/eccen_c_mark)%>%
  mutate(sd_grey_c = 1/sd_grey_c)%>%
  mutate(sd_grey_lb = 1/sd_grey_lb)
  
male_skewness_new <- male_normal %>%
  dplyr::select(-name)%>%
  skewness()

male_kurtosis_new <- male_normal %>%
  dplyr::select(-name)%>%
  kurtosis()

male_skewness_new
male_kurtosis_new

#should I look into box-cox transformation?
```

# Females LDA

The LDA algorithm starts by finding directions that maximize the separation between classes, then use these directions to predict the class of individuals. These directions, called linear discriminants, are a linear combinations of predictor variables.

LDA assumes that predictors are normally distributed (Gaussian distribution) and that the different classes have class-specific means and equal variance/covariance.

## Check the effect

```{r}
female_lda <- lda(name ~ ., data=female_normal)
female_variable <- female_normal %>%
  dplyr::select(-name)
female_lda
#solidity of the casque mark shows the most significant impact 
#the area of a particle divided by its convex hull area
```
## Test the accuracy 
```{r}
female_lda_p <- predict(female_lda, new_data = female_normal[,c(2:22)])$class
female_lda_p
table(female_lda_p, female_normal[,1])
mean(female_lda_p == female_normal$name)
# 100% prediction 
```
# 2D visualization 
```{r}
#convert list to dataframe 
female_lda_predict <- predict(female_lda)
female_lda_plot <- data.frame(name = female_normal[,1], LDA = female_lda_predict$x)

#plot 
ggplot(data = female_lda_plot, aes(x=LDA.LD1, y = LDA.LD2))+
  geom_point(aes(colour = name),
             size = 5,
             alpha = 0.5)+
  labs(title = "Females",
       x = "LDA1 (83%)",
       y = "LDA2 (17%)")+
  theme(plot.title = element_text(hjust = 0.5))

```

# Males LDA

## Check the effect 
```{r}
#check for the effect size 
male_lda <- lda(name ~ ., data=male_normal)
male_lda
```
## Test the model accuracy 
```{r}
#test the model 
male_lda_p <- predict(male_lda, new_data = male_normal[,c(2:16)])$class
male_lda_p
table(male_lda_p, male_normal[,1])
mean(male_lda_p == male_normal$name)
# 93% accuracy 
```

## 2D visualization 
```{r}
#convert list to dataframe 
male_lda_values <- predict(male_lda)
male_lda_plot <- data.frame(name = male_normal[,1], LDA = male_lda_values$x)

#plot 
ggplot(data = male_lda_plot, aes(x=LDA.LD1, y = LDA.LD2))+
  geom_point(aes(colour = name),
             size = 4,
             alpha = 0.5)+
  labs(title = "Males",
       x = "LDA1(60%)",
       y = "LDA2(22%)")+
  theme(plot.title = element_text(hjust = 0.5))

```
## 3D visualization 


```{r}
plot_ly(male_lda_plot, x = ~LDA.LD1, y = ~LDA.LD2, z = ~LDA.LD3,
        mode = "markers", color = ~name, showscale = TRUE) %>%
   layout(scene = list(xaxis = list(title = 'LDA1 (60%)'),
                                   yaxis = list(title = 'LDA2 (22%)'),
                                   zaxis = list(title = 'LDA3 (8%)')))
```
```{r}
male_lda_plot
write_xlsx(male_lda_plot,"3dplot.xlsx")
```


# Female Model finalization 

## Ver1
```{r}
#exclude the least significant variable 
female_normal_ver1 <- female_normal %>%
  select(-ratio_maj_min_c)

female_lda_ver1 <- lda(name ~ ., data=female_normal_ver1)
female_lda_ver1_p <- predict(female_lda_ver1, new_data = female_normal_ver1[,c(2:21)])$class
mean(female_lda_ver1_p == female_normal_ver1$name)
female_lda_ver1
```

## Ver2
```{r}
#exclude the least significant variable 
female_normal_ver2 <- female_normal_ver1 %>%
  select(-ratio_colour_c)

female_lda_ver2 <- lda(name ~ ., data=female_normal_ver2)
female_lda_ver2_p <- predict(female_lda_ver2, new_data = female_normal_ver2[,c(2:20)])$class
mean(female_lda_ver2_p == female_normal_ver2$name)
female_lda_ver2
```

## Ver3 
```{r}
#exclude the least significant variable 
female_normal_ver3 <- female_normal_ver2 %>%
  select(-eccen_lb_mark)

female_lda_ver3 <- lda(name ~ ., data=female_normal_ver3)
female_lda_ver3_p <- predict(female_lda_ver3, new_data = female_normal_ver3[,c(2:19)])$class
mean(female_lda_ver3_p == female_normal_ver3$name)
female_lda_ver3
```

## Ver4

```{r}
#exclude the least significant variable 
female_normal_ver4 <- female_normal_ver3 %>%
  select(-mean_grey_ub)

female_lda_ver4 <- lda(name ~ ., data=female_normal_ver4)
female_lda_ver4_p <- predict(female_lda_ver4, new_data = female_normal_ver4[,c(2:18)])$class
mean(female_lda_ver4_p == female_normal_ver4$name)
female_lda_ver4
```

## Ver5

```{r}
#exclude the least significant variable 
female_normal_ver5 <- female_normal_ver4 %>%
  select(-eccen_ub_mark)

female_lda_ver5 <- lda(name ~ ., data=female_normal_ver5)
female_lda_ver5_p <- predict(female_lda_ver5, new_data = female_normal_ver5[,c(2:17)])$class
mean(female_lda_ver5_p == female_normal_ver5$name)
female_lda_ver5
```
## Ver6
```{r}
#exclude the least significant variable 
female_normal_ver6 <- female_normal_ver5 %>%
  select(-sd_grey_ub)

female_lda_ver6 <- lda(name ~ ., data=female_normal_ver6)
female_lda_ver6_p <- predict(female_lda_ver6, new_data = female_normal_ver6[,c(2:16)])$class
mean(female_lda_ver6_p == female_normal_ver6$name)
female_lda_ver6
```

## Ver7
```{r}
#exclude the least significant variable 
female_normal_ver7 <- female_normal_ver6 %>%
  select(-ratio_colour_lb)

female_lda_ver7 <- lda(name ~ ., data=female_normal_ver7)
female_lda_ver7_p <- predict(female_lda_ver7, new_data = female_normal_ver7[,c(2:15)])$class
mean(female_lda_ver7_p == female_normal_ver7$name)
female_lda_ver7
```
## Ver8
```{r}
#exclude the least significant variable 
female_normal_ver8 <- female_normal_ver7 %>%
  select(-sd_grey_c)

female_lda_ver8 <- lda(name ~ ., data=female_normal_ver8)
female_lda_ver8_p <- predict(female_lda_ver8, new_data = female_normal_ver8[,c(2:14)])$class
mean(female_lda_ver8_p == female_normal_ver8$name)
female_lda_ver8
```
## Ver9
```{r}
#exclude the least significant variable 
female_normal_ver9 <- female_normal_ver8 %>%
  select(-convex_ub_mark)

female_lda_ver9 <- lda(name ~ ., data=female_normal_ver9)
female_lda_ver9_p <- predict(female_lda_ver9, new_data = female_normal_ver9[,c(2:13)])$class
mean(female_lda_ver9_p == female_normal_ver9$name)
female_lda_ver9
```
## Ver10
```{r}
#exclude the least significant variable 
female_normal_ver10 <- female_normal_ver9 %>%
  select(-ratio_maj_min_ub)

female_lda_ver10 <- lda(name ~ ., data=female_normal_ver10)
female_lda_ver10_p <- predict(female_lda_ver10, new_data = female_normal_ver10[,c(2:12)])$class
mean(female_lda_ver10_p == female_normal_ver10$name)
female_lda_ver10
```

## Ver11
```{r}
#exclude the least significant variable 
female_normal_ver11 <- female_normal_ver10 %>%
  select(-ratio_maj_min_lb)

female_lda_ver11 <- lda(name ~ ., data=female_normal_ver11)
female_lda_ver11_p <- predict(female_lda_ver11, new_data = female_normal_ver11[,c(2:11)])$class
mean(female_lda_ver11_p == female_normal_ver11$name)
female_lda_ver11
```


## Ver12
```{r}
#exclude the least significant variable 
female_normal_ver12 <- female_normal_ver11 %>%
  select(-sd_grey_lb)

female_lda_ver12 <- lda(name ~ ., data=female_normal_ver12)
female_lda_ver12_p <- predict(female_lda_ver12, new_data = female_normal_ver12[,c(2:10)])$class
mean(female_lda_ver12_p == female_normal_ver12$name)
female_lda_ver12
```

## Ver13
```{r}
#exclude the least significant variable 
female_normal_ver13 <- female_normal_ver12 %>%
  select(-mean_grey_lb)

female_lda_ver13 <- lda(name ~ ., data=female_normal_ver13)
female_lda_ver13_p <- predict(female_lda_ver13, new_data = female_normal_ver13[,c(2:9)])$class
mean(female_lda_ver13_p == female_normal_ver13$name)
female_lda_ver13
```

## Ver14
```{r}
#exclude the least significant variable 
female_normal_ver14 <- female_normal_ver13 %>%
  select(-mean_grey_c)

female_lda_ver14 <- lda(name ~ ., data=female_normal_ver14)
female_lda_ver14_p <- predict(female_lda_ver14, new_data = female_normal_ver14[,c(2:8)])$class
mean(female_lda_ver14_p == female_normal_ver14$name)
female_lda_ver14
```

## Ver15
```{r}
#exclude the least significant variable 
female_normal_ver15 <- female_normal_ver14 %>%
  select( -convex_lb_mark)

female_lda_ver15 <- lda(name ~ ., data=female_normal_ver15)
female_lda_ver15_p <- predict(female_lda_ver15, new_data = female_normal_ver15[,c(2:7)])$class
mean(female_lda_ver15_p == female_normal_ver15$name)
female_lda_ver15
```

## Ver16
```{r}
#exclude the least significant variable 
female_normal_ver16 <- female_normal_ver15 %>%
  select( -eccen_c_mark)

female_lda_ver16 <- lda(name ~ ., data=female_normal_ver16)
female_lda_ver16_p <- predict(female_lda_ver16, new_data = female_normal_ver16[,c(2:6)])$class
mean(female_lda_ver16_p == female_normal_ver16$name)
female_lda_ver16
```

## Ver17
```{r}
#exclude the least significant variable 
female_normal_ver17 <- female_normal_ver16 %>%
  select( -convex_c_mark)

female_lda_ver17 <- lda(name ~ ., data=female_normal_ver17)
female_lda_ver17_p <- predict(female_lda_ver17, new_data = female_normal_ver17[,c(2:5)])$class
mean(female_lda_ver17_p == female_normal_ver17$name)
female_lda_ver17
```

## Ver18
```{r}
#exclude the least significant variable 
female_normal_ver18 <- female_normal_ver17 %>%
  select( -solidity_c_mark)

female_lda_ver18 <- lda(name ~ ., data=female_normal_ver18)
female_lda_ver18_p <- predict(female_lda_ver18, new_data = female_normal_ver18[,c(2:4)])$class
mean(female_lda_ver18_p == female_normal_ver18$name)
female_lda_ver18
```

## Ver19
```{r}
#exclude the least significant variable 
female_normal_ver19 <- female_normal_ver18 %>%
  select(-ratio_colour_ub)

female_lda_ver19 <- lda(name ~ ., data=female_normal_ver19)
female_lda_ver19_p <- predict(female_lda_ver19, new_data = female_normal_ver19[,c(2:3)])$class
mean(female_lda_ver19_p == female_normal_ver19$name)
female_lda_ver19
```

## Ver20
```{r}
#exclude the least significant variable 
female_normal_ver20 <- female_normal_ver19 %>%
  select(-solidity_ub_mark)

female_lda_ver20 <- lda(name ~ ., data=female_normal_ver20)
female_lda_ver20_p <- predict(female_lda_ver20, new_data = female_normal_ver20[,c(2)])$class
mean(female_lda_ver20_p == female_normal_ver20$name)
female_lda_ver20
```
## Fin ver
```{r}
mean(female_lda_ver19_p == female_normal_ver19$name)
female_lda_ver19
#98% accuracy 
#includes solidity 
```
# Males Model finializaton 

## Ver1
```{r}
male_normal_ver1 <- male_normal %>%
  select(-ratio_colour_lb)
male_lda_ver1 <- lda(name ~ ., data=male_normal_ver1)
male_lda_ver1_p <- predict(male_lda_ver1, new_data = male_normal_ver1[,c(2:15)])$class
mean(male_lda_ver1_p == male_normal_ver1$name)
male_lda_ver1
```

## Ver2
```{r}
male_normal_ver2 <- male_normal_ver1 %>%
  select(-mean_grey_lb)
male_lda_ver2 <- lda(name ~ ., data=male_normal_ver2)
male_lda_ver2_p <- predict(male_lda_ver2, new_data = male_normal_ver2[,c(2:14)])$class
mean(male_lda_ver2_p == male_normal_ver2$name)
male_lda_ver2
```

## Ver3
```{r}
male_normal_ver3 <- male_normal_ver2 %>%
  select(-ratio_maj_min_ub)
male_lda_ver3 <- lda(name ~ ., data=male_normal_ver3)
male_lda_ver3_p <- predict(male_lda_ver3, new_data = male_normal_ver3[,c(2:13)])$class
mean(male_lda_ver3_p == male_normal_ver3$name)
male_lda_ver3
```

## Ver4
```{r}
male_normal_ver4 <- male_normal_ver3 %>%
  select(-sd_grey_c)
male_lda_ver4 <- lda(name ~ ., data=male_normal_ver4)
male_lda_ver4_p <- predict(male_lda_ver4, new_data = male_normal_ver4[,c(2:12)])$class
mean(male_lda_ver4_p == male_normal_ver4$name)
male_lda_ver4
```

## Ver5
```{r}
male_normal_ver5 <- male_normal_ver4 %>%
  select(-ratio_maj_min_c)
male_lda_ver5 <- lda(name ~ ., data=male_normal_ver5)
male_lda_ver5_p <- predict(male_lda_ver5, new_data = male_normal_ver5[,c(2:11)])$class
mean(male_lda_ver5_p == male_normal_ver5$name)
male_lda_ver5
#92%
```

## Ver6
```{r}
male_normal_ver6 <- male_normal_ver5 %>%
  select(-eccen_lb_mark)
male_lda_ver6 <- lda(name ~ ., data=male_normal_ver6)
male_lda_ver6_p <- predict(male_lda_ver6, new_data = male_normal_ver6[,c(2:10)])$class
mean(male_lda_ver6_p == male_normal_ver6$name)
male_lda_ver6
#87%
```