---
title: "R Notebook"
author: "Shuna Maekawa"
output:
  html_document:
    toc: true
    toc_float: true
---
# Import data 
```{r, message=FALSE, warning=FALSE}
#load the packages 
library(tidyverse)
library(tibble)
library(ggplot2)
library(vegan)
library(dplyr)
library(RVAideMemoire)
library(skimr)
library(ggpubr)
library(broom)
library(AICcmodavg)
```


```{r, message=FALSE, warning=FALSE}
#import data from the working directory 
oph <- read.csv("HornbillID.csv")
```



# Data wrangling 
```{r, message=FALSE, warning=FALSE}
#replace NA with 0
oph <- oph %>%
  replace(is.na(.), 0)
```

```{r, message=FALSE, warning=FALSE, results = "hide"}
# see the data structure 
glimpse(oph)
```
```{r}
#separate the dataframe

#male only
male <- oph %>%
  filter(sex == "m")

#female only
female <- oph %>%
  filter(sex == "f")
```

# Data analysis 

## Male NMDs
```{r, message=FALSE, warning=FALSE, results = "hide"}
# NMDs analysis for male hornbills
set.seed(46)
nmds_male <- male%>%
  # exclude non-numerical/empty variables
  select(-id,-name, -hornbill, -ID, -side, -sex, -ratio_colour_ub,
         -eccen_ub_mark, -convex_ub_mark, -sd_grey_ub,-mean_grey_ub,
         -solidity_ub_mark)%>%
  metaMDS(., distance="jaccard", k=3, trymax=50, trace = 2) 
```

```{r, message=FALSE, warning=FALSE, results = "hide"}
# see the data structure 
glimpse(nmds_male)
```

### biplot
```{r, message=FALSE, warning=FALSE}
#plot the raw result 
plot(nmds_male)
```
### stressplot
```{r, message=FALSE, warning=FALSE}
# shepherd plot 
# original dissimilarities are quite well preserved by the reduced dimensions
stressplot(nmds_male)
```
```{r, message=FALSE, warning=FALSE}
# add nmds scores to the original data
male_plot <-  nmds_male %>% 
  #extract nmds scores 
  scores(., choices = c(1:3), display=c("sites")) %>%
  #convert the nmds scores in a dataframe 
  data.frame(.) %>%
  bind_cols(male, .) %>%
  left_join(male)
```

### male NMDs plot
```{r, message=FALSE, warning=FALSE}

p1 <-  ggplot() +
  geom_point(data = male_plot, 
             aes(x=NMDS1, y=NMDS2, 
                 colour = name,
                 shape = side),
             size = 2,
             alpha = 0.7)

#stat_ellipse(data=males_plot, aes(x=NMDS1, y=NMDS2, colour = name), show.legend = F)


p1
```

### finn and tristan
```{r}
finn_tristan <- male_plot%>%
  filter(name == c( "finn", "tristan"))


p2 <- ggplot(data = finn_tristan, aes(x = NMDS1, y = NMDS2))+
  geom_point(aes(colour= name))
p2


```
### finn and toby
```{r}

finn_toby <- male_plot%>%
  filter(name == c( "finn", "toby"))
p3 <- ggplot(data = finn_toby, aes(x= NMDS1, y = NMDS2))+
  geom_point(aes(colour=name))
p3
```

## Female NMDs

```{r, message = FALSE, results = FALSE}
# NMDs analysis for female hornbills 
set.seed(37)
nmds_female <- female%>%
  # exclude non-numerical variables 
  select(-id,-name, -hornbill, -ID, -side, -sex)%>%
  metaMDS(., distance="jaccard", k=3, trymax=50, trace = 2) 
```

### stressplot
```{r}
stressplot(nmds_female)
```

```{r, message=FALSE, warning=FALSE}
# add nmds scores to the original data
female_plot <-  nmds_female %>% 
  #extract nmds scores 
  scores(., choices = c(1:3), display=c("sites")) %>%
  #convert the nmds scores in a dataframe 
  data.frame(.) %>%
  bind_cols(female, .) %>%
  left_join(female)
```

### female NMDs plot 
```{r, message=FALSE, warning=FALSE}
# plot the entire nmds 
p4 <-  ggplot() +
  geom_point(data = female_plot, 
             aes(x=NMDS1,
                 y=NMDS2, 
                 colour = name, 
                 shape = side),
             size = 2,
             alpha = 0.7)

#  stat_ellipse(data=female_plot,aes(x=NMDS1, y=NMDS2, colour = name), show.legend = F)
p4

```
### female boxplot 

```{r}
# plot NMDS1 values 
ggplot(data = female_plot, aes(x = id, y = NMDS1))+
  geom_boxplot(horizontal = TRUE)
```
#MANOVA 

## Female Analysis

### adonis
```{r}
#data prep
# dataframe that only contains variables
female_variables <- female_plot[,7:27]

set.seed(176)
#variable vs hornbills
adonis(female_variables ~ female_plot$name*female_plot$side, 
       permutation = 999)


```
Above shows that female individuals alone is statistically different, but sides are not (interaction model). 

### permanova 

```{r,message=FALSE, warning=FALSE}
set.seed(189)
#pairwise.perm.manova(fact = female_plot$name, resp = female_variables, nperm = 999)

```
Each female hornbill is statistically different from each other. 


## Male Analysis 

### adonis 
```{r,message=FALSE, warning=FALSE}
#data prep
# dataframe that only contains variables
male_variables <- male_plot[,7:27]%>%
  #remove unnecessary columns 
  select(-ratio_colour_ub, -eccen_ub_mark, -convex_ub_mark,
         -sd_grey_ub,-mean_grey_ub, -solidity_ub_mark)
set.seed(100)
#variable vs hornbills
adonis(male_variables ~ male_plot$name*male_plot$side, 
       permutation = 999)

```

hornbill*side may be statistically significant too (?) 

### permanova 
```{r,message=FALSE, warning=FALSE}
#hornbill individuals 
set.seed(12)
#pairwise.perm.manova(fact = male_plot$name, resp = male_variables, nperm = 999)


```

```{r,message=FALSE, warning=FALSE}
#hornbill sides
set.seed(10)
#pairwise.perm.manova(fact = male_plot$id, resp = male_variables, nperm = 999)

## Error in anova.mlm(lm(resp ~ fact), test = test) : residuals have rank 13 < 15
## cannot figure out what this 13 means, as the residual degree of freedom is 103-15-1 = 87 and is bigger than the number of dimensions
```

```{r}
#check for covarying variables
library(corrplot)
cor <- cor(male_variables)
corrplot(cor, method = "color")

```

```{r,message=FALSE, warning=FALSE}
male_nmds <- male_plot[, 28:30]
#pairwise.perm.manova(fact = male_plot$id, resp = male_nmds, nperm = 999)
```

```{r}
skim(female_variables)
```
```{r}
skim(male_variables)
```

# One way anova 

```{r}
summary(male)
```
### ratio_maj_min_c
```{r}
test_1 <- aov(ratio_maj_min_c ~ name, data = male)
summary(test_1)
# 0.094
```
### ratio_maj_min_ub
```{r}
test_2 <- aov(ratio_maj_min_ub ~ name, data = male)
summary(test_2)
# 0.118
```

### ratio_maj_min_lb
```{r}
test_3 <- aov(ratio_maj_min_lb ~ name, data = male)
summary(test_3)
# 0.068
```
### ratio_color_c
```{r}
test_4 <- aov(ratio_colour_c ~ name, data = male)
summary(test_4)
# significant 
```

