---
title: "R Notebook"
author: "Shuna Maekawa"
output:
  html_document:
    toc: true
    toc_float: true
---
# Import data 
```{r, message=FALSE, warning=FALSE}
#load the packages 
library(tidyverse)
library(tibble)
library(ggplot2)
library(vegan)
library(gganimate)
library(gifski)
library(transformr)
library(infer)
library(patchwork)
library(dplyr)
library(plyr)
library(cluster)    
library(factoextra) 
library(dendextend)
```


```{r, message=FALSE, warning=FALSE}
#import data from the working directory 
oph <- read.csv("HornbillID.csv")
```


```{r}
variables <- oph%>%
  select(ID,side,sex)

variables <- rename(variables, c("side"="side_dis", "sex"="sex_dis")) 
```

# Data wrangling 
```{r, message=FALSE, warning=FALSE}

#replace categorical variables with numerical variables \

#oph$side[oph$side == "r"] <- as.numeric("0")
#oph$side[oph$side == "l"] <- as.numeric("1")
#oph$sex[oph$sex == "m"] <- as.numeric("0")
#oph$sex[oph$sex == "f"] <-  as.numeric("1")

#oph$side <- as.numeric(oph$side)
#oph$sex <- as.numeric(oph$sex)

#replace NA with 0
oph <- oph %>%
  replace(is.na(.), 0)
```

```{r, message=FALSE, warning=FALSE, results = "hide"}
# see the data structure 
glimpse(oph)
```

# Data analysis 
```{r, message=FALSE, warning=FALSE, results = "hide"}
# NMDS 
nmds <- oph %>%
  # exclude non-numerical variables 
  select(-id,-name, -hornbill, -ID, -side, -sex)%>%
  metaMDS(., distance="jaccard", k=3, trymax=50, trace = 2) 
```

```{r, message=FALSE, warning=FALSE, results = "hide"}
# see the data structure 
glimpse(nmds)
```

# Data visualization 
```{r, message=FALSE, warning=FALSE}
#plot the raw result 
plot(nmds)
```
```{r, message=FALSE, warning=FALSE}
# shepherd plot 
# original dissimilarities are quite well preserved by the reduced dimensions
stressplot(nmds)
```
```{r, message=FALSE, warning=FALSE}
# add nmds scores to the original data
test <-  nmds %>% 
  #extract nmds scores 
  scores(., choices = c(1:3), display=c("sites")) %>%
  #convert the nmds scores in a dataframe 
  data.frame(.) %>%
  bind_cols(oph, .) %>%
  left_join(oph)
```

## all individuals
```{r, message=FALSE, warning=FALSE}
# plot the entire nmds 
p1 <-  ggplot() +
  geom_point(data = test, 
             aes(x=NMDS1, y=NMDS2, colour = id), 
             alpha = 0.7)


p1
```

## Sex

### males
```{r, message=FALSE, warning=FALSE}
#male vs female 
male <- test %>% 
  
  filter(sex == "m")

p2 <-  ggplot() +
  geom_point(data = male, 
             aes(x=NMDS1, y=NMDS2, colour = name, shape = side), 
             size = 3,
             position = "jitter")+
  labs(title = "Male")+
  theme(plot.title = element_text(hjust = 0.5))

p2
  
```

### females
```{r,message=FALSE, warning=FALSE}
#male vs female 

set.seed(36)
female <- test %>% 
  filter(sex == "f")

p3 <-  ggplot() +
  geom_point(data = female, 
             aes(x=NMDS1,
                 y=NMDS2,
                 colour = name,
                 shape= side), 
             size = 3)+
  labs(title = "Female")+
  theme(plot.title = element_text(hjust = 0.5))

#stat_ellipse(data=female, aes(x=NMDS1, y=NMDS2, colour = name), show.legend = F)
 
p3
```

## Sides

### Right
```{r,message=FALSE, warning=FALSE}
#side
right <- test %>% 
  filter(side == "r")

p4 <-  ggplot() +
  geom_point(data = right, 
             aes(x=NMDS1, y=NMDS2, colour = id,
                 shape = sex), 
             alpha = 0.7)+
  labs(title = "Right")+
  theme(plot.title = element_text(hjust = 0.5))

p4
```

### Left
```{r}
#side
left <- test %>% 
  filter(side == "l")

p5 <-  ggplot() +
  geom_point(data = left, 
             aes(x=NMDS1, y=NMDS2, colour = id,
                 shape = sex), 
             alpha = 0.7)+
  labs(title = "Left")+
  theme(plot.title = element_text(hjust = 0.5))

p5
```

## Each individual

### Oliver
```{r}
#individual
Oliver <- test %>% 
  filter(name == "oliver")

p6 <-  ggplot() +
  geom_point(data = Oliver, 
             aes(x=NMDS1, y=NMDS2, colour = id), 
             alpha = 0.7)+
  labs(title = "Oliver")+
  theme(plot.title = element_text(hjust = 0.5))+
  geom_text(data= Oliver,
            aes(x=NMDS1,y=NMDS2,label=hornbill),size=2.5,vjust=1)


p6
```

### Toby
```{r}
#individual
Toby <- test %>% 
  filter(name == "toby")

p7 <-  ggplot() +
  geom_point(data = Toby, 
             aes(x=NMDS1, y=NMDS2, colour = id), 
             alpha = 0.7)+
  labs(title = "Toby")+
  theme(plot.title = element_text(hjust = 0.5))+
  geom_text(data= Toby,
            aes(x=NMDS1,y=NMDS2,label=hornbill),size=2.5,vjust=1)


p7
```

### Rachel 
```{r}
#individual
Rachel <- test %>% 
  filter(name == "rachel")

p8 <-  ggplot() +
  geom_point(data = Rachel, 
             aes(x=NMDS1, y=NMDS2, colour = id), 
             alpha = 0.7)+
  labs(title = "Rachel")+
  theme(plot.title = element_text(hjust = 0.5))+
  geom_text(data= Rachel,
            aes(x=NMDS1,y=NMDS2,label=hornbill),size=2.5,vjust=1)


p8
```

### Finn
```{r}
#individual
Finn <- test %>% 
  filter(name == "finn")

p9 <-  ggplot() +
  geom_point(data = Finn, 
             aes(x=NMDS1, y=NMDS2, colour = id), 
             alpha = 0.7)+
  labs(title = "Finn")+
  theme(plot.title = element_text(hjust = 0.5))+
  geom_text(data= Finn,
            aes(x=NMDS1,y=NMDS2,label=hornbill),size=3,vjust=1)


p9
```

### Samantha 
```{r}
#individual
Samantha <- test %>% 
  filter(name == "samantha")

p10 <-  ggplot() +
  geom_point(data = Samantha, 
             aes(x=NMDS1, y=NMDS2, colour = id), 
             alpha = 0.7)+
  labs(title = "Samantha")+
  theme(plot.title = element_text(hjust = 0.5))+
  geom_text(data= Samantha,
            aes(x=NMDS1,y=NMDS2,label=hornbill),size=2.5,vjust=1)


p10
```

### Leo
```{r}
#individual
Leo <- test %>% 
  filter(name == "leo")

p11 <-  ggplot() +
  geom_point(data = Leo, 
             aes(x=NMDS1, y=NMDS2, colour = id), 
             alpha = 0.7)+
  labs(title = "Leo")+
  theme(plot.title = element_text(hjust = 0.5))+
  geom_text(data= Leo,
            aes(x=NMDS1,y=NMDS2,label=hornbill),size=2.5,vjust=1)


p11
```
### Aurora

```{r}
#individual
Aurora <- test %>% 
  filter(name == "aurora")

p12 <-  ggplot() +
  geom_point(data = Aurora, 
             aes(x=NMDS1, y=NMDS2, colour = id), 
             alpha = 0.7)+
  labs(title = "Aurora")+
  theme(plot.title = element_text(hjust = 0.5))+
  geom_text(data= Aurora,
            aes(x=NMDS1,y=NMDS2,label=hornbill),size=2.5,vjust=1)


p12
```

### Tristan 
```{r}
#individual
Tristan <- test %>% 
  filter(name == "tristan")

p13 <-  ggplot() +
  geom_point(data = Tristan, 
             aes(x=NMDS1, y=NMDS2, colour = id), 
             alpha = 0.7)+
  labs(title = "Tristan")+
  theme(plot.title = element_text(hjust = 0.5))+
  geom_text(data= Tristan,
            aes(x=NMDS1,y=NMDS2,label=hornbill),size=2.5,vjust=1)


p13
```

### Will 

```{r}
#individual
Will <- test %>% 
  filter(name == "will")

p14 <-  ggplot() +
  geom_point(data = Will, 
             aes(x=NMDS1, y=NMDS2, colour = id), 
             alpha = 0.7)+
  labs(title = "Will")+
  theme(plot.title = element_text(hjust = 0.5))+
  geom_text(data= Will,
            aes(x=NMDS1,y=NMDS2,label=hornbill),size=2.5,vjust=1)


p14
```

# ANOSIM Test

## all hornbills
```{r}
#make community matrix - extract columns with abundance information, turn data frame into matrix

#test_anosim = oph[,7:27]
#matrix_anosim= as.matrix(test_anosim)

#anosim = anosim(matrix, oph$id, distance = "euclidean", permutations = 9999)
#anosim

```


## male hornbills

```{r}

#male_test = male[,5:27]
#male_matrix = as.matrix(male_test)

#anosim_male = anosim(male_matrix, male$id, distance = "bray", permutations = 9999)
#anosim_male
```

## female hornbills
```{r}

#female_test = female[,5:27]
#female_matrix = as.matrix(female_test)

#anosim_female = anosim(female_matrix, female$id, distance = "bray", permutations = 9999)
#anosim_female


```

# Clustering Analysis 

## Agglomerative Hierarchical Clustering

```{r}
# exclude unnecessary columns 
hclust <- oph 
rownames(hclust) <- hclust$hornbill
hclust <- hclust %>%
  select(-id, -name, -ID, -hornbill, -side, -sex)

#convert table to dataframe 
hclust.df <- as.data.frame(hclust)
hclust.df <- scale(hclust.df)

# Dissimilarity matrix
hclust_dis <- dist(hclust.df, method = "euclidean")

# Hierarchical clustering using Complete Linkage
hclust_link <- hclust(hclust_dis, method = "complete" )

# Plot the obtained dendrogram
plot(hclust_link, cex = 0.6, hang = -1)
```

```{r}
# Compute with agnes
hc2 <- agnes(hclust.df, method = "complete")

# Agglomerative coefficient
hc2$ac
## 0.8534745

# methods to assess
m <- c( "average", "single", "complete", "ward")
names(m) <- c( "average", "single", "complete", "ward")

# function to compute coefficient
ac <- function(x) {
  agnes(hclust.df, method = x)$ac
}

map_dbl(m, ac)

```

```{r}
hc3 <- agnes(hclust.df, method = "ward")
pltree(hc3, cex = 0.6, hang = -1, main = "Dendrogram of agnes") 
```

## Divisive Hierarchical Clustering

```{r}
# compute divisive hierarchical clustering
hc4 <- diana(hclust.df)

# Divise coefficient; amount of clustering structure found
hc4$dc
## 0.8428457

# plot dendrogram
pltree(hc4, cex = 0.6, hang = -1, main = "Dendrogram of diana")
```

```{r}
# Ward's method
hc5 <- hclust(hclust_dis, method = "ward.D2" )

# Cut tree into 4 groups
sub_grp1 <- cutree(hc5, k = 17)
sub_grp2 <- cutree(hc5, k = 5)

# Number of members in each cluster
table(sub_grp1)

table(sub_grp2)

```

```{r}
plot(hc5, cex = 1)
rect.hclust(hc5, k = 17, border = 2:5)
```

```{r}
# Convert hclust into a dendrogram and plot
hc6 <- as.dendrogram(hc5)

nodePar <- list(lab.cex = 0.6, pch = c(NA, 19), 
                cex = 0.7, col = "blue")
plot(hc6,  
     xlab = "Height",
     nodePar = nodePar,
     horiz = TRUE)
```

```{r}
# Zoom in to the dendrogram
plot(hc6, xlim = c(10,-4), ylim = c(1,15),
     horiz = TRUE,
     nodePar = nodePar, 
     xlab = "Height")
```

### Cluster plot

```{r}
fviz_cluster(list(data = hclust.df, cluster = sub_grp2), geom = c("point"))

```
```{r}
fviz_cluster(list(data = hclust.df, cluster = sub_grp2), geom = c("text"))
```

## Female clusters 

#### Calculation

```{r}
# exclude unnecessary columns 
hclustf <- oph 
rownames(hclustf) <- hclustf$hornbill
hclust_f <- hclustf %>%
  filter(sex == 1)%>%
  select(-id, -name, -ID, -hornbill, -sex,-side)

#convert table to dataframe 
hclust_f.df <- as.data.frame(hclust_f)

#standardize the valeus 
hclust_f.df <- scale(hclust_f.df)

# Dissimilarity matrix
hclust_f_matrix <- dist(hclust_f.df, method = "euclidean")
# Ward's method
hc7 <- hclust(hclust_f_matrix, method = "ward.D2" )

# Cut tree into 4 groups
sub_grp3 <- cutree(hc7, k = 6)
sub_grp4 <- cutree(hc7, k = 3)

# Number of members in each cluster
table(sub_grp3)

table(sub_grp4)
```

#### Visualization
```{r}
hc8 <- as.dendrogram(hc7)

plot(hc8,  
     xlab = "Height",
     nodePar = nodePar,
     horiz = TRUE)

```
```{r}
# Zoom in to the dendrogram
plot(hc8, xlim = c(10,-4), ylim = c(1,15),
     horiz = TRUE,
     nodePar = nodePar, 
     xlab = "Height")
```
```{r}
# Zoom in to the dendrogram
plot(hc8, xlim = c(10,-4), ylim = c(15,30),
     horiz = TRUE,
     nodePar = nodePar)
```

```{r}
# Zoom in to the dendrogram
plot(hc8, xlim = c(10,-4), ylim = c(40,55),
     horiz = TRUE,
     nodePar = nodePar)
```

```{r}
fviz_cluster(list(data = hclust_f.df, cluster = sub_grp3), geom = c("text"))
```
```{r}
fviz_cluster(list(data = hclust_f.df, cluster = sub_grp3), geom = c("point"))
```
#Permanova Analysis 

```{r}
#https://rpubs.com/collnell/manova

matrix <- as.matrix(oph[,7:27])

dist<-vegdist(matrix, method='bray')
difference <-adonis2(dist~id, data=oph, permutations = 999, method="bray")
difference
```

