---
title: "Midterm Practice"
author: "Aditya Singhania"
date: "3/1/2022"
output:
  pdf_document:
    number_sections: true
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```

These are the packages that I will need for my solutions.

```{r}
library(tidyverse)
library(ggrepel)
library(Hmisc)
```

# Relation between tibbles, data frames, and ists

## (a) Tibble

```{r}
musicians_tbl <- tibble(
  name = c("Keith", "John", "Paul"),
  band = c("Stones", "Beatles", "Beatles"),
  instrument = c("guitar", "guitar", "bass")
)
```

## (b) Data frame

```{r}
musicians_dfr <- data.frame(
  name = c("Keith", "John", "Paul"),
  band = c("Stones", "Beatles", "Beatles"),
  instrument = c("guitar", "guitar", "bass")
)
```

## (c) List

```{r}
musicians_list <- list(
  name = c("Keith", "John", "Paul"),
  band = c("Stones", "Beatles", "Beatles"),
  instrument = c("guitar", "guitar", "bass")
)
```

## (d) Applying functions to all

```{r}
musicians <- Hmisc::llist(musicians_tbl, musicians_dfr, musicians_list)
sapply(musicians, function(x) {
  list(
    class = class(x),
    is_tib = is_tibble(x),
    is_df = is.data.frame(x),
    is_list = is.list(x)
  )
})
```

## (e) Attributes

```{r}
lapply(musicians, attributes)
```

## (f) Subsetting operations

```{r}
lapply(musicians, attributes)
```

## (g) Remove attributes

```{r}
remove_attr <- function(x) {
  attr(x, "class") <- NULL
  attr(x, "row.names") <- NULL
  # dropping appended "label" attribute
  attr(x,"label") <- NULL
  return(x)
}

musicians_cleaned <- lapply(musicians, remove_attr)

# to check attributes
# lapply(musicians_cleaned,attributes)

# Check whether identical

lapply(musicians_cleaned, function(x) identical(x,musicians_list))
```

## Map-family of functions

```{r}
mtcars
```

### (ai) Mean

```{r}
map(mtcars, mean)
```

### (aii) Type

```{r}
map(nycflights13::flights, typeof)
```

### (aiii) Is factor

```{r}
map(forcats::gss_cat, is.factor)
```

### (aiv) Unique values

```{r}
map_int(iris, n_distinct)
```

## (b) pmap

```{r}
pmap(list(10, 1:10, 2:11), runif)
```
What does pmap(list(10, 1:10, 2:11), runif) do? Why?

_Ans:_ `pmap()` allows us to specify a single list that will contain all the vectors that
we want to supply to any command or function. In this case, pmap allows us to accommodate 
3 vectors in a single list that will run with the runif() command, rather than 
using `runif()` and `map()` on each list 1) 10, 2) 1:10 and 3) 2:11 3 separate times.

# source: 
https://dcl-prog.stanford.edu/purrr-parallel.html#pmap



## (c) `map_dfr` `beaver1`, `beaver2`


```{r}
# Original function
map_dfr(
  list(beaver1, beaver2), function(dfr) {
    glm(activ ~ temp, data = dfr, family = binomial) |> pluck(coef)
  }
)

# Formula
map_dfr(
  list(beaver1, beaver2), ~ {
    glm(activ ~ temp, data = ., family = binomial) |> pluck(coef)
  }
)
```
Describe what this code chunk does.

_Ans:_ 
1. `map_dfr()` helps to bind the rows and columns that come from the two lists (beaver1
and beaver2) into a single dataframe.

2. Within the function, `glm()` calls for a generalized linear model to be fit to the dataframe.

3. (activ ~ temp) indicates that the line should be fitted based on the relationship
between temperature on activity. There will be two linear models due to it being 
applied to 2 datasets. 

4. (family = binomial) specifies that there is binomial data involved, we can see
that for the column `activ`, the values are either 0 (indicating no activity) or 1
(indicating activity).  

5. The coefficients of the two lines from beaver1 and beaver2 are retrieved, 
which includes the intercept in the 1st column and the gradient in the second column (temp). 

https://purrr.tidyverse.org/reference/map.html
https://data.princeton.edu/r/glms


# Tibbles with list columns

```{r}
data(starwars)
```

## (a) Names of all list columns

```{r}
starwars |>
  # Predicate functions must be wrapped in `where()`
  select(where(is.list)) |>
  names()
```

## (b) Milenium Falcon

```{r}
starwars |>
  select(name, starships) |>
  filter(map_lgl(starships, ~ "Millennium Falcon" %in% .)) |>
  select(name)

# Alternatively, and perhaps, more succintly
starwars$name[map_lgl(starwars$starships, ~ "Millennium Falcon" %in% .)]
```

## (c) Unique Films

```{r}
unique(unlist(starwars$films))
```

## (d) Feminine characters

```{r}
starwars_fem_percent <- unnest_longer(starwars, films) |>
  filter(!is.na(gender)) |>
  group_by(films) |>
  summarise(
    n_male = sum(gender == "masculine"),
    n_female = sum(gender == "feminine"),
    fem_percent = (n_female / (n_male + n_female)) * 100
  )
```

<https://stackoverflow.com/questions/41803446/group-data-frame-by-elements-from-a-variable-containing-lists-of-elements>

## (e) Plot feminine percentage by move release date

```{r}
release <- tribble(
  ~movies, ~year,
  "A New Hope", 1977,
  "The Empire Strikes Back", 1980,
  "Return of the Jedi", 1983,
  "The Phantom Menace", 1999,
  "Attack of the Clones", 2002,
  "Revenge of the Sith", 2005,
  "The Force Awakens", 2015
)

fem_by_year <-
  left_join(starwars_fem_percent, release, by = c("films" = "movies"))

ggplot(fem_by_year, aes(year, fem_percent, label = films)) +
  geom_point() +
  # fitting regression line 
  stat_smooth(method = "lm", col = "coral1") +
  geom_label_repel() +
  labs(
    y = "Feminine characters", 
    x = "Year", 
    title = "Gender ratio of Characters in Star Wars Movies", 
    caption = "Source: Star Wars API (https://swapi.dev/)"
  ) +
  # default scales::percent() multiplies its input value by 100, manually input scale value = 1
  scale_y_continuous(labels = scales::percent_format(scale = 1))
  
```

Source: https://thomasadventure.blog/posts/ggplot2-percentage-scale/

## (f) Comments

_Ans:_ There is a positive trend in the percentage of feminine characters
appearing onscreen with each Star Wars Movie released through the years from 1977 
(when a New Hope is released) to 2015 (The Force Awakens). 
We see one partciular outlier outside the general linear model which is from 
Attack of the Clones, where 33.3% of the characters were feminine characters.  
