---
title: "Zachary's karate-club network"
author: "Team Red"
date: "3/23/2022"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```

Loading the necessary packages:
```{r}
library(tidyverse)
library(ggraph)
library(igraph)
library(tidygraph)
library(ggforce)
library(concaveman)
```

(1) Import the data and create an undirected `tbl_graph` called `karate_netw`.
```{r}
karate_edges <- read_table("soc-karate.zip",
  col_names = FALSE, skip = 24
)

karate_nodes <- tibble(node = 1:34)

karate_netw <- tbl_graph(
  nodes = karate_nodes,
  edges = karate_edges,
  directed = FALSE
)
karate_netw
```
(2) How many nodes and edges are in the network?
```{r}
number_nodes <- gorder(karate_netw)
number_edges <- gsize(karate_netw)

number_nodes
number_edges
```
\Ans There are 34 nodes and 78 edges in the network. 

(3) Are there parallel edges or loops in the network?
```{r}
is_simple(karate_netw)
```
\Ans There are no parallel edges or loops in the network. 

(4) Node 1 is the karate instructor. Node 34 is the club president. Add a node attribute called `role` to the `tbl_graph` with the values `Instructor`, `President` and `Other`.
```{r}
karate_netw <- karate_netw |>
  activate(nodes) |>
  mutate(
    role = case_when(node == 1 ~ "Instructor",
                     node == 34 ~ "President",
                     TRUE ~ "Other"))

karate_netw
```

(5) Show that the two nodes with the highest degree are the president and the instructor.
```{r}
  karate_netw |>
  mutate(degree = centrality_degree()) |>
  filter(role != "Other")

```

(6) Find all shortest paths between the instructor and the president.
```{r}
all_shortest_paths(karate_netw, 1, 34)
```

(7) What is the local clustering coefficient of the instructor? What is the local clustering coefficient of the president? How do you interpret these numbers?
```{r}
  karate_netw |>
  mutate(local_transitivity = local_transitivity())  |>
  filter(role != "Other") 
```
##ANSWER
The fact that the clustering coefficient for the instructor and president is below 1, it shows that there are a larger number of triplets than triangles that can be formed. Since the instructor has a slightly higher local transitivity, it means that there are slightly more triangles that can be formed (that contain the instructor), compared to the president. 

(8) Which are the largest cliques in the network? Which nodes are in these cliques?
```{r}
largest_cliques(karate_netw)
```
\Ans One of the largest cliques in the network has the nodes 8, 1, 2, 3, 4, and the other largest clique in the network has the nodes 4, 1, 2, 3, 14.

(9) Confirm that the network is connected (i.e. there is only one component).
```{r}
count_components(karate_netw)
```
\Ans Since there is only one component, the network is connected.

(10) Which communities does the Louvain algorithm detect? Use **tidygraph** to add information about the communities to `karate_netw`.
```{r}
karate_netw <- 
  karate_netw |>
  activate(nodes) |>
  mutate(community_id = group_louvain())
karate_netw
```
(11) Make a plot similar to figure 1.
```{r}
karate_layout <- create_layout(karate_netw, layout = "stress")
layout_attributes <- attributes(karate_layout)
node_positions <- read_csv("karate_node_positions.csv")

karate_layout <-
  karate_layout |>
  select(-x, -y) |>
  left_join(node_positions, by = "node") |>
  select(x, y, everything())

attributes(karate_layout) <- layout_attributes

ggraph(karate_layout) +
  geom_edge_link() +
    geom_mark_hull(
    aes(
      x, 
      y, 
      fill = as.factor(community_id), 
      label = str_c("Community ", community_id)
      ), 
    colour = NA, 
    con.colour = "grey", 
    show.legend = FALSE
  ) +
  labs(
    title = "Zachary's Karate Club \nCommunities assigned by Louvain algorithm",
    caption = "Source: https://networkrepository.com"
  ) +
  geom_node_label(aes(label = node)) +
  geom_bezier() #???
```

(12) Briefly explain to a reader what the figure reveals about the karate-club network. 

\Ans From the figure, readers understand that there are 4 communities in the karate-club network. Community 2 has the original president (of the karate-club) and community 1 has the original instructor (community 1 has become his new organisation. From the black curve, we also know where the observed split happened between the members of the karate-club, the members that stayed in the old club were mainly from community 2 and 3 (with the exception of 9 and 10). 