---
title: "Solutions: Categorical data in the General Social Survey"
subtitle: "YSC2210 - DAVis with R"
author: "Michael T. Gastner"
date: "3 February 2022"
output:
  html_document:
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

For this exercise, we need the following libraries.

```{r, message=FALSE}
library(formattable)
library(kableExtra)
library(tidyverse)
```


# 'Race' of respondents

(a)

    ```{r}
    levels(gss_cat$race)
    ```

(b)

    ```{r}
    table(gss_cat$race)
    ```

(c)

    ```{r}
    nice_table <- function(gss_col_name, table_col_name) {
      table(gss_cat[[gss_col_name]]) |>
        as.data.frame() |>
        mutate(Freq = color_bar("lightgreen")(Freq)) |>
        kbl(
          col.names = c(table_col_name, "Frequency"),
          align = c("l", "r"),
          escape = FALSE
        ) |>
        kable_styling(c("striped", "condensed"), full_width = FALSE)
    }
    nice_table("race", "Race")
    ```

(d)

    ```{r}
    gss_cat$race_drop_unused <- fct_drop(gss_cat$race)
    levels(gss_cat$race_drop_unused)
    ```

    'Not applicable' is not a level of `"gss_cat$race_drop_unused"`.
(e)

    ```{r}
    nice_table("race_drop_unused", "Race")
    ```

    The row for 'Not applicable' has disappeared.
(f)

    ```{r}
    gss_cat$race_drop_unused <- fct_infreq(gss_cat$race_drop_unused)
    nice_table("race_drop_unused", "Race")
    ```


# Investigating the reported income

(a)

    ```{r}
    levels(gss_cat$rincome)
    ```

(b)

    ```{r}
    nice_table("rincome", "Reported income")
    ```
    
    The order of the rows is not intuitive.
    It would be more conventional to sort the income levels from smallest to
    largest.
    The non-numeric answers should be at the end.
    
    Another small issue is that some levels have the pattern '... to ...' and
    other levels '... - ...'.
    It would be nice to choose a uniform pattern.
    
    The abbreviation 'Lt' for 'Less than' is not self-explanatory.

(c)

     We could manually replace the levels with `" - "` by similarly named
     levels with `"to"`. However, in the following pipeline, I take a shortcut
     using `str_replace()` from the **stringr** package.
     
    ```{r}
    gss_cat$rincome_improved <-
      gss_cat$rincome |>
      fct_recode("Less than $1000" = "Lt $1000") |>
      fct_relevel("Not applicable", after = 0) |>
      fct_rev() |>
      fct_relabel(~ str_replace(.x, " - ", " to "))
    nice_table("rincome_improved", "Reported income")
    ```


# Party affiliation

(a)

    ```{r}
    levels(gss_cat$partyid)
    ```

(b)

    ```{r}
    gss_cat$partyid_aggregated <-
      gss_cat$partyid |>
      fct_collapse(
        Republican = c("Strong republican", "Not str republican"),
        Independent = c("Ind,near rep", "Independent", "Ind,near dem"),
        Democrat = c("Not str democrat", "Strong democrat"),
        Other = "Other party",
        Missing = c("No answer", "Don't know")
      ) |>
      fct_relevel("Republican", "Independent", "Democrat", "Other", "Missing")
    levels(gss_cat$partyid_aggregated)
    ```

(c)

    ```{r}
    nice_table("partyid_aggregated", "Party affiliation")
    ```


# Religion and denomination

(a)

    ```{r}
    gss_cat$is_protestant <- fct_other(gss_cat$relig, "Protestant")
    ```

(b)

    ```{r}
    gss_cat$has_denom <-
      fct_other(gss_cat$denom,
        keep = c(
          "No answer",
          "Don't know",
          "No denomination",
          "Not applicable"
        ),
        other_level = "Has a denomination"
      ) |>
      fct_relevel(c(
        "Has a denomination",
        "No denomination",
        "Not applicable",
        "Don't know"
      ))
    ```

(c)

    Conversion from two-way table to data frame requires
    `as.data.frame.matrix()`, see
    <https://stackoverflow.com/questions/47216508/create-two-way-table-and-change-to-data-frame-format>.

    ```{r}
    dfr <-
      table(gss_cat$has_denom, gss_cat$is_protestant) |>
      as.data.frame.matrix()
    ```

    I store the maxima of `Protestant` and `Other` in variables `max_prot` and
    `max_other`, respectively so that we can scale the column widths later on.
    
    ```{r}
    max_prot <- max(dfr$Protestant)
    max_other <- max(dfr$Other)
    ```

    The argument `color_bar_scale` in the next code chunk determines the
    overall width of the table.
    I aim to keep the bars relatively short to minimise data ink.
    
    ```{r}
    color_bar_scale <- 5e-4
    dfr |>
      mutate(
        Protestant = color_bar("lightgreen")(Protestant),
        Other = color_bar("lightblue")(Other)
      ) |>
      kbl(
        escape = FALSE,
        align = c("r", "r")
      ) |>
      kable_styling(c("striped", "condensed"), full_width = FALSE) |>
      column_spec(2, width = str_c(color_bar_scale * max_prot, "em")) |>
      column_spec(3, width = str_c(color_bar_scale * max_other, "em"))
    ```
