---
title: "Midterm Prep DAVis"
author: "shuna"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, fig.align = "centre")
```

```{r}
library(tidyverse)
library(ggplot2)
```

# L01 QR
How much do you remember from QR? (titanic and iris)

## Import Data
```{r}
titanic <- read.csv("titanic.csv")
```


## dimensions of the data 

```{r}
nrow(titanic)
ncol(titanic)
dim(titanic)
```
## unique 
```{r}
unique(titanic$class)
```
## table 
```{r}
table(titanic$class)
```

### two-way table 
```{r}
# rows, columns 
two_way_table <- table(titanic$survived, titanic$class)
two_way_table
```

## extract rows by column
```{r}
sec_class <- titanic[titanic$class == "2nd",]
```


## basic barplot 

```{r}
barplot(
  two_way_table,
  # change from stack to side-by-side plot
  beside = TRUE,
  main = "people on the Titanic",
  xlab = "Class",
  # not colour!!! col!!!
  col = c("lightgreen", "mediumpurple"),
  # change the x-axis to horizontal position
  las = 1
)

grid(
  # no line on x-axis 
  nx = NA,
  # add lines for y-axis 
  ny = NULL
)

legend(
  "topleft",
  fill = c("lightgreen", "mediumpurple"),
  legend = c("died", "survived")
)

abline(h = 0)
```

## mean by species 

```{r}
data(iris)
iris
```


```{r}
iris |>
  group_by(Species)|>
  summarize(Sepal.Length = mean(Sepal.Length))
```

```{r}
aggregate(
  Sepal.Length ~ Species,
  data = iris,
  FUN = mean
)
```

## use of par to facet 

```{r}
# row, column
par(mfrow = c(3,1))

iris_breaks <-seq(4.2, 8.0, 0.2)
iris_ylim <- c(0,12)
iris_xlab <- "Sepal length (cm)"

for (species in unique(iris$Species)){
  hist(
    iris$Sepal.Length[iris$Species == species],
    breaks = iris_breaks,
    ylim = iris_ylim,
    xlab = iris_xlab,
    main = species
  )
  grid()
}

par(mfrow = c(1,1))

```

## normal facted hist 
```{r}
ggplot(
  data = iris,
  aes(x = Sepal.Length)
) +
  geom_histogram(
    color = "black",
    fill = "grey",
    binwidth = 0.2
  ) +
  facet_wrap(ncol = 1, ~Species) +
  labs(
    x = "Sepal Length (cm)",
    y = "Frequency"
  ) +
  scale_x_continuous(breaks = seq(2, 9, 1)) +
  scale_y_continuous(breaks = seq(0, 16, 4))
```

# L02 Vector

## vector 

```{r}
1:10
```

## seq
```{r}
seq(3,36, by = 3)
```

## rep
```{r}
rep(c(2,3), each = 2)
```

## subsetting 

```{r}
x <- 1:10
y <- c(FALSE, TRUE)
x[y]
```
because x is longer than y, R recycles the elements of y. 
Only the elements of x at TRUE are filtered out. 

## Alphabets 
```{r}
letters[3]
LETTERS[4]
```
```{r}
c(rbind(letters,LETTERS))
```

```{r}
r <- rep(0:16, each = 4)
r
```

```{r}
r + 1:4
```

format(v, scientific = FALSE)

## sequence involving exponentiation 

```{r}
s <- seq(2,20, by = 2)
r <- rep(s, each = 2)
s
r
```

```{r}
exponent <- r - 0:1
exponent
```
```{r}
a <- (1:20)^exponent
a
```


```{r}
format(a, scientific = FALSE)
```

## continued fraction 


```{r}
n <- 1000
a <- c(2, rep(1,n-1))
b <- 1 / (1:n)
g <- contfrac::gconvergents(a, b, b0 = 2)
g$A[n] / g$B[n]
```

# L03 Vector 

Online Shopping 

## function 

```{r}
d_customers <- c(33L, 22L, 41L, 39L, 38L, 46L, 47L)
d_customers
```

```{r}
d_sales <- c(56, 67, 75, 89, 98, 87, 75)
```

```{r}
weekly_profit <- function(d_customers, d_sales){
  sales_per_day <- d_customers * d_sales
  sum(sales_per_day)
}

weekly_profit(d_customers, d_sales)
```

## which 

```{r}
max(d_sales)
```

```{r}
which.max(d_sales)
```


# L04 

## Quote 

An election for Mayor of New York City was held on November 7, 2017.

>"Incumbent Democrat Bill de Blasio won reelection to a second term with 66.2% of the vote against Republican Nicole Malliotakis." ^[Quote from <https://en.wikipedia.org/wiki/2017_New_York_City_mayoral_election>, accessed on 28 January 2022]

## kable table

```{r}
library(kableExtra)
#https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_html.html#Table_Styles
```

```{r}
ny <- read_csv("ny_mayor.csv")
# add a row for sum of each column
ny_with_total<- ny %>%
  # reference :<https://stackoverflow.com/questions/4946873/add-row-to-a-data-frame-with-total-sum-for-each-column>
  adorn_totals("row")
# convert it to a data frame 
ny <- as.data.frame(ny_with_total)
```

```{r}
# making a table
ny %>%
  # add a caption
  kbl(
    caption = "New York Mayoral Election 2017",
    # add commas for numbers
    format.args = list(big.mark = ",")
  ) %>%
  # group columns
  add_header_above(c(
    " " = 1,
    "Candidate" = 3
  )) %>%
  # add footnote
  footnote(
    general = "<small>Source: CNN</small>",
    # remove title
    general_title = "",
    # keep the footnote small
    fixed_small_size = TRUE,
    # title will not be excaped against HTML
    escape = FALSE
  ) %>%
  # add alternative row colors
  kable_paper("striped",
    # reduce the size
    full_width = F
  ) %>%
  # add a solid line for total
  row_spec(5,
    extra_css = c("border-bottom: 1px solid black")
  )
```

```{r}
ny_mayor <- read_csv("ny_mayor.csv")

# I use bind_rows to add the total votes as bottom row of the table. The idea
# is from
# https://stackoverflow.com/questions/4946873/add-row-to-a-data-frame-with-total-sum-for-each-column
# (accessed on 19 January 2022).
# The idea for the solid horizontal line above the row for the total votes
# is from
# https://stackoverflow.com/questions/53655983/row-spec-function-from-kableextra-does-not-create-a-horizontal-line-in-html-ou
# I make 'CNN' in the caption a clickable hyperlink.
ny_mayor |>
  bind_rows(summarise(
    ny_mayor,
    across(where(is.numeric), sum),
    across(where(is.character), ~"Total")
  )) |>
  rename_with(str_to_title) |>
  rename("de Blasio" = "De_blasio") |>
  kbl(
    format.args = list(big.mark = ","),
    caption = "New York Mayoral Election 2017"
  ) |>
  kable_styling(
    bootstrap_options = c("striped", "condensed"),
    full_width = FALSE
  ) |>
  
  add_header_above(c(" " = 1, "Candidate" = 3)) |>
  # add a solid line 
  row_spec(nrow(ny_mayor), extra_css = "border-bottom: 1px solid") |>
  # footnote 
  # hyperlink 
  footnote(
    general = str_c(
      "<small>Source: <a href=",
      "https://edition.cnn.com/election/2017/results/new-york-city-mayor",
      ">CNN</a></small>"
    ),
    escape = FALSE,
    general_title = ""
  )
```

# L05 If_else
Practicing R programming 


## if_else
```{r}
r_function <- function(x){
  if_else(
    x == 0,
    "it is zero",
    if_else(
      x > 0,
      "it is bigger than zero",
      NA_character_
    )
  )
}

r_function(10)
```

## case_when
```{r}
case_when_function <- function(x){
  case_when(
    x == 0 ~ "it is zero",
    x == 1 ~ "it is one", 
    x == 2 ~ "it is two",
    TRUE ~ NA_character_
  )
}
case_when_function(10)
```

## recode

```{r}
third_function <- function(x){
  recode(
    x, 
    "1" = "it is one",
    "2" = "it is two",
    .default = NA_character_
  )
}

third_function(3)
```

## overflow & underflow 

check r_programming_solutions 

## floating point 

Floating point numbers are only represented with finite precision on a computer, following the IEEE 754 standard. The internal representation of these numbers is in binary format. Numbers that have an exact decimal representation cannot be represented accurately as a binary number with finite precision. Consequently, there are round-off errors, which lead to 0.1 + 0.2 and 0.3 to differ in the least significant bias 

Instead of ==,  we should use `all.equal()`, which is designed to test whether two objects are equal or almost equal. 


# L06 factors 
General Social Survey 

```{r}
gss_cat$is_protestant <- fct_other(gss_cat$relig, "Protestant")
gss_cat
```

```{r}
dfr <-
  table(gss_cat$has_denom, gss_cat$is_protestant) |> 
  as.data.frame.matrix()
```

```{r}
library(formattable)
```

```{r}
dfr |> 
  mutate(
    Protestant = color_bar("lightgreen")(Protestant),
    Other = color_bar("lightblue")(Other)
  ) |> 
  kbl(
    escape = FALSE,
    align = c("r", "r")
  ) |>
  kable_styling(c("striped", "condensed"), full_width = FALSE)
```
# L07 dplyr 

Dow Jons monthly performance 

```{r}
library(lubridate)
library(readxl)
```
```{r}
dj <- read_excel("dja-performance-report-monthly.xls", skip = 4)
dj <- dj[!(is.na("Effective Date"))]
```

```{r}
dj <- read_excel("dja-performance-report-monthly.xls", skip = 4) |> 
  filter(!is.na("Effective Date")) |> 
  select(date = "Effective Date",
         close_value = "Close Value") |> 
  mutate(month = mdy(date) |> month())

dj
```

```{r}
dj <- dj |>
  mutate(
    year = mdy(date) |> year(),
    next_month_in_dj = lead(month),
    year_of_next_month = lead(year),
    next_month = if_else(month < 12, month+1, 1),
    year_of_next_month_in_c = year + (month == 12)
  ) |> 
  slice_head(n = nrow(dj))

dj
```

# L08 ggplot 
bubble chart - Preston curve 

```{r}
library(ggrepel)
lq <- read.csv("life_quality.csv")
show_label <- 
  rank(lq$gdp_per_capita) <= 5 |
  rank(-lq$gdp_per_capita) <= 5 |
  rank(lq$life_expectancy) <= 5 |
  rank(-lq$life_expectancy) <= 5 |
  rank(lq$pop) <= 5

lq$country_label <- if_else(show_label, lq$country_name, "")
```

```{r, fig.width=7.5, fig.height=5.0}
g <- ggplot(data = lq, aes(
  gdp_per_capita,
  life_expectancy,
  size = pop,
  colour = continent
)) + 
  geom_point(alpha = 0.5) + 
  # labels 
  labs(
    x = "GDP",
    y = "Life Expectancy", 
    title = "Yay",
    caption = "Source: World Bank"
  ) + 
  theme(plot.title = element_text(hjust = 0.5))+
  # scaling 
  scale_x_log10(labels = dollar,
                minor_breaks = 2:9 %o% 10^(2:5)) + 
  # size of the symbols 
  scale_size_area(
    name = "Population",
    breaks = 10^(6:9),
    labels = c("1 million", "10 million", "100 million", "1 billion"),
    max_size = 20
  ) + 
  # colour of the symbols 
  scale_color_brewer(name = "Continent", palette ="Set1") +
  # override legends 
  guides(colour = guide_legend(override.aes = list(alpha = 1, size = 4))) +
  # abline 
  geom_smooth(
    aes(weight = pop),
    colour = "grey",
    method = "loess",
    show.legend = FALSE
  ) + 
  # expand ylim so that more space 
  ylim(NA,86) +
  geom_text_repel(
    aes(label = country_label,
        size = 1e6 * log(pop)),
    min.segment.length = 0.25,
    max.overlaps = 20,
    show.legend = FALSE
  )
g
```
```{r}
2:9 %o% 10^(2:6)
```

# L09 Measles 

```{r}
measles <- read.csv("measles_for_ggplot2.csv")
```

```{r}
p <- ggplot(
  measles, 
  aes(year+0.5, state,fill = incidence + 1e-6)
) + 
  # add space around the tiles 
  geom_tile(
    width = 0.95,
    height = 0.95,
    colour = "white"
  ) + 
  labs(
    x = "year",
    y = "state",
    title = "Measles cases in US"
  ) + 
  scale_x_continuous(
    breaks = seq(1930, 2000, by = 10),
    expand = expansion(0)
  ) + 
  # invert the y-axis in alphabetical order
  scale_y_discrete(limits = rev) +
  # deal with the color and legend 
  scale_fill_distiller(
    name = "Incidence rate\n(per 100,000)",
    # breaks in legend
    breaks = 10^(-2:3),
    # set legend range
    limits = c(2e-3, NA),
    # use actual numbers 
    labels = prettyNum,
    trans = "log10", 
    oob = scales::squish,
    palette = "Reds",
    # invert the gradation 
    direction = 1,
    guide = guide_colourbar(
      frame.colour = "black",
      ticks.colour = "black",
      barheight = 10,
      # order = 1 for the NA legend 
      order = 1
    )
  ) + 
  geom_tile(
    data = subset(measles, is.na(incidence)),
    aes(colour = "Missing data"),
    width = 0.9,
    fill = "gray"
  ) + 
  scale_colour_manual(
    name = NULL,
    # outline for the legend and tiles
    values = "white",
    guide_legend(order = 2)
  ) + 
  geom_vline(xintercept = 1963, size = 1) + 
  annotate(
    "text",
    label = "Vaccine introduced",
    x = 1963.3,
    y = 53.5,
    # in the plot
    # if 0, on the plot 
    vjust = 1,
    hjust = 0
  ) +
  theme(
    axis.text.y = element_text(size = 6),
    panel.background = element_rect(fill = NA),
    panel.grid.major.x = element_line(size = 1),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.ontop = TRUE,
    plot.title = element_text(face = "bold")
    
  )
  
p
```
# R practice 

```{r, message=FALSE}
library(MASS)
library(tidyverse)
library(scales)
```

## 4 factors 

```{r}
answers <- factor(c(rep("yes", 6), rep("no", 6)))

answers <- factor(answers, levels = c("Yes", "no"))
table(answers)
answers

```
factors not in levels will be NAs 

## 5 ggplot 

```{r}
diamonds
```


```{r}
ggplot(data = diamonds, aes(x = carat, y = price)) + 
  facet_wrap( ~color,
              nrow = 2,
              ncol = 4) +
  labs(x = "Carat",
       y = "Price(US$)",
       title = "Relation between Price, Carat and COlour of Diamonds",
       caption = "Source: R package 'ggplot2'") + 
  theme(plot.title = element_text(hjust = 0.5)) + 
  geom_bin_2d()+
  geom_smooth(data=select(diamonds, -color),method = 'lm', color = "black") + 
  scale_x_log10(minor_breaks = 1:6 %o% 10^(-3:-1)) + 
  scale_y_log10(
    minor_breaks = 1:5 %o% 10^(1:5),
                labels = comma) + 
  scale_fill_distiller(palette = "Reds",
                       direction = 1,
                       name = "Count") + 
  theme(
    legend.position = c(0.9, 0.15),
    legend.title = element_text(size = 6),
    legend.text = element_text(size = 6)
  )
```

## masses 

```{r}
masses <- as.tibble(Animals, rownames = "species")
masses$group <-
  fct_collapse(masses$species,
    "Dinosaurs" = c("Dipliodocus", "Triceratops", "Brachiosaurus"),
    other_level = "Mammals"
  )

masses$group <- factor(masses$group, c("Mammals", "Dinosaurs"))
levels(masses$group)
```

```{r}
ggplot(masses, aes(x = body, y = brain, shape = group, color = group)) + 
  geom_point() +
  theme(legend.title =element_blank())+
  scale_x_log10(labels=trans_format('log10',math_format(10^.x)),
                breaks = 10^(-1:5),
                minor_breaks = NULL)+
  scale_y_log10(labels=trans_format('log10',math_format(10^.x)),
                minor_breaks = NULL) + 
  coord_fixed(ratio = 1)+
  scale_colour_brewer(palette = "Dark2") + 
  guides(colour = guide_legend(override.aes = list(size = 4)))
```
```{r}
asean <- read.csv("asean.csv")
asean <- asean |>
  mutate(per = scales::percent(million_visits/sum(million_visits),
                               accuracy = 0.1))
asean
```

```{r}
ggplot(data = asean, aes(y = million_visits, x = reorder(country, +million_visits)))+
  geom_bar(stat = 'identity') + 
  coord_flip() + 
  ylim(0, 21)+
  geom_text(aes(label = per), 
            hjust = -0.2,
            ) +
  scale_y_continuous(expand = c(0,0),
                     limits = c(-0.1,22))
```
```{r}
?scales::percent
```

