---
title: "New York Mayoral Election 2017"
author: "Team Spirit"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    toc: true
---

```{r setup, include=FALSE}
library(kableExtra)
library(scales)
library(tidyverse)
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
```


# Introduction

In this document, we look at different ways to present data for the 2017
election for Mayor of New York City.
As Wikipedia[^wikipedia] notes,

> 'Incumbent Democrat Bill de Blasio won reelection to a second term with
  66.2% of the vote against Republican Nicole Malliotakis.'

[^wikipedia]:
  <https://edition.cnn.com/election/2017/results/new-york-city-mayor>,
  accessed on 19 January 2022.
  
We consider the number of votes in each borough.


# Data presented as a table

```{r nyTable}
library(kableExtra)
library(tidyverse)
ny_mayor <- read_csv("ny_mayor.csv")

# I use bind_rows to add the total votes as bottom row of the table. The idea
# is from
# https://stackoverflow.com/questions/4946873/add-row-to-a-data-frame-with-total-sum-for-each-column
# (accessed on 19 January 2022).
# The idea for the solid horizontal line above the row for the total votes
# is from
# https://stackoverflow.com/questions/53655983/row-spec-function-from-kableextra-does-not-create-a-horizontal-line-in-html-ou
# I make 'CNN' in the caption a clickable hyperlink.
ny_mayor |>
  bind_rows(summarise(
    ny_mayor,
    across(where(is.numeric), sum),
    across(where(is.character), ~"Total")
  )) |>
  rename_with(str_to_title) |>
  rename("de Blasio" = "De_blasio") |>
  kbl(
    format.args = list(big.mark = ","),
    caption = "New York Mayoral Election 2017"
  ) |>
  kable_styling(
    bootstrap_options = c("striped", "condensed"),
    full_width = FALSE
  ) |>
  add_header_above(c(" " = 1, "Candidate" = 3)) |>
  row_spec(nrow(ny_mayor), extra_css = "border-bottom: 1px solid") |>
  footnote(
    general = str_c(
      "<small>Source: <a href=",
      "https://edition.cnn.com/election/2017/results/new-york-city-mayor",
      ">CNN</a></small>"
    ),
    escape = FALSE,
    general_title = ""
  )
```


# Data presented as a bar plot

```{r nyBarPlot, fig.width=4, fig.height=3, fig.align='center', out.width='50%'}
ny_mayor_long <- read_csv("ny_mayor.csv") |>
  mutate(borough = str_replace(borough, " ", "\n")) |>
  pivot_longer(-borough) |>
  mutate(name = fct_rev(name))
ggplot(ny_mayor_long, aes(borough, value, fill = name)) +
  geom_col(position = "dodge") +
  scale_x_discrete(limits = rev) +
  coord_flip() +
  scale_y_continuous(labels = label_comma()) +
  scale_fill_manual(
    labels = c("Other", "Malliotakis", "de Blasio"),
    values = c("palegreen4", "firebrick2", "steelblue")
  ) +
  labs(
    title = "New York Mayoral Election 2017",
    caption = "Source: CNN",
    x = "Borough",
    y = "Votes",
    fill = "Candidate"
  ) +
  guides(fill = guide_legend(reverse = TRUE)) +
  theme(legend.key.size = unit(1, "lines"))
```


# What is better: table or bar plot?

The table contains the exact numbers of votes cast in each borough, whereas
it is impossible to retrieve exact numbers from the bar plot.
It would be possible to print the numbers inside the bars with the
`geom_bar_text()` function from the __ggfittext__ package to overcome this
disadvantage.

The table has the disadvantage that reading numbers is a slower process than
judging the lengths of bars in the bar plot.
The table could be improved by adding bars under each number with the help of
the __formattable__ package.
