---
title: "General Social Survey, Team A"
author: "Team A"
date: "2/9/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

We have loaded the following packages for this exercise.

```{r}
library(forcats)
library(tidyverse)
library(dplyr)
library(kableExtra)
library(formattable)
```

## Exercise 1

(a)
    ```{r}
    levels(gss_cat$race)
    ```

(b)
    ```{r}
    table(gss_cat$race)
    ```

(c)
    ```{r}
    table(gss_cat$race) |>
      as.data.frame() |>
      mutate(Freq = color_bar("lightgreen")(Freq)) |>
      kbl(
        col.names = c("Race", "Frequency"),
        align = c("l", "r"),
        escape = FALSE
      ) |>
      kable_styling(c("striped", "condensed"),
        full_width = FALSE
      )
    
        nice_table <- function(column, name) {
      table(column) |>
        as.data.frame() |>
        mutate(Freq = color_bar("lightgreen")(Freq)) |>
        kbl(
          col.names = c(name, "Frequency"),
          align = c("l", "r"),
          escape = FALSE
        ) |>
        kable_styling(c("striped", "condensed"),
          full_width = FALSE
        )
    }

    nice_table(gss_cat$race, "Race")
    ```

(d)
    ```{r}
    gss_cat$race_drop_unused <- fct_drop(gss_cat$race)
    nlevels(gss_cat$race_drop_unused)
    ```

(e)
    ```{r}
    nice_table(gss_cat$race_drop_unused, "Race")
    ```

(f)
    ```{r}
    race_sorted <- fct_infreq(gss_cat$race_drop_unused)
    levels(race_sorted)
    nice_table(race_sorted, "Race")
    ```

## Exercise 2

(a)
    ```{r}
    levels(gss_cat$rincome)
    ```

(b)
    ```{r}
    nice_table(gss_cat$rincome, "Income levels") 
    ```
This table would not be suited to be a published report because of 
several important reasons:

1. Majority of the data are in the "$25000 or more" or "Not applicable" 
categories, and are not clearly represented. Instead, smaller categories 
could be created to break up the "$25000 or more" category to give viewers
more insight.

2. The "income levels" column lacks comma separators in the numbers
(e.g. $5000 than $5,000).

3. The income labels are inconsistent. Some use "to" and some use "-" 
to indicate a range. For example, one label is "$15000 - $19999", while
another is "$7000 to $7999" which is inconsistent.

4. The range sizes are inconsistent. For example, one range is, 
"$1000 to $2999". This range size is $1999, while another range 
"$7000 to $7999" has a range size of $999. Similar inconsistencies
are present between other levels, and does not provide a legitimate
base for comparison.

5. The income level "Lt $1000" has "Lt", which means "lower than" but 
this is not a widely known abbreviation and may be confusing to readers.
Instead, it should just say "$1000 or lower".

6. "No answer", "Don't know", "Not applicable" and "Refused" can be 
subsumed together into one level, because each one does not provide any 
important information individually.

(c)
    ```{r}
    levels(gss_cat$rincome)
    
    # Fixing ranges, standaridisng labels, adding commas
    gss_cat_rincome_recode <-
      fct_recode(gss_cat$rincome,
        "$0 to 4,999" = "Lt $1000",
        "$0 to 4,999" = "$1000 to 2999",
        "$0 to 4,999" = "$3000 to 3999",
        "$0 to 4,999" = "$4000 to 4999",
        "$5,000 to 9,999" = "$5000 to 5999",
        "$5,000 to 9,999" = "$6000 to 6999",
        "$5,000 to 9,999" = "$7000 to 7999",
        "$5,000 to 9,999" = "$8000 to 9999",
        "$10,000 to 14,999" = "$10000 - 14999",
        "$15,000 to 19,999" = "$15000 - 19999",
        "$20,000 to 24,999" = "$20000 - 24999",
        "$25,000 or more" = "$25000 or more",
        Missing = "No answer",
        Missing = "Don't know",
        Missing = "Refused"
      )
    
    # Ordering properly
    gss_cat_rincome_recode <-
      fct_relevel(
        gss_cat_rincome_recode,
        c(
          "$25,000 or more",
          "$20,000 to 24,999",
          "$15,000 to 19,999",
          "$10,000 to 14,999",
          "$5,000 to 9,999",
          "$0 to 4,999",
          "Missing"
        )
      )
    
    levels(gss_cat_rincome_recode)
    
    nice_table(gss_cat_rincome_recode, "Income")
    ```
Even though we have improved it, most of the information is still concentrated
in the "$25000 or more" category, and the "Missing" category. It would be better
if we could have more categories that break up the "$25000 or more" category. 

## Exercise 3

(a)
    ```{r}
    levels(gss_cat$partyid)
    ```

(b)
    ```{r}
    gss_cat$partyid_aggregated <- fct_collapse(
      gss_cat$partyid,
      Republican = c(
        "Strong republican",
        "Not str republican"
      ),
      Independent = c(
        "Ind,near rep", "Independent",
        "Ind,near dem"
      ),
      Democrat = c(
        "Not str democrat",
        "Strong democrat"
      ),
      Other = c("Other party"),
      Missing = c("No answer", "Don't know")
    )
    
    # Ordering properly
    gss_cat$partyid_aggregated <-
      fct_relevel(
        gss_cat$partyid_aggregated,
        c(
          "Republican",
          "Independent",
          "Democrat",
          "Other",
          "Missing"
        )
      )
    ```

(c)
    ```{r}
    nice_table(gss_cat$partyid_aggregated, "Party Affiliation")
    ```
    
## Exercise 4

(a)
    ```{r}
    gss_cat$is_protestant <- fct_other(gss_cat$relig, keep = c("Protestant"))
    nice_table(gss_cat$is_protestant, "Religion")
    ```

(b)
    ```{r}
    gss_cat$has_denom <- fct_other(gss_cat$denom,
      keep = c(
        "No denomination",
        "Not applicable",
        "No answer",
        "Don't know"
      )
    )
    levels(gss_cat$has_denom) <- c(
      "Has a denomination", "No denomination",
      "Not applicable", "No answer", "Don't know"
    )

    # Ordering correctly
    gss_cat$has_denom <- fct_relevel(gss_cat$has_denom,
      c(
        "Has a denomination", "No denomination", "Not applicable",
        "Don't know", "No answer"
      ),
      after = 0L
    )

    nice_table(gss_cat$has_denom, "Denomination")
    ```

(c)
    ```{r}
    kbl(table(gss_cat$has_denom, gss_cat$is_protestant)) |>
      kable_styling(
        bootstrap_options = "striped", full_width = F, position =
          "left"
      )
    ```