---
title: "Network Data"
author: "shuna"
date: "3/17/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message= FALSE) 

```

```{r}
library(tidyverse)
demo_nodes <- read_csv("demo_nodes.csv")
demo_edges <- read_csv("demo_edges.csv")
```

```{r}
demo_nodes
demo_edges
```
```{r}
library(tidygraph)
```

```{r}
demo_undirected_netw <- tbl_graph(
  nodes = demo_nodes,
  edges = demo_edges,
  directed = FALSE
)
demo_undirected_netw
```
```{r}
class(demo_undirected_netw)
```

```{r}
active(demo_undirected_netw)
```

```{r}
demo_undirected_netw |>
  activate(edges)|>
  as_tibble()
```

```{r}
email_nodes <- read_csv("email_nodes.csv")
email_edges <- read_csv("email_edges.csv")
email_directed_netw <- tbl_graph(
nodes = email_nodes,
edges = email_edges
)
```

```{r}
library(igraph)
```

```{r}
is_directed(email_directed_netw)
is_directed(demo_undirected_netw)
```


```{r}
# the number of nodes in the network
gorder(demo_undirected_netw)
# the number of edges in the network 
gsize(demo_undirected_netw)

```
```{r}
library(ggplot2)
library(ggraph)
```


```{r}
ggraph(demo_undirected_netw) +
  geom_edge_link() +
  geom_node_label(aes(label = name))
```
```{r}
ggraph(demo_undirected_netw, layout = "kk") +
  geom_edge_link() +
  geom_node_label(aes(label = name))
```

```{r}
layout <- create_layout(demo_undirected_netw, layout = "kk")
ggraph(layout) +
geom_edge_link() +
geom_node_label(aes(label = name))

layout
```


```{r}
layout$y[layout$name == "K"] <- 0.5
ggraph(layout) +
geom_edge_link() +
geom_node_label(aes(label = name))
```

```{r}
library(tidyverse)
ggraph(demo_undirected_netw, layout = "kk") +
  geom_edge_bend(strength = 0.8) +
  geom_node_circle(
    aes(r = if_else(name < "G", 0.05, 0.1)),
    fill = "green"
  ) +
  geom_node_text(aes(label = name), repel = TRUE) +
  coord_equal()
```

```{r}
set.seed(-688508752)
email_layout <- create_layout(email_directed_netw, layout = "fr")
email_layout$x[email_layout$name == "Hossam"] <- min(email_layout$x)
email_layout$y[email_layout$name == "Hossam"] <- min(email_layout$y)
ggraph(email_layout) +
  geom_edge_fan(
    aes(
      label = emails,
      end_cap = label_rect(node2.name)
    ),
    arrow = arrow(length = unit(2, "mm")),
    strength = 1.5,
    angle_calc = "along",
    label_dodge = unit(2, "mm")
  ) +
  geom_edge_loop(
    aes(label = emails, end_cap = label_rect(node2.name)),
    arrow = arrow(length = unit(2, "mm")),
    angle_calc = "along",
    label_dodge = unit(2, "mm")
  ) +
  geom_node_label(aes(label = name, colour = location)) +
  scale_colour_brewer(name = "Location", palette = "Dark2") +
  theme(legend.position = c(1, 1), legend.justification = c(1, 1))
```


```{r}
library(tidygraph)
email_directed_netw |>
  activate(nodes) |>
  mutate(
    continent = case_when(
      location == "London" ~ "Europe",
      location == "Los Angeles" ~ "North America",
      location == "Singapore" ~ "Asia"
    )
  ) |>
  activate(edges) |>
  rename(electronic_mail = emails)
```

```{r}
email_directed_netw <-
email_directed_netw |>
activate(edges) |>
mutate(
from_id = .N()$name[from],
to_id = .N()$name[to]
)
email_directed_netw
```
```{r}
email_undirected_netw <-
email_directed_netw |>
convert(to_undirected)
```


```{r}
email_layout
```

```{r}
set.seed(-68858752)
email_layout <- create_layout(email_directed_netw, layout = "fr")
email_layout$x[email_layout$name == "Hossam"] <- min(email_layout$x)
email_layout$y[email_layout$name == "Hossam"] <- min(email_layout$y)
ggraph(email_layout) +
  geom_edge_fan(
    aes(
      label = emails,
      end_cap = label_rect(node2.name)
    ),
  ) +
  geom_edge_loop(
    aes(label = emails, end_cap = label_rect(node2.name)),
    arrow = arrow(length = unit(2, "mm")),
    angle_calc = "along",
    label_dodge = unit(2, "mm")
  ) +
  geom_node_label(aes(label = name, colour = location)) +
  scale_colour_brewer(name = "Location", palette = "Dark2") +
  theme(legend.position = c(1, 1), legend.justification = c(1, 1))
```


```{r}
is_simple(email_undirected_netw)
```

```{r}
email_simple_netw <- 
  email_undirected_netw |>
  convert(to_simple)

email_simple_edges <- 
  email_simple_netw |>
  activate(edges) |>
  as_tibble()

email_simple_edges


```

```{r}
factor_location <-
  email_directed_netw |>
  activate(nodes) |>
  pull(location) |>
  factor()
email_contracted_netw <-
  email_directed_netw |>
  activate(nodes) |>
  mutate(level_as_integer = as.integer(factor_location)) |>
  convert(to_contracted, level_as_integer) |>
  mutate(location = levels(factor_location)) |>
  activate(edges) |>
  mutate(emails = map_dbl(.orig_data, ~ sum(pull(., emails))))
ggraph(email_contracted_netw, layout = "stress") +
  geom_edge_fan(aes(label = emails, end_cap = label_rect(node2.location)),
    arrow = arrow(length = unit(4, "mm")),
    angle_calc = "along",
    label_dodge = unit(2, "mm")
  ) +
  geom_node_label(aes(label = location)) +
  scale_x_continuous(expand = expansion(c(0.1, 0.06)))
```

# Extracting information about nodes 

```{r}
email_directed_netw
```

```{r}
email_directed_netw |>
  activate(edges) |>
  filter(edge_is_to(.N()$name == "Fatimah")) |>
  select(ends_with("_id")) |>
  as_tibble()
```


```{r}
email_directed_netw |>
  activate(edges) |>
  filter(edge_is_incident(.N()$name == "Fatimah")) |>
  select(ends_with("_id"))|>
  as_tibble()
```

```{r}
email_directed_netw |>
  activate(nodes) |>
  filter(node_is_adjacent(name == "Fatimah")) |>
  pull(name)
```

By default, tidygraph considers each node to be adjacent to itself. This is why Fatimah is part of the output above although there is no edge from Fatima to herself. If we want to avoid that, we can use this 

```{r}
email_directed_netw |>
  activate(nodes) |>
  filter(node_is_adjacent(name == "Fatimah", include = FALSE)) |>
  pull(name)
```

```{r}
# those who sent email to Fatimah 
email_directed_netw |>
  activate(nodes) |>
  filter(node_is_adjacent(
    name == "Fatimah", 
    include = FALSE,
    mode = "in"
  )) |>
  pull(name)

# those who received emails from Fatimah 
email_directed_netw |>
  activate(nodes) |>
  filter(node_is_adjacent(
    name == "Fatimah", 
    include = FALSE,
    mode = "out"
  )) |>
  pull(name)
```

# calculate degree and strength of a node 

```{r}
demo_undirected_netw |>
  activate(nodes) |>
  mutate(degree = centrality_degree()) |>
  as_tibble()
```

We can also add argument to centrality_degree, like mode = "in", mode = "out"
, and mode = "all" 

## calculating strength

```{r}
email_directed_netw |>
activate(nodes) |>
mutate(
in_strength = centrality_degree(weights = emails, mode = "in"),
out_strength = centrality_degree(weights = emails, mode = "out"),
strength = centrality_degree(weights = emails, mode = "all")
) |>
as_tibble()
```

# distance 

```{r}
email_directed_netw |>
activate(nodes) |>
mutate(
distance_to_pauline = node_distance_to(name == "Pauline"),
distance_from_pauline = node_distance_from(name == "Pauline")
) |>
as_tibble()
```

```{r}
email_directed_netw |>
  activate(nodes) |>
  mutate(
    neighbors = map(local_members(order = 2), ~ name[.])
  ) |>
  filter(name == 'Lakshmi')|>
  pull(neighbors)
```

the diameter is the maximum of the shortest path lengths between any two nodes 
```{r}
diameter(email_directed_netw)
```

determine the shortest path 
```{r}
pauline_to_simon <-
email_directed_netw |>
convert(to_shortest_path, name == "Pauline", name == "Simon")
```

```{r}
pauline_to_simon |>
activate(nodes) |>
as_tibble()
```
```{r}
pauline_to_simon |>
  activate(edges) |>
  as_tibble()
```

```{r}
email_directed_netw
```

```{r}
shortest_paths(email_directed_netw, "Pauline", "Simon")$vpath
```

```{r}
email_directed_netw |>
convert(
to_shortest_path,
name == "Pauline",
name == "Simon",
weights = emails)
```
ggplot on page 287 


# betweeness 


```{r}
email_directed_netw |>
activate(nodes) |>
mutate(betw = centrality_betweenness()) |>
arrange(desc(betw)) |>
as_tibble()
```


























