---
title: "week09_davis_codes"
author: "shuna"
date: "3/16/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, messag= FALSE)
```

```{r}
library(tidyverse)
```


1. Relation between tibbles, data-frames and lists 
(a) Create a tibble 
```{r}
musicians_tbl <- tibble(
  name = c("Keith", "John", "Paul"),
  band = c("Stones", "Beetles", "Beetles"),
  instrument = c("guitar", "guitar", "bass")
)

musicians_tbl
```

(b) data frame
```{r}
musicians_df <- data.frame(
  name = c("Keith", "John", "Paul"),
  band = c("Stones", "Beetles", "Beetles"),
  instrument = c("guitar", "guitar", "bass")
)
musicians_df
```

(c) list
```{r}
musicians_list <- list(
  name = c("Keith", "John", "Paul"),
  band = c("Stones", "Beetles", "Beetles"),
  instrument = c("guitar", "guitar", "bass")
)
str(musicians_list)
```

(4) 

```{r}
data <- list(musicians_tbl, musicians_df, musicians_list)
```

```{r}
class(musicians_tbl)
class(musicians_df)
class(musicians_list)
```

```{r}
map_at(data, function(x) class(x))
```

```{r}
is_tibble(musicians_tbl)
is_tibble(musicians_df)
is_tibble(musicians_list)
```



```{r}
is.data.frame(musicians_tbl)
is.data.frame(musicians_df)
is.data.frame(musicians_list)
```

```{r}
is.list(musicians_tbl)
is.list(musicians_df)
is.list(musicians_list)
```

tibble and data frame are lists. 


(e) Compare the return values of attribute() when applied to tibble, data frame, and list 

```{r}
attributes(musicians_tbl)
attributes(musicians_df)
attributes(musicians_list)

```

tibble and data frame have three attributes, namely class, row.names and names, but list only has one attribute, name. 

(f) 

```{r}
musicians_df[1:2,]
musicians_tbl[1:2,]
#musicians_list[1:2,]
#musicians_list[,1:2]
```


(g)

```{r}
class(musicians_df) <- NULL
row.names(musicians_df) <- NULL

class(musicians_tbl) <- NULL
row.names(musicians_tbl) <- NULL
```


```{r}
identical(musicians_list, musicians_tbl)
identical(musicians_list, musicians_df)
identical(musicians_df, musicians_tbl)
```

```{r}
nycflights13::flights
```

(a)

(1)
```{r}
map_dbl(mtcars, mean)
```


(ii)

```{r}
map_chr(nycflights13::flights, function(x) typeof(x))
```

(iii)

```{r}
map_lgl(forcats::gss_cat, is.factor)
```

(iv)
```{r}
map_dbl(iris, function(x) length(unique(x)))
map_dbl(iris, n_distinct)
```

(b)

```{r}
pmap(list(10, 1:10, 2:11), runif)
```
`pmap()` iterates over multiple arguments simultaneously. In this case, `pmap()` takes three arguments, n, min, and max, and run `runif()` using these arguments. 

The `runif()` function generates random deviates of the uniform distribution. Using n, min, max, we can generate n number of random samples within the range that is defined by the min and max argument. 

In the actual code, the first 10 specifies the number of random deviates generated from the uniform distribution. The second 1:10 specifies the minimum of the deviates, and 2:11 specifies the maximum of the deviates. So the code generates 10 random samples for each min and max pair (10 times). 


(c)

```{r}
beaver1
```

```{r}
map_dfr(
  list(beaver1, beaver2),
  function(dfr) {
    glm(activ ~ temp, data = dfr, family = binomial) |>
      pluck(coef)
  }
)
```

```{r}
map2_dfr(
  list(beaver1,beaver2),
  ~. glm(.x$activ ~ .x$temp, data = dfr, family = binomial) |> pluck(coef)
)


```


```{r}
starwars
```

(a)
```{r}
colnames(starwars)
```

(b)
```{r}
starwars |>
  filter(map_lgl(starships, ~ "Millennium Falcon" %in% .)) |>
  select(name) |>
  pull()
```


```{r}
starwars |>
  map_chr(films, unique)exp_ma
```

