---
title: "Week07_Davis_Fri_codes"
author: "shuna"
date: "3/4/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, fig.align = "centre")
```

```{r}
library(tidyverse)
library(nycflights13)
library(ggplot2)
```

# 1
## a
```{r}
flights |> 
  filter(dep_delay >= 2) |>
  nrow()
```
## b
```{r}
flights |> 
  filter(origin == "JFK" & dest == c("IAH", "HOU")) |> 
  nrow()
```
## c 
```{r}
flights |>
  filter(dep_time == 2400 | dep_time <= 600) |>
  nrow()
```
# 2

## a
```{r}
q2_a <- flights |>
  filter(is.na(dep_time)) |>
  nrow()

q2_a
```
## b

```{r}
flights |> 
  filter(is.na(dep_time) & is.na(arr_time)) |>
  nrow() |> 
  all.equal(q2_a)
```
## c

cancelled flights 

# 3

```{r}
flights %>%
  group_by(dest) %>%
  summarise(mean_air_time = mean(air_time, na.rm = TRUE)) %>%
  arrange(desc(mean_air_time)) %>%
  head(10) %>%
  ungroup()

# alternatively 
# slice_max(mean_air_time, n = 10)
```

# 4

```{r}
flights %>% 
  mutate(speed = distance / air_time * 60) %>%
  select(dest,air_time, distance, speed) %>%
  arrange(speed) %>%
  head(10)
```

# 5

```{r}
flights %>%
  select(ends_with("arr_time"))
```

# 6
```{r}
q6 <- flights %>%
  select(contains("dep"))
q6
```

# 7

## a
```{r}
q6_a <- q6 %>%
  mutate(diff_time = dep_time - sched_dep_time) %>%
  arrange(diff_time)
q6_a
```

## b
we expected two of them to be the same, which in reality turns out not the case. 

# c

```{r}
calc_time <- function(dep, sch){
  dep_h = dep %/% 100
  sch_h = sch %/% 100
  dep_m = dep %% 100
  sch_m = sch %% 100
  d_time = dep_h * 60 + dep_m
  s_time = sch_h * 60 + sch_m
  if_else(
    sch_h - dep_h > 1,
    d_time + 24*60 - s_time,
    d_time - s_time
  )
}
```


```{r}
q6 %>%
  mutate(diff_time = calc_time(dep_time, sched_dep_time)) %>%
  filter(all.equal(diff_time, dep_delay))
```

```{r}
q7_c <- q6 %>%
  mutate(diff_time = calc_time(dep_time, sched_dep_time),
         same_or_not = diff_time - dep_delay)

unique(q7_c$same_or_not)
```

# 8
Make a scatter plot in which each point represents one day and the coordinates are:
• x: the mean departure delay (conditional on the depaprture delay being known).
• y: the percentage of cancelled flights. We consider a flight as cancelled if the departure time is NA.
Label the axes, give the plot a title and credit the data source. Use code chunk options fig.width,
fig.height and out.width to adjust figure dimensions. All parts of the figure should be clearly legible
without appearing disproportionately large compared to the font size of the running text in the knitted
R Markdown file.
Judging from the plot, is the proportion of cancelled flights related to the mean departure delay?

```{r}
q8 <- flights %>% 
  select(year, month, day, dep_time, dep_delay) %>%
  group_by(year, month, day) %>% 
  mutate(mean_dep_delay = mean(dep_delay, na.rm = TRUE),
         perc = 100 * sum(is.na(dep_time))/n()) %>%
  select(-dep_time, -dep_delay) %>%
  ungroup()

q8 <- distinct(q8)


q8 %>%
  #filter(perc < 40) %>%
  ggplot(aes(mean_dep_delay, perc))+
  geom_point(alpha = 0.3, colour = "blue", na.rm = TRUE) + 
  labs(
    x = "The Mean Departure Delay (mins)",
    y = "Cancelled flights (%)",
    title = "Departure Delay and Cancellation",
    caption = "Source: R package 'nycflights13'"
  ) +
  theme(plot.title = element_text(hjust = 0.5)) + 
  scale_y_continuous(breaks = c(5,10,20,40,60), trans = "sqrt")
```
# 9

Make a plot that shows the mean departure delay by hour. What time of day should you fly if you
want to avoid departure delays?


```{r}
flights %>%
  select(hour, dep_delay) %>%
  filter(dep_delay >= 0) %>%
  group_by(hour) %>%
  summarise(mean_dep_delay = mean(dep_delay, na.rm = TRUE))%>%
  filter(!is.na(mean_dep_delay)) %>%
  ungroup() %>%
  ggplot(aes(hour, mean_dep_delay)) + 
  geom_point(size = 2) + 
  labs(
    x = "Hour of the day",
    y = "Mean Departure Delay (mins)",
    title = "Mean Departure Delay by Hour",
    caption = "Source: R package 'nycflights13'"
  ) + 
  theme(plot.title = element_text(hjust = 0.5)) + 
  scale_x_continuous(breaks = seq(5,24,by = 1),
                     minor_breaks = NULL)
  
```

5am



 

