---
title: "General Social Survey"
author: "Team B"
date: "2/4/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```

```{r, message= FALSE}
library(tidyverse)
library(kableExtra)
library(formattable)
```


### Exercise 1: About the 'Race' of respondents
    
(a)  what are the levels of `gss_cat$race`?

```{r}
levels(gss_cat$race)
```

We see that there are four levels in `gss_cat$race`.

(b) tabulate the frequency of the `races` with `table()`

```{r}
table(gss_cat$race)
```

Additionally, we can see the frequency of each level. "Not applicable" seems to be, well, *not applicable*; that is, none of our data fall in that category. (?) 

(c) Now, we can create our own function `nice_table()` to display a nice table.

```{r}
nice_table <- function(col, name) {
  # col refers to the name of the column in the data set 
  # name refers to the column name to be printed  
  table(gss_cat[, col]) |>
    # convert it to a data frame 
    as.data.frame() |>
    # create a new column for frequency 
    mutate(Freq = color_bar("lightgreen")(Freq)) |>
    kbl(
      # edit a column name 
      col.names = c(name, "Frequency"),
      align = c("l", "r"),
      escape = FALSE
    ) |>
    kable_styling(c("striped", "condensed"), full_width = FALSE)
}

# Check if the function works
nice_table("race", "Race")
```

(d) Create a new column `gss_cat$race_drop_unused` that ahs the same data as `gss_cat$race`, but drop the unused level. 

```{r}
gss_cat$race_drop_unused <- droplevels(gss_cat$race)

# Check if the unused level "Not applicable" has been dropped
!"Not applicable" %in% levels(gss_cat$race_drop_unused)
```

(e) Now, we can form a `nice_table` again!

```{r}
nice_table("race_drop_unused", "Race")
```

(f) Sort the levels in the order of frequency. 

It may be better to organize our data by frequency. The `forcats` package provides a function for this: `fct_infreq()`. This will sort the levels based on the frequency they appear in. Thus, we can form a *nicer* `nice_table`.

```{r}
gss_cat$race_drop_unused <- fct_infreq(gss_cat$race_drop_unused)

# Check if the levels have been ordered
nice_table("race_drop_unused", "Race")
```

### Exercise 2: Investigating the reported income 

(a) What are the levels? 

```{r}
levels(gss_cat$rincome)
```

(b) tabulate the frequency of the income levels with the `nice_table()`.

```{r}
nice_table("rincome", "Income")
```

(c) 

From the above table output, we realise that there is an inconsistent use of "-", and "to". Further, the levels are of unequal size. Finally, it may be a good idea to add commas to the numbers.

```{r}
levels(gss_cat$rincome_reformatted)

gss_cat$rincome_reformatted <- gss_cat$rincome |>
  # merge different factors 
  fct_collapse(
    "$5,000 - 9,999" = c(
      "$8000 to 9999",
      "$7000 to 7999",
      "$6000 to 6999",
      "$5000 to 5999"
    ),
    "$4,999 or less" = c(
      "$4000 to 4999",
      "$3000 to 3999",
      "$1000 to 2999",
      "Lt $1000"
    )
  ) |>
  # reorder the factor levels 
  fct_relevel(c(
    "$25000 or more",
    "$20000 - 24999",
    "$15000 - 19999",
    "$10000 - 14999",
    "$5000 - 9999",
    "$4999 or less",
    "No answer",
    "Don't know",
    "Refused",
    "Not applicable"
  ))

# plot a table 
nice_table("rincome_reformatted", "Income")
```

```{r}
gss_cat[["rincome"]] <- fct_recode(gss_cat[["rincome"]],
                      "Not applicable" = "No answer",
                      "Not applicable"= "Don't know",
                      "Not applicable" = "Refused",
                      "$25,000 or more"= "$25000 or more",
                      "$20,000 to 24,999" = "$20000 - 24999",
                      "$15,000 to 19,999" = "$15000 - 19999",
                      "$10,000 to 14,999" = "$10000 - 14999",
                      "$5,000 to 9,999" = "$8000 to 9999", 
                      "$5,000 to 9,999" = "$7000 to 7999", 
                      "$5,000 to 9,999" = "$6000 to 6999",
                      "$5,000 to 9,999" = "$5000 to 5999",
                      "$1,000 to 4,999" = "$4000 to 4999",
                      "$1,000 to 4,999" = "$3000 to 3999",
                      "$1,000 to 4,999" = "$1000 to 2999",
                      "Less than $1,000" = "Lt $1000")

gss_cat[["rincome"]] <- fct_shift(fct_drop(gss_cat[["rincome"]]), n = 1)

nice_table("rincome", "Income")
```


### Exercise 3: About the 'Political Association' of respondents

(a)

```{r}
levels(gss_cat$partyid)
```

(b)

```{r}
gss_cat$partyid_aggregated <- fct_collapse(gss_cat$partyid,
  "Republican" = c(
    "Strong republican",
    "Not str republican"
  ),
  "Independent" = c(
    "Ind,near rep",
    "Independent",
    "Ind,near dem"
  ),
  "Democrat" = c(
    "Not str democrat",
    "Strong democrat"
  ),
  "Other" = c("Other party"),
  "Missing" = c(
    "Don't know",
    "No answer"
  )
) |>
  fct_infreq()
```

(c)

```{r}
nice_table("partyid_aggregated", "Party affiliations")
```

### Exercise 4: About the 'Religious Association' of respondents

(a)

```{r}
gss_cat$is_protestant <- fct_collapse(gss_cat$relig,
  "Protestant" = c("Protestant"),
  other_level = "Other"
)
```

(b)


```{r}
gss_cat$has_denom <- fct_collapse(gss_cat$denom,
  "No answer" = c("No answer"),
  "No denomination" = c("No denomination"),
  "Not applicable" = c("Not applicable"),
  "Don't know" = c("Don't know"),
  other_level = "Has a denomination"
) |> fct_relevel(c(
  "Has a denomination", "No denomination", "Not applicable",
  "Don't know", "No answer"
))
```

(c)


```{r}
data <- gss_cat[, c("is_protestant", "has_denom")]
data <- as.data.frame.matrix(table(data$has_denom, data$is_protestant))


data |>
  mutate(
    Protestant = color_bar("lightgreen")(Protestant),
    Other = color_bar("lightgreen")(Other)
  ) |>
  kbl(
    col.names = c("Protestant", "Other"),
    align = c("r"),
    escape = FALSE
  ) |>
  kable_styling(
    c("striped", "condensed"),
    full_width = FALSE
  )
```

```{r}
```{r}
gss_cat[["rincome"]] <- fct_recode(gss_cat[["rincome"]],

  # merging similar items
  "Not applicable" = "No answer",
  "Not applicable" = "Don't know",
  "Not applicable" = "Refused",

  # correcting formatting
  "$25,000 or more" = "$25000 or more",
  "$20,000 to 24,999" = "$20000 - 24999",
  "$15,000 to 19,999" = "$15000 - 19999",
  "$10,000 to 14,999" = "$10000 - 14999",
  "$5,000 to 9,999" = "$8000 to 9999",
  "$5,000 to 9,999" = "$7000 to 7999",
  "$5,000 to 9,999" = "$6000 to 6999",
  "$5,000 to 9,999" = "$5000 to 5999",
  "$1,000 to 4,999" = "$4000 to 4999",
  "$1,000 to 4,999" = "$3000 to 3999",
  "$1,000 to 4,999" = "$1000 to 2999",
  "Less than $1,000" = "Lt $1000"
)

# reorder the rows so that Not applicable is at the bottom 
gss_cat[["rincome"]] <- fct_shift(fct_drop(gss_cat[["rincome"]]), n = 1)

# plot a table
nice_table("rincome", "Income")
```
```

