---
title: "General Social Survey"
author: "Team B"
date: "2/4/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```

```{r}
library(tidyverse)
library(kableExtra)
library(formattable)
```


### Exercise 1: About the 'Race' of respondents

#### a. What are the levels of `gss_cat$race`?

```{r}
levels(gss_cat$race)
```

We see that there are four levels in `gss_cat$race`.

#### b. tabulate the frequency of the `races` with `table()`

```{r}
table(gss_cat$race)
```

Additionally, we can see the frequency of each level. "Not applicable" seems to be, well, *not applicable*; that is, none of our data fall in that category. 

#### c. Now, we can create our own function `nice_table()` to display a nice table.

```{r}
nice_table <- function(col, name) {
  # col refers to the name of the column in the data set
  # name refers to the column name to be printed
  table(gss_cat[, col]) |>
    # convert it to a data frame
    as.data.frame() |>
    # create a new column for frequency
    mutate(Freq = color_bar("lightgreen")(Freq)) |>
    kbl(
      # edit a column name
      col.names = c(name, "Frequency"),
      align = c("l", "r"),
      escape = FALSE
    ) |>
    kable_styling(c("striped", "condensed"), full_width = FALSE)
}

# Check if the function works
nice_table("race", "Race")
```

#### d. Create a new column `gss_cat$race_drop_unused` that has the same data as `gss_cat$race`, but drop the unused level.

```{r}
# Dropping unused levels
gss_cat$race_drop_unused <- droplevels(gss_cat$race)

# Check if the unused level "Not applicable" has been dropped
!"Not applicable" %in% levels(gss_cat$race_drop_unused)
```

#### e. Now, we can form a `nice_table()` again!

```{r}
nice_table("race_drop_unused", "Race")
```

#### f. Sort the levels in the order of frequency.

It may be better to organize our data by frequency. The `forcats` package provides a function for this: `fct_infreq()`. This will sort the levels based on the frequency they appear in. Thus, we can form a *nicer* `nice_table`.

```{r}
# Sorting based on frequency
gss_cat$race_drop_unused <- fct_infreq(gss_cat$race_drop_unused)

# Check if the levels have been ordered
nice_table("race_drop_unused", "Race")
```

### Exercise 2: Investigating the reported income

#### a. What are the levels?

```{r}
levels(gss_cat$rincome)
```

#### b. Tabulate the frequency of the income levels with the `nice_table()`.

```{r}
nice_table("rincome", "Income")
```

#### c. Improve the table. 

From the above table output, we realise that there is an inconsistent use of "-", and "to". Further, it may be a good idea to add commas to the numbers. We should rename `Lt $1000` to the same format as `$25000 or more`. Finally, "No answer", "Don't know", "Refused" and "Not applicable" seem to be semantically the same. We should also include these at the end rather than the top. Thus, we will merge them into "Not applicable". This also has the benefit of shifting all of them to the end (because "Not applicable" is already at the end).


```{r}
gss_cat[["rincome"]] <- fct_recode(gss_cat[["rincome"]],

  # merging similar items
  "Not applicable" = "No answer",
  "Not applicable" = "Don't know",
  "Not applicable" = "Refused",

  # correcting formatting
  "$25,000 or more" = "$25000 or more",
  "$20,000 to 24,999" = "$20000 - 24999",
  "$15,000 to 19,999" = "$15000 - 19999",
  "$10,000 to 14,999" = "$10000 - 14999",
  "$5,000 to 9,999" = "$8000 to 9999",
  "$5,000 to 9,999" = "$7000 to 7999",
  "$5,000 to 9,999" = "$6000 to 6999",
  "$5,000 to 9,999" = "$5000 to 5999",
  "$1,000 to 4,999" = "$4000 to 4999",
  "$1,000 to 4,999" = "$3000 to 3999",
  "$1,000 to 4,999" = "$1000 to 2999",
  "Less than $1,000" = "Lt $1000"
)

# reorder the rows so that Not applicable is at the bottom 
gss_cat[["rincome"]] <- fct_shift(fct_drop(gss_cat[["rincome"]]), n = 1)

# plot a table
nice_table("rincome", "Income")
```


### Exercise 3: About the 'Political Association' of respondents

#### a. What are the levels of `gss_cat$partyid`?

```{r}
levels(gss_cat$partyid)
```

#### b. Add a factor column `gss_cat$partyid_aggregated`.

```{r}

# combining appropriate parties.
gss_cat$partyid_aggregated <- fct_collapse(gss_cat$partyid,
  "Republican" = c(
    "Strong republican",
    "Not str republican"
  ),
  "Independent" = c(
    "Ind,near rep",
    "Independent",
    "Ind,near dem"
  ),
  "Democrat" = c(
    "Not str democrat",
    "Strong democrat"
  ),
  "Other" = c(
    "Other party"
  ),
  "Missing" = c(
    "Don't know",
    "No answer"
  )
) |>
  fct_infreq()
```

#### c. Make a frequency table of `gss_cat$partyid_aggregated` with the `nice_table()` function from (1)-c.

```{r}
nice_table("partyid_aggregated", "Party affiliations")
```

### Exercise 4: About the 'Religious Association' of respondents

#### a. Add a factor column `gss_cat$is_protestant` with levels "Protestant" and "Other".

```{r}
gss_cat$is_protestant <- fct_collapse(gss_cat$relig,
  "Protestant" = c(
    "Protestant"
  ),
  other_level = c(
    "Other"
  )
)
```

#### b. Add a factor column `gss_cat$has_denom` with levels.

```{r}
gss_cat$has_denom <- fct_collapse(gss_cat$denom,
  "No answer" = c(
    "No answer"
  ),
  "No denomination" = c(
    "No denomination"
  ),
  "Not applicable" = c(
    "Not applicable"
  ),
  "Don't know" = c(
    "Don't know"
  ),
  other_level = c(
    "Has a denomination"
  )
) |> fct_relevel(c(
  "Has a denomination",
  "No denomination",
  "Not applicable",
  "Don't know",
  "No answer"
))
```

#### c. Make a two-way table as indicated in figure 2.

```{r}
# extract the necessary factors 
data <- gss_cat[, c("is_protestant", 
                    "has_denom")]
# convert it to a data frame 
data <- as.data.frame.matrix(table(data$has_denom, data$is_protestant))


data |>
  mutate(
    # add bars
    Protestant = color_bar("lightgreen")(Protestant),
    Other = color_bar("lightgreen")(Other)
  ) |>
  # name the columns and edit the style 
  kbl(
    col.names = c("Protestant", 
                  "Other"),
    align = c("r"),
    escape = FALSE
  ) |>
  kable_styling(
    c("striped", 
      "condensed"),
    full_width = FALSE
  )
```
