---
title: "DAVisR"
author: "shuna"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
---

# Conditional element selection 

```{r}
v <- c(31, -50, 93, 29, -44, 93)
```

## Comparison Operators 


1. dplyr::between(x,left,right) TRUE iff left <= x <= right 
2. x %in% y TRUE iff x is an element in y 

## if_else()

we want to turn `v <- c(31, -50, 93, 29, -44, 93)` into a chanracter string
`("odd', "even", "odd", "odd", "even", "odd")`. 

We can turn `v` into a logical vector 
```{r}
v %% 2 == 0
```
we use a function `if_else()`

`if_else(condition, true, false)`

true and false kust be vectors of the same class 
- true[i] if condition[i] is TRUE
- false[i] if condition[i] is FALSE

```{r, message=FALSE}
library(dplyr)
if_else(c(TRUE, TRUE, FALSE, FALSE), 1:4, -1:-4)
```
```{r}
if_else(v %% 2 == 0, "even", "odd")
```

## case_when()

```{r}
x <- c(31, -50, 93, 29, -44, 93, 91.5)
case_when(
  x %% 2 == 0 ~ "even",
  x %% 2 == 1 ~ "odd",
  TRUE ~ "neither even nor odd"
)
```
## Logical operators 

bit演算ですね。

```{r}
pin <- c(TRUE, FALSE, FALSE, TRUE)
password <- c(FALSE, FALSE, TRUE, TRUE)
```

### AND

the operator AND returns TRUE iff both input are TRUE. 
```{r}
pin & password
```
### OR

the operator OR returns TRUE when either is TRUE or both are TRUE
```{r}
pin | password
```
### NOT

```{r}
!pin
```

# Data type conversion 

## Mixing numeric and character elements

When mixing numeric vector and character vector, you get character vector. 


```{r}
"42" > 8
```

```{r}
"42" < 8
```
The letter `4` comes alphabetically before the letter `8`.


## General rules of coercion 

```{r}
# logical vector 
logical <- c(TRUE, FALSE, TRUE)

# numeric vector 
x 

# int 
int <- 1:3

# character 

chr <- c("odd", "even", "neither")
```

```{r}
c(logical, int)
```
TRUE and FALSE are turned into the integers 1 and 0. 

## as. 

```{r}
as.numeric(c(TRUE, FALSE))
```

```{r}
as.logical(seq(-3,6, by = 0.5))
```
FALSE if the number is exactly 0 

FALSE if the element is either FALSE, false, F 

```{r}
as.logical(c(TRUE, "true", T, FALSE, "false", F, NA))
```
```{r}
# if negative, round up 
# if positive, round down
as.integer(c(-2.111, -2.99999, -1.2, 0, 1.2, 2.99))
```
## Taking advantage 

```{r}
nchar(c("TRUE", "Odd"))
```
```{r}
nchar(c(42, -6857))
```

```{r}
logical_value <- rep(c(TRUE, FALSE), c(500, 100))
sum(logical_value)
```
# Missing and Undefined values 

```{r}
x <- cat()
x
```
## NA
```{r}
class(NA_real_)
```

```{r}
class(NA_integer_)
```

```{r}
nchar(NA)
```

```{r}
test <- c(1,2,3,4,5,NA, 6,7,8,9,10)
```

```{r}
is.na(test)
```
```{r}
any(is.na(test))
```
```{r}
anyNA(test)
```

```{r}
mean(test[!is.na(test)])
```
## INF

```{r}
Inf + 1

```
```{r}
1/Inf
```

## NaN

```{r}
0 / 0
```
## NULL

```{r}
v_1 <- c(1, Inf, NA, NULL, NaN)
v_1
```
# Factors 

## builtin factors 
```{r}
state.name
```

```{r}
state.region
```

```{r}
class(state.name)
class(state.region)
```
## Create factors 

```{r}
v <- c("apple", "orange", "banana")
factor(v)
```
we can add levels 

```{r}
v2 <- factor(v, levels = c("apple", "banana", "orange", "grape"))
v2
```
factors not in the levels will be NAs 
```{r}
v3 <- factor(c("apple", "banana", "orange", "peach"),
             levels = c("apple", "banana", "orange", "grape"))
v3
```

## Subsetting of factors 

```{r}
v2[1:2]
```

```{r}
v3[-4]
```

```{r}
v[c(TRUE, TRUE, FALSE)]

```
v1[c(TRUE, TRUE, FALSE)]

```{r}
v == "orange"
v[v == "orange"]
```
## Retrieving and changing the levels 

```{r}
levels(v2)
```

```{r}
nlevels(v2)
```

```{r}
library(forcats)
v4 <- fct_recode(v2,
                 A = "apple", 
                 B = "banana",
                 O = "orange",
                 G = "grape")
v4
```

```{r}
as.character(v4)
```

## R's internal representation of factors 

```{r}
f1 <- factor(c("Grande", "Venti", "Grande", "Tall"), 
             levels = c("Grande", "Venti", "Tall", "Short"))
str(f1)
```
```{r}
typeof(f1)
f1
```

```{r}
attributes(f1)
```

```{r}
as.integer(f1)
```
```{r}
f2 <- fct_recode(
  f1, 
  '16' = "Grande", 
  "17" = "Short", 
  "18" = "Venti",
  "19" = "Tall")
f2
```
```{r}
as.integer(f2)
```

```{r}
as.integer(levels(f2)[f2])

```
## Dropping unused levels 
```{r}
unique(f1)
```
```{r}
f_drop <- fct_drop(f1)
f_drop
```

## Arrangng the order of levels according to their frequency 

```{r}
favorite <- factor(c(
  rep("coffee", 61),
  rep("tea", 13),
  rep("merchandise", 1),
  rep("food", 25)
))
```

```{r}
levels(favorite)
```

```{r}
table(favorite)
```

```{r}
favorite_sorted <- fct_infreq(favorite)
table(favorite_sorted)
```

## Ordered factors 

```{r}
f1 <- factor(c("Grande", "Venti", "Grande", "Tall"), 
             levels = c("Grande", "Venti", "Tall", "Short"), 
             ordered = TRUE)
f1
```
```{r}
f1 <- ordered(c("Grande", "Venti", "Grande", "Tall"), 
             levels = c("Grande", "Venti", "Tall", "Short"))
f1
```

```{r}
f1 < "Tall"
```

```{r}
opinion <- factor(c(
  "Disagree",
  "Agree",
  "Disagree",
  "Neutral",
  "Agree",
  "Agree"
))
str(opinion)
```
```{r}
size <- c("Grande", "Venti", "Grande", "Tall")
f <- ordered(size, levels = c("Short", "Tall", "Grande", "Venti"))
f > "Tall"
```

```{r}
library(ggplot2)
#ggplot(mpg, aes(displ, hwy)) +
  geom_point(colour = "blue")
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(colour = "blue"))
ggplot(mpg, aes(displ, hwy, colour = "blue")) +
  geom_point()
```
```{r}
data(harelynx)
ggplot(harelynx, aes(hare, lynx, colour = year)) +
  geom_path() +
  geom_point()
```


```{r}
writers_c <- list(
  c("lemmon", "Mccartney"),
  c("Harrison", "lennon", "maccartney", "starkey")
)
side_c <- list(
  1:2,
  c("the fool on the hill", "flying"),
  writers_c,
  c(180,136)
)
side_c
```

```{r}
side_c[c(1,3)]
```
```{r}
names(side_c) <- c("track", "title", "writers", "duration")
side_c
```

```{r}
library(purrr)
side_c$title
side_c |> 
  pluck("title") 
```

```{r}
library(tibble)
side_c_tibble <- 
  tibble(
    track = 1:2,
    title = c("The fool on the hill", "Flying"),
    writers = writers_c,
    duration = c(180,136)
  )
side_c_tibble
```

```{r}
side_c_tibble$writers[[1]][2]
```

```{r}
titles <- list(
  c("magical", "your mother"),
    "I am Ali", 
  c("the fool", "is flying")
)

map(titles, length)
map_dbl(titles, length)
```

```{r}
durations <- list(
  c(168, 153),
  275,
  c(180,136),
  230
)
str(durations)
map_dbl(durations, sum)
```

