---
title: "Solutions: Choropleth maps of country-level statistics"
subtitle: "YSC2210 - DAVis with R"
author: "Michael T. Gastner"
output:
  pdf_document:
    citation_package: natbib
urlcolor: blue
header-includes: 
    - \usepackage{hyperref}
bibliography: preston.bib
biblio-style: natbib
---

```{r setup, echo=FALSE}
knitr::opts_chunk$set(message = FALSE)
```

I need the following packages for this exercise.

```{r}
library(countrycode)
library(readxl)
library(sf)
library(tidyverse)
library(tmap)
```

(1)
    Code from the exercise about relational data.

    ```{r}
    gdp_per_cap <- read_excel(
      "data/API_NY.GDP.PCAP.KD_DS2_en_excel_v2_3731742.xls",
      skip = 3
    )
    life_exp <- read_excel(
      "data/API_SP.DYN.LE00.IN_DS2_en_excel_v2_3731513.xls",
      skip = 3
    )

    # Next function can pivot any of the World Bank tibbles
    pivot_longer_wb <- function(tb, name) {
      tb |>
        select(-starts_with("Indicator")) |>
        pivot_longer(
          matches("^\\d{4}$"), # String with exactly 4 digits
          names_to = "year",
          values_to = {{ name }},
          names_transform = list(year = as.integer)
        )
    }
    wb <-
      Hmisc::llist(life_exp, gdp_per_cap) |>
      imap(pivot_longer_wb) |>
      reduce(left_join, by = c("Country Name", "Country Code", "year")) |>
      semi_join(codelist, by = c("Country Code" = "iso3c"))
    ```

(2)

    ```{r}
    import_and_simplify <- function(zip_file) {
      zip_file |>
        unzip() |>
        str_subset(".shp$") |>
        read_sf() |>
        rmapshaper::ms_simplify()
    }
    country_borders <-
      import_and_simplify("data/wb_countries_admin0_10m.zip")
    ```
    
(3)

    ```{r}
    country_borders <-
      country_borders |>
      mutate(ISO_A3 = case_when(
        NAME_EN == "Norway" ~ "NOR",
        NAME_EN == "France" ~ "FRA",
        TRUE ~ ISO_A3
      ))
    ```
    
(4)

    ```{r}
    countries <-
      country_borders |>
      inner_join(wb, by = c("ISO_A3" = "Country Code")) |>
      filter(year == 2015) |>
      select(
        name = "Country Name",
        code = "ISO_A3",
        gdp_per_cap,
        life_exp,
        geometry
      ) |>
      print()
    ```

    We check whether there are any empty geometries.
    
    ```{r}
    any(st_is_empty(countries))
    ```
    
    In my case, there are no empty geometries.
    You may find empty geometries if you pursue another strategy to merge
    `countries` with `country_borders`.
    
(5)

    ```{r}
    tm_shape(countries) +
      tm_polygons()
    ```
    
    Western Sahara (claimed by Morocco), Aksai Chin, Arunachal Pradesh
    (both disputed between India and China) and Antarctica are not on this
    map.
    
(6)

    ```{r}
    land <- import_and_simplify("data/wb_land_10m.zip")
    tm_shape(land) +
      tm_polygons() +
      tm_shape(countries) +
      tm_polygons()
    ```

(7)
    The projection of the map above is the projection of the `sf` object
    `land` because `land` is the argument when `tm_shape()` is called for the
    first time.
    
    ```{r}
    st_crs(land)$proj4string
    ```
    
    The coordinate reference system is a plain longitude-latitude grid, which
    is not an equal-area projection.
    
(8) The following URL contains information about generating a logarithmic
    scale:
    <https://github.com/mtennekes/tmap/issues/128>.
    However, for me the suggested solution (`style = "log10_pretty"`) results
    in too few bins.
    Therefore, I set the breaks manually in the next code chunk.

    ```{r fig.width=10, fig.height=8, fig.align='centre'}
    choropleth <- tm_shape(land, projection = "+proj=eck4") +
      tm_polygons(col = "grey") +
      tm_shape(countries) +
      tm_polygons(
        c("gdp_per_cap", "life_exp"),
        border.col = "grey30",
        palette = list("Oranges", "Greens"),
        breaks = list(
          c(-Inf, 1000, 2000, 5000, 10000, 20000, 50000, Inf),
          c(-Inf, 60, 65, 70, 75, 80, 85, Inf)
        ),
        colorNA = "grey",
        title = c(
          "GDP per capita\n(US$, PPP 2015)",
          "Life expectancy at birth\n(years)"
        ),
        lwd = 0.5
      ) +
      tm_text("code", size = "AREA", col = "black") +
      tm_credits(
        c("", "Source: World Bank"),
        position = c("left", "bottom")
      ) +
      tm_layout(
        bg.color = "lightblue",
        frame = FALSE,
        inner.margins = c(0.00, 0.15, 0.01, 0.01),
        earth.boundary = TRUE,
        space.color = "white",
        main.title = "Wealth and Health by Country in 2015"
      )
    choropleth
    ```

    Run the next code chunk to export the map to a file.
    
    ```{r}
    tmap_save(
      choropleth,
      "images/choropleth_map.pdf",
      width = 10,
      height = 8
    )

    # Crop white space around the map
    knitr::plot_crop("images/choropleth_map.pdf")
    ```
    
(9)
    By comparing the two maps, we notice an overall tendency that per-capita
    GDP and average life expectancy are correlated.
    Most countries that are poor and have a short average life expectancy are
    in central Africa.
    Afghanistan is the only comparable country outside Africa, with similarly
    low per-capita GDP and short life expectancy.
    Western Europe is home to many of the wealthiest and longest-living
    nations.
    The group of top performing countries also includes the USA, Canada,
    Australia and New Zealand.
    Most Latin American countries are in the midrange for both metrics.
    Asia is very heterogeneous: Afghanistan is at the low end of the spectrum,
    whereas Japanese and South Koreans enjoy very wealthy and long lives.

(10)
    Geographic maps show clearly where countries are located, whereas the
    colour code in figure \ref{fig:gapminder} can only imply coarse-grained
    locations.
    For example, green bubbles represent all countries in Asia, which spans a
    huge area.
    However, the scatter plot contains information about population
    size (represented by the areas of the bubble) that is not present on
    our choropleth maps.
    We could have overlaid proportional symbol maps on top of the choropleth
    maps, using bubbles with an area representing population size centred on
    the countries' centroids.
    However, the bubbles would have obstructed the country labels and many of
    the boundaries.
    A more elegant solution than a proportional symbol map would be a
    cartogram in which each country appears with an area in proportion to its
    population.
    R currently does not have a mature package for cartogram construction.

    Another disadvantage of the faceted maps is that viewers must shift their
    gaze between two spatially separated facets to recognise that there is a
    correlation between per-capita GDP and life expectancy.
    An interactive map in which viewers can toggle between the two variables
    would alleviate the problem, but that technique does not work on printed
    maps.
    The correlation is more directly obvious in the scatter plot because the
    two variables are on the coordinate axes.
    
```{r gapminder, echo=FALSE, fig.cap='\\label{fig:gapminder}Scatter plot of life expectancy vs.\ GDP per capita by country.'}
knitr::include_graphics("images/gg_preston_curve.pdf")
```