---
title: "final_practice"
author: "shuna"
date: "5/4/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(lubridate)
library(tidyverse)
library(tmap)
library(ggraph)
library(igraph)
library(tidygraph)
library(countrycode)
library(readxl)
library(sf)
library(tmap)
library(rmapshaper)
```
(a) Create a tbl_graph from these data. Note that the network is undirected.

```{r}
f_nodes <- read_csv("florence_nodes.csv")
f_edges <- read_csv("florence_edges.csv") 
```

```{r}
# tidygraph 
florence <- tbl_graph(
  nodes = f_nodes,
  edges= f_edges,
  directed = FALSE
)
florence
```

(b) How many edges are in the network? 
```{r}
#nodes
gorder(florence)
#edges 
gsize(florence)
```

(b_2) are there anuy parallel edges or loops in the network? 
```{r}
is_simple(florence)
```

(c)what is the density of the network? 

The density of a network is the proportion of edges in a network that actually do exist out of all edges that could potentially exist if all node pairs were connected.

```{r}
edge_density(florence)
```

(d) what is the transitivity of the network? 

transitivity is a tendency for triplets to form triangles in undirected networks.

```{r}
transitivity(florence)
```

For each node in the network we can calculate how many triangles are involved. 
```{r}
florence |>
  activate(nodes) |>
  mutate(local_trans = local_triangles()) |>
  as_tibble()
```

We can also compare nodes in the same network in terms of their tendency to form local triangles 
```{r}
florence |>
  activate(nodes) |>
  mutate(cluster_coef = local_transitivity()) |>
  as_tibble()
# number of undirected triangles that contain x 
# divided by number of undirected troplets with x as midpoint 
```

(e) what is the diameter of the network? 

the diameter of a network is the longest finite distance between any pair of nodes. i.e., the max of the shortest path length between any two nodes. 
```{r}
diameter(florence)
```

```{r}
florence
```

(f) what is the shortest path from the Guadagni family to the Barbadori family? 

to and from the node...
```{r}
florence |>
  activate(nodes) |>
  mutate(distance_to_ginori = node_distance_to(name == "Ginori"),
         distance_from_ginari = node_distance_from(name == "Ginori")) |>
  as_tibble()
```

shortest path between two nodes 
```{r}
florence |>
  convert(to_shortest_path, name == "Guadagni", name == "Barbadori") |> as_tibble()
```
```{r}
florence |>
  convert(
    to_shortest_path, 
    name == "Guadagni", 
    name == "Barbadori"
  ) |> activate(edges) |>
  as_tibble()
  
```
```{r}
all_shortest_paths(florence, from = "Guadagni", to = "Barbadori")
```


# highest degree 
```{r}
florence |>
  activate(nodes) |>
  mutate(
    degree = centrality_degree(mode = "all")
  ) |>
  as_tibble()
```


## highest betweenness centrality 
```{r}
florence |>
  activate(nodes) |>
  mutate(
    betw = centrality_betweenness()
  ) |>
  as_tibble() |>
  arrange(betw) |>
  tail(1) |>
  pull()
```

```{r}
florence |>
  activate(nodes) |>
  mutate(
    eigen = centrality_eigen()
  ) |>
  as_tibble() |>
  arrange(eigen) |>
  tail(1) |>
  pull()
```

## components 

```{r}
florence |>
  activate(nodes) |>
  mutate(comp = group_components()) |>
  as_tibble() |>
  select(comp) |>
  distinct()
```

## max cliques 
```{r}
largest.cliques(florence)
```

```{r}
ggraph(florence, layout = "stress") +
  geom_edge_link() + 
  geom_node_label(aes(label = name))
```


# Data wrangling 


```{r}
epl <- read_csv("epl_results.csv", na = "-")
head(epl)
```
```{r}
abbr <- colnames(epl)[2:21]
teams <- epl |>
  cbind(abbr) |>
  select(name = home, abbr) |>
  print()
```



```{r}
epl_pivotted <- epl |> 
  pivot_longer(!home, names_to = "abbr", values_to = "result") |>
  print()
```

```{r}
new_teams = teams
for (i in 1:19){
  new_teams <- rbind(new_teams, teams)
}
new_teams
```



```{r}
epl_replaced <- left_join(epl_pivotted, teams, by = "abbr") |>
  filter(!is.na(result)) |>
  select(home, away = name, result)
```

```{r}
library(tidyr)
epl_sep <- separate(
  data = epl_replaced, 
  col = result, 
  into = c("home_goals", "away_goals"), 
  sep = "\\-"
)
```

```{r}
epl_pts <- epl_sep |>
  mutate(
    home_pts = case_when(
      home_goals > away_goals ~ 3,
      home_goals < away_goals ~ 0,
      TRUE ~ 1
    ),
    away_pts = case_when(
      home_goals < away_goals ~ 3,
      home_goals > away_goals ~ 0,
      TRUE ~ 1
    )
  )
```

```{r}
epl_pi <- epl_pts |>
  pivot_longer(
    cols = c(home, away),
    names_to = "home_or_away",
    values_to = "team"
  ) |>
  select(team, home_or_away, home_goals, away_goals, home_pts, away_pts) |>
  print()
```

```{r}
epl_i <- epl_pi |>
  mutate(pts = if_else(
    home_or_away == "home", 
    home_pts,
    away_pts
  ),
  goals_for = if_else(
    home_or_away == "home",
    home_goals,
    away_goals
  ),
  goals_against = if_else(
    home_or_away == "away",
    home_goals,
    away_goals
  )) |>
  select(-home_pts, -away_pts, -home_goals, -away_goals, -home_or_away) |>
  print()
```


```{r}
epl_j <- epl_i |>
  mutate(
    goals_for = as.integer(goals_for),
    goals_against = as.integer(goals_against)
  ) |>
  group_by(team) |>
  summarise(
    pts = sum(pts),
    goals_for = sum(goals_for),
    goals_against = sum(goals_against)
  ) |>
  mutate(
    goals_diff = goals_for - goals_against
  ) |>
  select(-goals_against) |>
  ungroup() |>
  print()
```

```{r}

```

```{r}
epl_k <- epl_j |>
  arrange(-pts, -goals_diff) |>
  rowid_to_column() |>
  dplyr::rename(pos = rowid) |>
  print()
```

# string 

```{r}
ipsum <- read_file("corporate_ipsum.txt")
ipsum <- str_split(ipsum, "\\s+") |> unlist()
str(ipsum)
```


```{r}
str_match("The chromosome had duplicated duplicated hard themselves", 
          "(\\b\\w+\\b) \\1")
```

```{r}
ipsum <- str_remove_all(ipsum, "[[:punct:]]+$")
ipsum <- str_remove_all(ipsum, "^[[:punct:]]+")
```

```{r}
ipsum <- str_to_lower(ipsum)
```

```{r}
ipsum <- ipsum[!(ipsum %in% tm::stopwords())]
```

```{r}
ipsum <- ipsum[ipsum != "."]
```



```{r}
f <- tibble(
  word = as.vector(ipsum)
)
```
f
```{r}
f
```

```{r}
st <- f |>
  group_by(word) |>
  summarise(count = n()) |>
  filter(count >= 20) |>
  print()
```


```{r}
library(ggwordcloud)
st |>
  ggplot(aes(label = word, size = count)) +
  geom_text_wordcloud() +
  scale_size_area(max_size = 8) + 
  
  theme_minimal()
```

# choropleth 

```{r}
atm <- read_excel("API_FB.ATM.TOTL.P5_DS2_en_excel_v2_3932325.xls", skip = 2)
```



```{r}
atm <- atm |>
  select("Country Name","Country Code", "2018") |>
  print()
```

```{r}
sum(is.na(atm$`2018`))
```

```{r}
import_and_simplify <- function(zip_file) {
  zip_file |>
    unzip() |>
    str_subset(".shp$") |>
    read_sf() |>
    rmapshaper::ms_simplify()
  }

wb_c <- import_and_simplify("wb_countries_admin0_10m.zip")
wb_l <- import_and_simplify("wb_land_10m.zip")
```

```{r}
View(atm)
```

```{r}
wb_c <- wb_c |>
  mutate(
    ISO_A3 = 
      case_when(
        FORMAL_EN == "French Republic" ~"FRA",
        FORMAL_EN == "Kingdom of Norway" ~"NOR",
        TRUE ~ISO_A3
      )
  )
```


```{r}
colnames(atm)
```

```{r}
countries <- wb_c|>
  inner_join(atm, by = c("ISO_A3" = "Country Code")) |>
  select(
    code = "ISO_A3",
    atm = "2018",
    geometry
  )
```


```{r}
countries
```

```{r}
tm_shape(wb_l, projection = "+proj=eck4") +
  tm_polygons("grey") +
  tm_shape(countries, projection = "+proj=eck4") +
  tm_polygons(
    "atm",
    pallete = "OrRd",
    colorNA = "grey")
```

# lubridate 

```{r}
covid19 <- read_csv("covid_19_data.csv") 
head(covid19)
```

```{r}
covid19 <- covid19 |>
  filter(`Country/Region` == "Mainland China" | `Country/Region` == "Singapore")
```

```{r}
covid <- covid19 |>
  mutate(ObservationDate = mdy(ObservationDate)) |>
  print()
```

```{r}
covid |>
  group_by(ObservationDate, `Country/Region`) |>
  summarise(count = sum(Confirmed)) |>
  ggplot(aes(ObservationDate, count, color = `Country/Region`)) +
  geom_line()
```

```{r}
?slice_head
```

```{r}
MC <- covid |>
  filter(`Country/Region` == "Mainland China") |>
  group_by(ObservationDate, `Country/Region`) |>
  summarise(count = sum(Confirmed)) |>
  filter(count > 10000) |>
  head(1) |>
  pull(ObservationDate)
```

```{r}
MS <- covid |>
  filter(`Country/Region` == "Singapore") |>
  group_by(ObservationDate, `Country/Region`) |>
  summarise(count = sum(Confirmed)) |>
  filter(count > 10000) |>
  head(1) |>
  pull(ObservationDate)
```

```{r}
MC - MS
```

