---
title: "Solutions: Zachary's karate-club network"
subtitle: "YSC2210 - DAVis with R"
author: "Michael T. Gastner"
output:
  pdf_document:
    citation_package: natbib
    number_sections: false
urlcolor: blue
header-includes: 
    - \usepackage{hyperref}
bibliography: karate.bib
biblio-style: natbib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```

I need the following packages for these exercises.

```{r}
library(ggforce)
library(ggraph)
library(igraph)
library(tidygraph)
library(tidyverse)
```

(1)

    ```{r}
    karate_edges <- read_table(
      "data/soc-karate.zip",
      col_names = c("from", "to"),
      col_types =  list(.default = col_integer()),
      skip = 24
    )
    ```
    
    I assume that all nodes in the network appear at least once in the edge
    list.
    
    ```{r}
    karate_nodes <-
      tibble(node = unique(flatten_int(karate_edges))) |>
      arrange(node)
    karate_netw <- tbl_graph(
      nodes = karate_nodes,
      edges = karate_edges,
      directed = FALSE
    )
    ```

(2)

    ```{r}
    gorder(karate_netw)
    gsize(karate_netw)
    ```

    In conclusion, there are `r gorder(karate_netw)` nodes and
    `r gsize(karate_netw)` edges.
(3) 

    ```{r}
    is_simple(karate_netw)
    ```

    The network is simple.
    Thus, there are no parallel edges or loops.
(4)

    ```{r}
    instructor_id <- 1
    president_id <- 34
    karate_netw <-
      karate_netw |>
      activate(nodes) |>
      mutate(role = case_when(
        node == instructor_id ~ "Instructor",
        node == president_id ~ "President",
        TRUE ~ "Other"
      ))
    ```

(5)

    ```{r}
    top_2_degree <-
      karate_netw |>
      activate(nodes) |>
      mutate(degree = centrality_degree()) |>
      as_tibble() |>
      slice_max(degree, n = 2) |>
      print()
    ```

    The next code chunk answers the question: Are the two nodes with the
    highest degree the president and the instructor?
    
    ```{r}
    setequal(top_2_degree$role, c("President", "Instructor"))
    ```

(6)

    ```{r}
    asp <-
      all_shortest_paths(karate_netw, instructor_id, president_id)$res |>
      print()
    ```

    The distance between president and instructor
    is `r distances(karate_netw, president_id, instructor_id)`.
    The shortest paths from president and instructor go through one of the
    following intermediate nodes:
    
    ```{r}
    sort(map_int(asp, 2))
    ```
    
(7)

    ```{r}
    karate_netw |>
      activate(nodes) |>
      mutate(cluster_coef = local_transitivity()) |>
      filter(role %in% c("Instructor", "President")) |>
      as_tibble()
    ```
   
    The instructor has a higher clustering coefficient than the president.
    Therefore, two nodes adjacent to the instructor are more likely to be
    adjacent themselves than two nodes adjacent to the president.
    This fact can be interpreted as a sign that the instructor participates in
    more tightly knit relations. 
(8)

    ```{r}
    largest_cliques(karate_netw)
    ```

    Hence, there two cliques of size `r clique_num(karate_netw)` with the
    following nodes:
    
    - `r largest_cliques(karate_netw)[[1]] |> sort()`
    - `r largest_cliques(karate_netw)[[2]] |> sort()`
(9)

    ```{r}
    is_connected(karate_netw)
    ```

(10) The next two code chunks demonstrate two different options to calculate
     communities with the Louvain algorithm.

     - The first option is based on `group_louvain()` in the **tidygraph**
       package.
       At the end of the first pipeline, I add identifiers for the
       communities, which I need when I plot the network.
       
       ```{r}
       karate_netw <-
         karate_netw |>
         activate(nodes) |>
         mutate(community_id = group_louvain())
       karate_netw |>
         as_tibble() |>
         group_by(community_id) |>
         group_split() |>
         map(pluck("node"))
       ```
       
     - The second option uses `cluster_louvain()` in the **igraph** package.
     
       ```{r}
       karate_netw |>
         cluster_louvain() |>
         membership() |>
         tibble() |>
         mutate(node = row_number()) |>
         group_by(across(1)) |>
         group_split() |>
         map(pluck("node"))
       ```
       
       ```{r, include=FALSE}
       comm <-
         karate_netw |>
         as_tibble() |>
         group_by(community_id) |>
         group_split() |>
         map(pluck("node"))
       ```
       
     In conclusion, the Louvain algorithm identifies `r length(comm)`
     communities with the following nodes:
     
     - `r comm[[1]]`
     - `r comm[[2]]`
     - `r comm[[3]]`
     - `r comm[[4]]`
(11) First, I create the network layout.

     ```{r}
     karate_layout <- create_layout(karate_netw, layout = "stress")
     layout_attributes <- attributes(karate_layout)
     node_positions <- read_csv("karate_node_positions.csv")
     karate_layout <-
       karate_layout |>
       select(-x, -y) |>
       left_join(node_positions, by = "node") |>
       select(x, y, everything())
     attributes(karate_layout) <- layout_attributes
     ```

     Then, I create the Bezier curve.
     
     ```{r}
     bezier <- tribble(
       ~x, ~y,
       0.2, 1.3,
       -1.6, 0.0,
       0.7, -0.6
     )
     ```
     
     Finally, I make the plot.
     
     ```{r, fig.width = 7.5, fig.height = 6.7}
     ggraph(karate_layout) +
       geom_mark_hull(
         aes(
           x,
           y,
           fill = as.factor(community_id),
           label = str_c("Community ", community_id)
         ),
         colour = NA,
         con.colour = "grey",
         show.legend = FALSE
       ) +
       geom_edge_link(colour = "grey50") +
       geom_node_point(
         aes(colour = role, shape = role),
         size = 8,
         fill = "white",
         stroke = 2
       ) +
       geom_node_text(aes(label = node)) +
       geom_bezier(aes(x = x, y = y), data = bezier, size = 2) +
       annotate(
         "text",
         x = 0.1,
         y = 1.4,
         label = "Observed split",
         fontface = "bold"
       ) +
       labs(
         title = "Zachary's Karate Club",
         subtitle = "Communities assigned by Louvain algorithm",
         caption = "Source: https://networkrepository.com"
       ) +
       scale_y_continuous(expand = expansion(mult = c(0.1, 0.06))) +
       scale_colour_brewer(name = "Role", palette = "Dark2") +
       scale_shape_manual(
         name = "Role",
         values = 23:21,
         guide = guide_legend(override.aes = list(size = 7))
       ) +
       theme(
         legend.key.height = unit(9, "mm"),
         legend.key.width = unit(9, "mm"),
         legend.position = c(1, 1),
         legend.justification = c(1, 1)
       )
     ggsave("karate_community_split.pdf", width = 7.5, height = 6.7)
     ```

(12) The Louvain algorithm is agnostic about which nodes ended up on which
     side when the conflict arose in the karate club.
     Still, there appears to be a relation between the communities detected by
     the Louvain algorithm and the empirically observed split of the club.
     With only one swap (nodes 9 and 10), the observed split almost perfectly
     follows the division between communities 1 and 4 on the one hand, and
     communities 2 and 3 on the other hand.