---
title: "Solutions: Lists and the map-family"
subtitle: "YSC2210 - DAVis with R"
author: "Michael T. Gastner"
output:
  pdf_document:
    citation_package: natbib
    number_sections: true
header-includes: 
    - \usepackage{hyperref}
bibliography: pyramid.bib
biblio-style: natbib
---

```{r include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```

The following packages are needed for this exercise:

```{r}
library(ggrepel) # For geom_text_repel()
library(DescTools) # For StripAttr()
library(kableExtra)
library(tidyverse)
```


# Relation between tibbles, data frames and lists

(a)

    ```{r}
    musicians_tbl <- tribble(
      ~name, ~band, ~instrument,
      "Keith", "Stones", "guitar",
      "John", "Beatles", "guitar",
      "Paul", "Beatles", "bass"
    )
    ```
    
(b)

    ```{r}
    musicians_dfr <- data.frame(
      name = c("Keith", "John", "Paul"),
      band = c("Stones", "Beatles", "Beatles"),
      instrument = c("guitar", "guitar", "bass")
    )
    ```
    
    Here are two options for determining whether all columns are character
    vectors.
    
    (i) Elementary solution:
    
        ```{r}
        is.character(musicians_dfr$name) &
          is.character(musicians_dfr$band) &
          is.character(musicians_dfr$instrument)
        ```
        
    (ii) Less hard-coded. This option takes advantage of data frames
         being lists, a fact that we will learn by the end of this exercise.
    
         ```{r}
         map_lgl(musicians_dfr, is.character) |> all()
         ```
        
(c)

    ```{r}
    musicians_list <- list(
      name = c("Keith", "John", "Paul"),
      band = c("Stones", "Beatles", "Beatles"),
      instrument = c("guitar", "guitar", "bass")
    )
    ```
    
(d) The straightforward solution is to run
    
    ```{r, eval=FALSE}
    class(musicians_tbl)
    class(musicians_dfr)
    class(musicians_list)
    # etc.
    ```
    
    and then to replace `class()` by `is_tibble()`, `is.data.frame()` and
    `is.list()`.
    Here is a slightly less repetitive alternative with nicer output.
    First, I create a list containing the variables of interest.
    The function `get()` turns a character string into a variable name.

    ```{r}
    objects <-
      c("musicians_tbl", "musicians_dfr", "musicians_list") |>
      set_names() |>
      map(get)
    ```

    The function `classes()` returns the elements of `class()` as a list.
    This function is useful for printing the classes as a list column later
    on.
    
    ```{r}
    classes <- function(x) list(class(x))
    ```
    
    Next, I store `functions` as a tibble so that the column names are picked
    up in the pipeline below that prints a table.
    
    ```{r}
    functions <- c("classes", "is_tibble", "is.data.frame", "is.list") |>
      set_names() |>
      map_dfr(identity)
    objects |>
      map_dfr(~ map(functions, exec, .)) |>
      mutate(object = names(objects)) |>
      relocate(object) |>
      kbl() |>
      kable_styling(c("striped", "condensed", "bordered"), full_width = FALSE)
    ```

    - The results for `musicians_list` are as expected: its class is `"list"`,
      and it is not a tibble or a data frame.
    - The results for `musicians_dfr` are more surprising: its class is
      `data.frame`, but `is.list()` also returns `TRUE`.
      We also learn that a data frame is usually not a tibble.
    - A tibble is also always a data frame and a list.
(e)

    ```{r}
    attributes(musicians_tbl)
    attributes(musicians_dfr)
    attributes(musicians_list)
    ```
    
    All three objects have the same `names` attribute, but the tibble and 
    the data frame also have attributes `class` and `row.names`.
(f) We can perform matrix-style subsetting on `musicians_tbl` and
    `musicians_dfr`, but not on `musicians_list`.
    
    ```{r error=TRUE}
    musicians_tbl[2, 3]
    musicians_dfr[2, 3]
    musicians_list[2, 3] # This line throws an error
    ```

(g) There are two options to remove the attributes.

    (i) Elementary solution:
    
        ```{r}
        attr(musicians_tbl, "class") <- NULL
        attr(musicians_dfr, "class") <- NULL
        attr(musicians_tbl, "row.names") <- NULL
        attr(musicians_dfr, "row.names") <- NULL
        ```
    
    (ii) We can use `StripAttr()` from the **DescTools** package to remove
         the attributes `class` and `row.names` in a single step.

         ```{r}
         musicians_dfr <- StripAttr(musicians_dfr, c("class", "row.names"))
         musicians_tbl <- StripAttr(musicians_tbl, c("class", "row.names"))
         identical(musicians_tbl, musicians_dfr) &
           identical(musicians_tbl, musicians_list)
         ```
    
    In conclusion, tibbles and data frames are lists with two additional
    attributes: `class` and `row.names`.
    Thanks to these attributes, we can perform matrix-style subsetting.
    If we remove these attributes from a tibble or data frame, they turn into
    a simple list.


    
# Map-family of functions

(a)
    (i)

        ```{r}
        map_dbl(mtcars, mean)
        ```

        In the output, the name of the column in `mtcars` appears above the
        corresponding mean.

    (ii)
    
         ```{r}
         map_chr(nycflights13::flights, typeof)
         ```
        
    (iii)
    
          ```{r}
          map_lgl(gss_cat, is.factor)
          ```
    
    (iv)
    
         ```{r}
         map_int(iris, n_distinct)
         ```
         
(b) First, the function `pmap()` takes the first element of each list element
    in the first argument (`list(10, 1:10, 2:11)`) and applies the function
    in the second argument (`runif`).
    That is, the first element returned by `pmap()` is a vector containing 10
    random numbers that are uniformly distributed between 1 and 2.
    
    Next, the function takes the second element of each list element in
    `list(10, 1:10, 2:11)`.
    Because the first list element only has one element (`10`), this number
    is recycled.
    That is, the second element returned by `pmap()` is a vector containing
    10 random numbers that are randomly distributed between 2 and 3.

    This procedure repeats until the last elements in each element of
    `list(10, 1:10, 2:11)` is reached.
    Thus, the last element returned by `pmap()` is a vector containing 10
    random numbers that are randomly distributed between 10 and 11.
(c) Here is the code chunk rewritten.

    ```{r}
    map_dfr(
      list(beaver1, beaver2),
      ~ glm(activ ~ temp, data = ., family = binomial) |>
        pluck(coef)
    )
    ```
    
    The function call `glm(activ ~ temp, data = ., family = binomial)`
    returns the slope and intercept of a generalised linear model.
    To be specific, the model is logistic regression because of the argument
    `family = binomial`.
    The next step in the pipe (`pluck(coef)`) extracts the regression
    coefficients (i.e.\ intercept and slope) as a numeric vector.
    The function `map_dfr()` takes the data frames `beaver1` and `beaver2` and
    puts the regression coefficients into a tibble with two rows (one for
    `beaver1` and another for `beaver2`) and two columns (one for the
    intercept and another for the slope).


# Tibbles with list columns

(a) There are (at least) two options.

    (i) Here is a **dplyr**-option:

        ```{r}
        starwars |>
          select(where(is.list)) |>
          names()
        ```

    (ii) Here is an alternative using `map_lgl()`:
    
         ```{r}
         starwars |>
           map_lgl(is.list) |>
           {
             \(x) names(starwars)[x]
           }()
         ```
    
(b) Again, there are (at least) two options.

    (i) We can use `map_lgl()`:
    
        ```{r}
        starwars |>
          filter(map_lgl(starships, ~ ("Millennium Falcon" %in% .))) |>
          pull(name)
        ```
        
    (ii) We can use `rowwise()` from the **dplyr** package:
    
         ```{r}
         starwars |>
           rowwise() |>
           mutate(flew_millenium_falcon = ("Millennium Falcon" %in% starships)) |>
           filter(flew_millenium_falcon) |>
           pull(name)
         ```

(c)

    ```{r}
    starwars |>
      pull(films) |>
      flatten_chr() |>
      unique()
    ```

(d)

    ```{r}
    sw_gender <-
      starwars |>
      select(name, gender, films) |>
      unnest_longer(films) |>
      group_by(films) |>
      summarise(prop_fem = mean(gender == "feminine", na.rm = TRUE))
    ```

(e)

    ```{r fig.width=5.5, fig.height=4.5, fig.align='center', out.width='67%'}
    release <- tribble(
      ~movie,                   ~year,
      "A New Hope",              1977,
      "The Empire Strikes Back", 1980,
      "Return of the Jedi",      1983,
      "The Phantom Menace",      1999,
      "Attack of the Clones",    2002,
      "Revenge of the Sith",     2005,
      "The Force Awakens",       2015
    )
    sw_gender <-
      sw_gender |>
      left_join(release, by = c("films" = "movie"))
    ggplot(sw_gender, aes(year, prop_fem, label = films)) +
      geom_smooth(method = "lm") +
      geom_point() +
      geom_text_repel() +
      labs(
        title = "Gender Ratio of Characters in Star Wars Movies",
        caption = "Source: Star Wars API (https://swapi.dev/)",
        x = "Year",
        y = "Feminine characters"
      ) +
      scale_y_continuous(labels = scales::percent_format(accuracy = 1))
    ```

(f) The regression line has a positive slope, so the percentage of feminine
    characters has an increasing tendency.
    However, with a p-value of 0.0808, the increase is not statistically
    significant at an alpha level of 0.05.
    
    ```{r}
    model <- lm(prop_fem ~ year, data = sw_gender)
    broom::glance(model)
    ```
    

    