---
title: "Solutions: Flights from New York City"
subtitle: "YSC2210 - DAVis with R"
author: "Michael T. Gastner"
output:
  bookdown::pdf_document2:
  number_sections: false
toc: false
header-includes: 
  - \usepackage{hyperref}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, fig.align = "center")
```

I need the following packages for these solutions:

```{r}
library(nycflights13)
library(tidyverse)
```


(1)

    (a) The following pipeline returns a numeric vector of length 1.
        Instead of `nrow()`, we may use `count()` or `tally()`, which return
        a tibble with 1 row and 1 column.

        ```{r}
        flights |>
          filter(arr_delay >= 120) |>
          nrow()
        ```

    (b)
    
        ```{r}
        flights |>
          filter(origin == "JFK", dest %in% c("IAH", "HOU")) |>
          nrow()
        ```
    
    (c) Midnight appears as `2400` in `dep_time`.
    
        ```{r}
        flights |>
          filter(dep_time <= 600 | dep_time == 2400) |>
          nrow()
        ```
    
(2)

    (a)
    
        ```{r}
        flights_with_na_dep_time <-
          flights |>
          filter(is.na(dep_time))
        nrow(flights_with_na_dep_time)
        ```
    
    (b) The return value of the next pipeline answers: "Do all of these
        flights also have a missing `arr_time`?"
    
        ```{r}
        flights_with_na_dep_time |>
          pull(arr_time) |>
          is.na() |>
          all()
        ```
    
    (c) Flights that have missing departure times are probably cancelled
        flights in the sense that they never departed.
        The fact that none of these flights has a known arrival time is
        circumstantial evidence that supports this hypothesis.
    
(3)

    ```{r}
    flights |>
      group_by(dest) |>
      summarise(mean_air_time = mean(air_time, na.rm = TRUE)) |>
      slice_max(order_by = mean_air_time, n = 10)
    ```
    
(4)
    
    ```{r}
    flights |>
      mutate(speed_mph = 60 * distance / air_time) |>
      select(air_time, distance, speed_mph, dest) |>
      slice_min(speed_mph, n = 10)
    ```
    
(5)

    ```{r}
    select(flights, ends_with("dep_time"))
    ```

(6) We can use the function `contains()`.

    ```{r}
    dep_info <- 
      flights |>
      select(contains("dep")) |>
      print()
    ```

(7)

    (a)

        ```{r}
        dep_info <-
          dep_info |>
          mutate(diff_time = dep_time - sched_dep_time) |>
          print()
        ```
    
        ```{r echo=FALSE}
        n_mismatches <- dep_info |>
          filter(diff_time != dep_delay) |>
          nrow()
        ```
        
    (b) At first glance, we would expect `diff_time` and `dep_delay` to be
        equal.
        However, as the next code chunk shows, there are
        `r format(n_mismatches, big.mark = ",")` flights for which `diff_time`
        and `dep_delay` differ.
        
        ```{r}
        dep_info |>
          filter(diff_time != dep_delay) |>
          nrow()
        ```

        The explanation is that times are treated as if they were measured in
        a decimal system.
        However, one hour has only 60 minutes, not 100.
    (c)
    
        ```{r}
        dep_info <-
          dep_info |>
          mutate(
            dep_hour = dep_time %/% 100,
            dep_min = dep_time %% 100,
            sched_dep_hour = sched_dep_time %/% 100,
            sched_dep_min = sched_dep_time %% 100,
            diff_time =
              60 * (dep_hour - sched_dep_hour) + dep_min - sched_dep_min
          ) |>
          select(dep_time, sched_dep_time, diff_time, dep_delay)
        ```
    
        In most rows, we now have `diff_time == dep_delay`, but not in all.
        Let us print the first few rows in which `diff_time` and `dep_delay`
        are different.
    
        ```{r}
        filter(dep_info, diff_time != dep_delay)
        ```
    
        In each of these rows, the flight departed the following day.
        A proper fix would require working with dates and times.
        We will learn more about R's date-time structures later in this
        course.
        Here is a quick and dirty fix that does the job in the present
        example.
        We add 1440 to the time difference, which is the number of minutes
        in a day.
    
        ```{r}
        dep_info <-
          dep_info |>
          mutate(
            dep_hour = dep_time %/% 100,
            dep_min = dep_time %% 100,
            sched_dep_hour = sched_dep_time %/% 100,
            sched_dep_min = sched_dep_time %% 100,
            diff_time_tmp =
              60 * (dep_hour - sched_dep_hour) + dep_min - sched_dep_min,
            diff_time = if_else(
              diff_time_tmp == dep_delay,
              diff_time,
              diff_time_tmp + 1440
            )
          ) |>
          select(dep_time, sched_dep_time, diff_time, dep_delay)
        ```
    
        Now we find the expected equality between `diff_time` and `dep_delay`.
    
        ```{r}
        all(dep_info$diff_time == dep_info$dep_delay, na.rm = TRUE)
        ```
    
(8)

    ```{r fig.width=3.75, fig.height=3, out.width='45%'}
    daily_info <-
      flights |>
      group_by(year, month, day) |>
      summarise(
        mean_dep_delay = mean(dep_delay, na.rm = TRUE),
        prop_cancelled = mean(is.na(dep_time))
      )
    ggplot(daily_info, aes(mean_dep_delay, prop_cancelled)) +
      geom_point(alpha = 0.5) +
      labs(
        title = "Daily Cancellations vs. Departure Delay",
        caption = "Source: RITA, Bureau of Transportation Statistics",
        x = "Mean departure delay (minutes)",
        y = "Cancelled flights"
      ) +
      scale_y_continuous(labels = scales::percent)
    ```

    The percentage of cancelled flights tends to increase as the mean departure
    delay increases.
    
(9) In hour 1, there is only one flight, and its `dep_delay` equals `NA`.
    Thus, `mean_dep_delay` in hour 1 is also `NA`.
    Let us remove this row with `drop_na()`.
    I set the breaks along the x-axis at multiples of 6 because hours follow a
    duodecimal system.

    ```{r fig.width=3.75, fig.height=3, out.width='45%'}
    hourly_info <-
      flights |>
      group_by(hour) |>
      summarise(mean_dep_delay = mean(dep_delay, na.rm = TRUE)) |>
      drop_na()
    ggplot(hourly_info, aes(hour, mean_dep_delay)) +
      geom_line() +
      geom_point() +
      labs(
        title = "Departure Delay vs. Hour of Day",
        caption = "Source: RITA, Bureau of transportation statistics",
        x = "Hour",
        y = "Mean departure delay (minutes)"
      ) +
      scale_x_continuous(breaks = c(6, 12, 18))
    ```

    If we want to avoid departure delays, we should fly early in the morning.

