---
title: "Solutions: Singapore population pyramid"
subtitle: "YSC2210 - DAVis with R"
author: "Michael T. Gastner"
output:
  pdf_document:
    citation_package: natbib
header-includes: 
    - \usepackage{hyperref}
---

```{r include=FALSE}
knitr::opts_chunk$set(message=FALSE)
```

First, we load the packages that we need for this task.

```{r}
library(readxl)
library(scales)
library(tidyverse)
```

(1) The most convenient option is to import sheet 'T2', which contains
    'Singapore Residents By Age Group'.
    It would be possible to use sheet 'T3', which has more fine-grained data
    ('Singapore Residents By Single Year'), but the additional level of detail
    is not useful in the present context.
    
    Missing data (e.g.\ in the column for the year 1999 and residents in the
    age bracket '85--89 years') is encoded by the character string 'na'.
    
    The relevant rows to import are 38--61 (male residents) and 63--86 (female
    residents).
    However, to set the correct argument for `slice()` in the code chunk
    below, we must skip the first 10 rows, which contain metadata, and one
    more row for column header.
    Thus, we must subtract 11 from the row numbers in the Excel file.
    
    ```{r}
    sg <- read_excel(
      "outputFile.xlsx",
      sheet = "T2",
      na = "na",
      skip = 10
    ) |>
      slice(27:50, 52:75)
    ```
    
(2)

    ```{r}
    sg <- select(sg, age = `Data Series`, ends_with("0"))
    ```

(3)

    ```{r}
    sg <- mutate(
      sg,
      age = str_replace(age, "(.*) Years & Over$", ">\\1") |>
        str_remove(" Years") |>
        (\(x) {
          factor(x, levels = unique(x))
        })()
    )
    ```

(4)

    ```{r}
    sg <- mutate(sg, sex = rep(c("male", "female"), each = nrow(sg) / 2))
    ```

(5) The following helper function, named `fill_highest_age_group()`,
    identifies the first age group with `NA` (the variable called `x`) and the
    last age group without `NA` (`y`) in an input vector `v`.
    The function then shifts the value from `y` to `x`.

    ```{r}
    fill_highest_age_group <- function(v) {
      x <-
        which(is.na(v)) |>
        first()
      y <-
        which(!is.na(v)) |>
        last()
      v[x] <- v[y]
      v
    }
    ```

    The next helper function, `fill_highest_age_groups()`, where `groups` is
    in plural, performs `fill_highest_age_group()`, where `group` is in
    singular, for both genders.
    
    ```{r}
    fill_highest_age_groups <- function(w) {
      male <- w[1:(length(w) / 2)]
      female <- w[(length(w) / 2 + 1):length(w)]
      c(fill_highest_age_group(male), fill_highest_age_group(female))
    }
    ```
    
    We can now mutate `age` with `fill_highest_age_groups()`.
    
    ```{r}
    sg <- mutate(sg, across(ends_with("0"), fill_highest_age_groups))
    ```
    
(6)

    ```{r}
    sg <- filter(sg, str_starts(age, ">", negate = TRUE) | age == ">90")
    ```
    
(7)

    ```{r}
    sg <- pivot_longer(
      sg,
      -c(age, sex),
      names_to = "year",
      values_to = "pop",
      values_drop_na = TRUE
    )
    ```

(8) I flip the sign of the male population so that the male bars will extend
    into the right direction.
    
    ```{r}
    sg <- mutate(sg, pop = if_else(sex == "male", -pop, pop))
    ```
    
    I change axis tick labels to non-negative numbers with
    `scale_x_continuous(labels = ~ comma(abs(.)))`
    
    The ColorBrewer palette 'Set1' starts with red and blue, which is a common
    colour code for male and female in population pyramids.
    The colours are a little bit too intense for my taste; thus, I reduce the
    alpha-level.

    Finally, I use `theme()` to shift the legend to the bottom right.

    ```{r fig.width = 8, fig.height = 6}
    ggplot(sg, aes(pop, age, fill = sex)) +
      geom_col(alpha = 0.75) +
      labs(
        x = "Population",
        y = "Age (years)",
        fill = "Sex",
        title = "Singapore Resident Population between 1960 and 2020",
        caption = "Source: Department of Statistics Singapore",
      ) +
      scale_x_continuous(labels = ~ comma(abs(.))) +
      scale_fill_brewer(palette = "Set1", direction = -1) +
      facet_wrap(~year, nrow = 2) +
      theme(
        legend.position = c(1, 0),
        legend.justification = c(1, 0)
      )
    ```

    A figure width of 8 and figure height of 6 leads to a reasonable font size
    (e.g.\ in the axis labels).
    For your information, here is how you save the plot to a file.
    
    ```{r}
    ggsave("sg_pyramid.pdf", width = 8, height = 6)
    ```
    
(9) Since 1960, Singapore's population has become older, as we can see from
    the age groups represented by the topmost bars in the panels.
    The bar plots also have undergone a dramatic change in shape.
    In 1960, the age structure resembled a triangular shape in which the largest
    age group was at the bottom, and the bars tapered off towards the top.
    As time went by, a noticeable bulge formed in the middle-aged group.
    In 2020, there are more Singaporeans aged between 65 and 69 than between
    0 and 4 years.
    Towards older age groups, all pyramids are slightly asymmetric because
    women tend to live longer than men.