---
title: "Solutions: Relational data for country-level statistics"
subtitle: "YSC2210 - DAVis with R"
author: "Michael T. Gastner"
output:
  pdf_document:
    citation_package: natbib
header-includes: 
    - \usepackage{hyperref}
bibliography: preston.bib
biblio-style: natbib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```

I need the following packages for this exercise.

```{r}
library(countrycode)
library(gapminder)
library(Hmisc)
library(readxl)
library(rlang)
library(tidyverse)
```

(1) I import the data with `read_excel()`.
    In each input file, the first three lines contain metadata and must be
    skipped.

    ```{r}
    gdp_per_cap <- read_excel(
      "data/API_NY.GDP.PCAP.KD_DS2_en_excel_v2_3731742.xls",
      skip = 3
    )
    life_exp <- read_excel(
      "data/API_SP.DYN.LE00.IN_DS2_en_excel_v2_3731513.xls",
      skip = 3
    )
    pop <- read_excel(
      "data/API_SP.POP.TOTL_DS2_en_excel_v2_3759026.xls",
      skip = 3
    )
    ```

(2) The next command answers the question: are the '`Country Code`' columns in
    all three spreadsheets identical?

    ```{r}
    all(
      gdp_per_cap$`Country Code` == life_exp$`Country Code`,
      gdp_per_cap$`Country Code` == pop$`Country Code`
    )
    ```
    
(3) First, I write a utility function `pivot_longer_wb()` that pivots any of
    the World Bank tibbles (passed as variable `tb`) into long format.

    ```{r}
    pivot_longer_wb <- function(tb, name) {
      tb |>
        select(-starts_with("Indicator")) |>
        pivot_longer(
          matches("^\\d{4}$"), # String with exactly 4 digits
          names_to = "year",
          values_to = {{ name }},
          names_transform = list(year = as.integer)
        )
    }
    ```

    Next, I pivot all three World Bank tibbles into long format.
    The function `llist()` from the **Hmisc** package creates a list
    in which the elements acquire the names of the arguments.
    Then, I use `imap()` so that both the list elements and their names are
    passed as arguments to `pivot_longer_wb()`.
    Finally, I use `reduce()` from the **purrr** package to iteratively apply
    `left_join()`.
    
    ```{r}
    wb <-
      llist(pop, life_exp, gdp_per_cap) |>
      imap(pivot_longer_wb) |>
      reduce(left_join, by = c("Country Name", "Country Code", "year"))
    ```
    
    ```{r, include=FALSE}
    n_initial_rows_in_wb <- nrow(wb)
    ```
    
    There are `r n_initial_rows_in_wb` rows in `wb`.
    
    ```{r}
    nrow(wb)
    ```
    
    Here are the first few rows of `wb`:
    
    ```{r}
    wb
    ```
    
    By looking at the last few rows, it becomes clear that not all elements
    in the column `gdp_per_cap` are `NA`:
    
    ```{r}
    tail(wb)
    ```

(4) Here are all countries that cannot be matched with an ISO 3166-1 alpha-3
    code.

    ```{r}
    anti_join(wb, codelist, by = c("Country Code" = "iso3c")) |>
      pull(`Country Name`) |>
      unique()
    ```

    Most country codes that cannot be matched represent groups of countries
    instead of individual countries.
    The exceptions are:
    
    - *Channel Islands*: they are part of the British crown but have some
      rights to establish direct relations with foreign governments because
      they are not part of the United Kingdom.
    - *Kosovo*: Kosovo's status as an independent country is disputed.
      Thus, it is to be expected that some data bases (such as R's
      **countrycode** package) do not include it.
(5)

    ```{r}
    wb <- semi_join(wb, codelist, by = c("Country Code" = "iso3c"))
    ```
    
    The number of rows decreased from `r n_initial_rows_in_wb`
    to `r nrow(wb)`.
    
    ```{r}
    nrow(wb)
    ```
    
(6)

    ```{r}
    missing_values <-
      wb |>
      filter(is.na(gdp_per_cap) | is.na(life_exp) | is.na(pop)) |>
      group_by(year) |>
      count()
    missing_values
    ```

(7)

    ```{r fig.width=5, fig.height=3.5, fig.align='center', out.width='70%'}
    ggplot(missing_values, aes(year, n)) +
      geom_point() +
      geom_line() +
      labs(
        x = "Year",
        y = "Number of countries",
        title = "Countries with Missing Information in World Bank Data",
        caption = "Source: World Bank"
      )
    ```

    The number of missing countries decreased slowly from
    `r missing_values$na_countries[1]` in (`r missing_values$year[1]`) to
    $\approx$ 25 countries since the start of the millenium
    In 2019 and 2020, the number suddenly jumps to much higher values because
    none of the data are known yet.
    Presumably, it takes a few years for the World Bank to collate the data.

    A closer look at the data reveals that many of the missing entries in the
    first few years come from countries that did not exist at that time.
    For example, Ukraine only gained independence in 1991, and the World Bank
    apparently did not want to speculate how much of the Soviet Union's GDP
    can be attributed to the Ukraine prior to 1987.
    
(8)

    ```{r}
    gap <-
      gapminder_unfiltered |>
      filter(country != "Netherlands Antilles") |>
      mutate(country_code = countrycode(country, "country.name", "iso3c"))
    ```
    
    The next code answers the question: are there countries in `gap` without a
    country code?
    
    ```{r}
    anyNA(gap$country_code)
    ```
    
    Finally, I check whether there are countries with the same country code.
    
    ```{r}
    gap |>
      group_by(country_code) |>
      summarise(n_countries = n_distinct(country)) |>
      transmute(code_is_duplicated = (n_countries > 1)) |>
      any()
    ```
    
(9) Here are all countries in `gap` that are not in `wb`.

    ```{r}
    gap |>
      anti_join(wb, by = c("country_code" = "Country Code")) |>
      pull(country) |>
      unique()
    ```

    French Guiana, Guadeloupe, Martinique and La Réunion are French overseas
    departments.
    Presumably, the World Bank lumps these data together with the rest of
    France.
    Taiwan is not recognised as an independent country by the World Bank.
    
    Next, I determine all countries in `wb` that are not in `gap`.
    
    ```{r}
    wb |>
      anti_join(gap, by = c("Country Code" = "country_code")) |>
      pull(`Country Name`) |>
      unique()
    ```
    
    Most countries missing from `gap` are small except South Sudan, which only
    gained independence in 2011.
(10)
    I merge the GDP data for 2007 in the World Bank data set and in
    Gapminder.
    
    ```{r}
    wb_2007 <-
      wb |>
      filter(year == 2007) |>
      select("Country Code", "Country Name", gdp_wb = gdp_per_cap)
    gap_2007 <-
      gap |>
      filter(year == 2007) |>
      select(country_code, gdp_gap = gdpPercap)
    wb_gap <-
      inner_join(
        wb_2007,
        gap_2007,
        by = c("Country Code" = "country_code")
      ) |>
      print()
    ```

    Here are the countries in `wb_2007` that are not in `wb_gap`:
    
    ```{r}
    wb_2007 |>
      anti_join(wb_gap, by = "Country Code") |>
      pull(`Country Name`)
    ```
    
    In addition to the countries that I found to be generally absent from the
    Gapminder data in (9), Gapminder does not have data for Luxembourg in 2007
    either.
(11)
    I add a row that contains the percentage difference.

    ```{r}
    wb_gap <-
      wb_gap |>
      mutate(pct_diff = 100*gdp_gap/gdp_wb - 100) |>
      print()
    ```

(12)
    Here are the 5 largest values in `pct_diff`.
    
    ```{r}
    slice_max(wb_gap, pct_diff, n = 5)
    ```
    
    Here are the 5 smallest values.
    
    ```{r}
    slice_min(wb_gap, pct_diff, n = 5)
    ```

    The differences between the two data sets are substantial, for some
    countries more than a factor 2.