---
title: "Solutions: Solar cycles"
subtitle: "YSC2210 - DAVis with R"
author: "Michael T. Gastner"
output:
  pdf_document:
    citation_package: natbib
urlcolor: blue
header-includes: 
    - \usepackage{hyperref}
bibliography: preston.bib
biblio-style: natbib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```

I need the following packages for these exercises:

```{r}
library(lubridate)
library(tidyverse)
```

I also use `summarise_by_time()` from the **timetk** package.
Because I only use this function once, I do not load the entire package.
Instead, I use the double colon operator: `timetk::summarise_by_time()`.

(1) Some numbers are preceded by whitespace.
    In the next code chunk, I remove leading whitespace with the argument
    `trim_ws = TRUE`.

    ```{r}
    spots <-
      read_delim(
        "SN_d_tot_V2.0.csv",
        delim = ";",
        col_names = c(
          "year",
          "month",
          "day",
          "year_fraction",
          "ss_number",
          "sd",
          "n_obs",
          "definitive"
        ),
        na = c("-1", "-1.0"),
        trim_ws = TRUE
      ) |>
      print()
    ```
    
(2)

    ```{r}
    spots <-
      spots |>
      mutate(date = make_date(year, month, day)) |>
      print()
    ```

(3) The next pipeline prints the latest date with a sun spot number equal to
    `NA`.

    ```{r}
    spots |>
      filter(is.na(ss_number)) |>
      pull(date) |>
      last() |>
      stamp("Tuesday, 3 December, 2019", quiet = TRUE)()
    ```

(4) The output from the next pipeline answers the question: are all dates
    consecutive?
    After `mutate()`, the last element in the column
    `is_next_date_consecutive` is `NA` because there is no next date.
    I use `slice_head(-1)` to remove this value from
    `is_next_date_consecutive`.

    ```{r}
    spots |>
      mutate(
        date_diff = date %--% lead(date) / ddays(1),
        is_next_date_consecutive = (date_diff == 1)
      ) |>
      slice_head(n = -1) |>
      pull(is_next_date_consecutive) |>
      all()
    ```

(5)

    ```{r}
    spots <-
      spots |>
      filter(year >= 1975) |>
      timetk::summarise_by_time(
        date,
        .by = "month",
        mean_daily_ss_number = mean(ss_number)
      ) |>
      print()
    ```

(6)

    ```{r}
    solar_cycles <-
      read_csv("solar_cycles.csv") |>
      mutate(
        `Start (Minimum)` = ym(`Start (Minimum)`),
        Maximum = ym(Maximum)
      ) |>
      glimpse()
    ```

(7)

    ```{r}
    solar_cycles <-
      solar_cycles |>
      mutate(
        rise_in_years = round(

          # Either:
          # (`Start (Minimum)` %--% Maximum) / dyears(1),
          # or:
          time_length(`Start (Minimum)` %--% Maximum, unit = "year"),
          digits = 1
        ),
        is_rise_equal = (`Time of Rise (years)` == rise_in_years)
      )
    table(solar_cycles$is_rise_equal)
    ```
    
    There are two solar cycles in which the numbers do not match.

    ```{r}
    solar_cycles |>
      filter(!is_rise_equal) |>
      select(
        "Solar Cycle",
        "Start (Minimum)",
        "Maximum",
        "Time of Rise (years)",
        rise_in_years
      )
    ```

    In the last column, only the last decimals differ.
    It is unclear how Wikipedia calculated the numbers.
    One might suspect that the difference is caused by `dyears(1)` having a
    duration of 365.25 days instead of 365 days.
    However, it does not explain the difference in the last row of the
    following tibble.

    ```{r}
    solar_cycles |>
      filter(!is_rise_equal) |>
      mutate(
        rise_in_units_of_365_days =
          (`Start (Minimum)` %--% Maximum) / ddays(365),
        rise_in_units_of_dyears =
          (`Start (Minimum)` %--% Maximum) / dyears(1)
      ) |>
      select(contains("rise", ignore.case = TRUE)) |>
      glimpse()
    ```

(8)

    ```{r}
    solar_cycles <-
      solar_cycles |>
      filter(`Start (Minimum)` >= "1975-01-01") |>
      print()
    ```

(9) After `mutate()` the last element in the column `mid_date` is `NA` because
    it is not possible to determine the middle of a time interval if the end
    point of the interval is unknown.
    I remove the last row with `slice_head(-1)`.

    ```{r}
    solar_cycles <-
      solar_cycles |>
      rename(
        number = "Solar Cycle",
        start = "Start (Minimum)"
      ) |>
      mutate(
        number = str_extract(number, "\\d+$"),
        mid_date = as_date(int_end(start %--% lead(start) / 2))
      ) |>
      slice_head(n = -1) |>
      print()
    ```

(11) I make points semitransparent with `alpha = 0.5` so that overlapping
     points are visible.

     ```{r fig.width=5, fig.height=4}
     ggplot(spots, aes(date, mean_daily_ss_number)) +
       geom_point(colour = "deepskyblue4", alpha = 0.5) +
       geom_smooth(
         aes(colour = ""), # Add empty colour to obtain a legend
         method = "loess",
         se = FALSE,
         span = 0.1
       ) +
       geom_label(
         aes(
           mid_date,
           label = number,
           fill = "" # Add empty fill to obtain a legend
         ),
         data = solar_cycles,
         y = 0,
         label.r = unit(0, "cm"), # Give label same appearance as legend key
         label.size = 0
       ) +
       scale_colour_manual(values = "black") +
       labs(
         x = "Year",
         y = "Mean daily sunspot number",
         colour = "LOESS fit",
         fill = "Solar cycle",
         title = "Solar Activity",
         subtitle = "Sunspot number averaged by month",
         caption = "Source: Royal Observatory of Belgium"
       ) +
       guides( # Give legend key same appearance as label
         fill = guide_legend(
           keyheight = 1,
           override.aes = aes(label = "21")
         )
       ) +
       theme(
         legend.position = c(1, 1),
         legend.justification = c(0.9, 1),
         legend.spacing = unit(0, "cm")
       )
     ```
     
     ```{r, include=FALSE}
     ggsave("sunspot.pdf", width = 5, height = 4)
     ```
     
(11) Since 1975, the sunspot number has exhibited oscillations with a period
     of approximately 11 years.
     While the frequency appears to remain constant, the amplitude of the
     maximum has decreased.
     The latest complete cycle (2008-2019) has been called "Modern Minimum"
     (<https://en.wikipedia.org/wiki/Solar_cycle#Cycle_history>).
     However, there is nothing unusual about the current activity level when
     we compare it with data from past centuries.
     Much smaller minima have been observed before (e.g.\ during the Maunder
     Minimum between 1645 and 1715).
     