---
title: "Solutions: Emails at a large European research institution"
subtitle: "YSC2210 - DAVis with R"
author: "Michael T. Gastner"
output:
  pdf_document:
    citation_package: natbib
urlcolor: blue
header-includes: 
    - \usepackage{hyperref}
bibliography: eu.bib
biblio-style: natbib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```

I need the following packages for these exercises:

```{r}
library(ggraph)
library(igraph)
library(tidygraph)
library(tidyverse)
```


(1) I import node data and edge data as tibbles.
    The function `read_table()` can import gzipped text files; thus, there is
    no need to explicitly unzip them.

    ```{r}
    eu_nodes <-
      read_table(
        "email-Eu-core-department-labels.txt.gz",
        col_names = c("person", "dept")
      ) |>
      print()
    eu_edges <-
      read_table(
        "email-Eu-core.txt.gz",
        col_names = c("from", "to")
      ) |>
      print()
    ```

(2) I add 1 to all person identifiers.

    ```{r}
    eu_nodes <-
      eu_nodes |>
      mutate(person = person + 1) |>
      print()
    eu_edges <-
      eu_edges |>
      mutate(across(.fns = ~ . + 1)) |>
      print()
    ```


(3) I combine the node and edge information into a `tbl_graph`.

    ```{r}
    eu_netw <-
      tbl_graph(nodes = eu_nodes, edges = eu_edges) |>
      print()
    ```

(4)

    ```{r cache=TRUE}
    ggraph(eu_netw, layout = "kk") +
      geom_edge_bend(
        alpha = 0.1,
        arrow = arrow(length = unit(2, "mm"))
      ) +
      geom_node_point(colour = "brown", alpha = 0.5)
    ```

    There are so many nodes and edges that it is difficult to extract
    information from this 'hairball' plot.
    There is no space for labels, and, even when representing the edges as
    semitransparent lines, there is no way to tell who is sending emails to
    whom.
    Changing the layout does not help much.
    In similar situations, a hive plot may be more useful
    (<https://datavizproject.com/data-type/hive-plot/>).
    However, a hive plot requires that the set of nodes can be split into a
    small number of groups (ideally only two or three).
    Such a split is not evident from the description of EU-email data.
(5) Contract the network by department.

    ```{r}
    contracted_igraph <-
      eu_netw |>
      contract(eu_nodes$dept + 1)
    vertex_attr(contracted_igraph, "dept") <-
      0:(gorder(contracted_igraph) - 1)
    contracted_directed_netw <-
      contracted_igraph |>
      as_tbl_graph() |>
      convert(to_simple) |>
      activate(edges) |>
      mutate(weight = lengths(.tidygraph_edge_index)) |>
      print()
    ```

(6) I terminate the arrow head just outside the label with the `end_cap`
    aesthetic.

    ```{r cache=TRUE}
    ggraph(contracted_directed_netw, layout = "stress") +
      geom_edge_fan(
        aes(colour = weight, end_cap = label_rect(node2.dept)),
        alpha = 0.5,
        arrow = arrow(length = unit(0.2, "cm"))
      ) +
      geom_node_label(aes(label = dept), colour = "brown") +
      theme(legend.position = c(1, 1), legend.justification = c(1, 1))
    ```

    The nodes can be easily identified from this plot, but there are too many
    edges to see clearly which departments are connected to each other.
(7)

    ```{r}
    undirected_simple_netw <-
      contracted_directed_netw |>
      convert(to_undirected) |>
      convert(to_simple) |>
      activate(edges) |>
      mutate(weight = map_int(.orig_data, ~sum(.$weight))) |>
      print()
    ```
    
(8)

    ```{r}
    hierarchy <-
      undirected_simple_netw |>
      convert(to_hierarchical_clusters, weights = weight) |>
      print()
    ```

(9)

    ```{r fig.width=4, fig.height=8, fig.align='center', out.width='0.5\\linewidth'}
    layout <- create_layout(hierarchy, "dendrogram")
    g_hierarchy <- ggraph(layout) +
      geom_edge_elbow() +
      geom_node_text(
        aes(label = dept, filter = leaf),
        nudge_y = 0.8,
        hjust = 1
      ) +
      coord_flip() +
      scale_x_continuous(expand = c(0.01, 0.01)) +
      scale_y_reverse(breaks = 0:max(layout$y)) +
      ggtitle("Department hierarchy") +
      theme(
        panel.background = element_rect(fill = "grey90"),
        panel.grid.major.x = element_line(colour = "white"),
        plot.background = element_rect(fill = "grey90")
      )
    g_hierarchy
    ggsave("hierarchy.pdf", g_hierarchy, width = 4, height = 8)
    ```

(10)
    The dendrogram is more legible than the hairball plot in problem 5.
    Consequently, it is easier to identify connectivity from the dendrogram.
    A disadvantage of the dendrogram is that the lines cannot be interpreted
    as edges in the original network; thus, the dendrogram needs more
    explanation.
    Another disadvantage is that different clustering algorithms may generate
    substantially different dendrograms.
    Therefore, one should not overinterpret details of the dendrogram.
(11)
    The heat map shows a lot more detailed information than the dendrogram.
    In principle, the heat map contains similar information as the
    network plot in (6), but is not as cluttered as our earlier plot.
    A disadvantage compared to the dendrogram is that that the heat map does
    not immediately reveal whether there is a hierarchical structure among
    the departments.
    The numerical alignment of the departments along the axes is also less
    meaningful than the alingment in the dendrogram, where nearby departments
    tend to be more closely connected.

    For documentation, here is how I made the heat map.
    I deliberately did not use any **tidygraph** functions so that I can check
    whether my previous results are plausible.
    
    ```{r, fig.width=7, fig.height=6}
    departments_with_self_edges <-
      eu_edges |>
      left_join(eu_nodes, by = c("from" = "person")) |>
      left_join(eu_nodes, by = c("to" = "person")) |>
      select(dept_from = dept.x, dept_to = dept.y) |>
      count(dept_from, dept_to, name = "weight")
    g_heatmap <-
      ggplot(
        departments_with_self_edges,
        aes(x = dept_from, y = dept_to, fill = weight)
      ) +
      geom_tile(width = 0.9, height = 0.9) +
      geom_hline(
        aes(yintercept = pos),
        data = tibble(pos = seq(0.5, 41.5, by = 1)),
        size = 0.5,
        colour = "grey90"
      ) +
      geom_vline(
        aes(xintercept = pos),
        data = tibble(pos = seq(0.5, 41.5, by = 1)),
        size = 0.5,
        colour = "grey90"
      ) +
      geom_hline(
        aes(yintercept = pos),
        data = tibble(pos = seq(9.5, 39.5, by = 10)),
        size = 1,
        colour = "grey70"
      ) +
      geom_vline(
        aes(xintercept = pos),
        data = tibble(pos = seq(9.5, 39.5, by = 10)),
        size = 1,
        colour = "grey70"
      ) +
      coord_equal() +
      labs(
        title = "Number of People Exchanging Emails between Departments",
        caption = "Source: Yin et al. (2017)",
        x = "Sending Department",
        y = "Receiving Department"
      ) +
      scale_x_continuous(expand = c(0, 0)) +
      scale_y_continuous(expand = c(0, 0)) +
      scale_fill_distiller(
        name = "Number of people\nsending emails",
        trans = "log10",
        palette = "Reds",
        direction = 1
      ) +
      guides(fill = guide_colourbar(
        frame.colour = "black",
        ticks.colour = "black"
      )) +
      theme(
        legend.background = element_rect(fill = "grey90"),
        panel.background = element_rect(fill = "white"),
        plot.background = element_rect(fill = "grey90")
      )
    g_heatmap
    ggsave("heatmap.pdf", g_heatmap, width = 7, height = 6)
    ```
