---
title: "First steps with ggplot2"
author: 'Group 4: Alex Tai, Momo Mbarek, Zhang Zichen'
date: "16/02/2022"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = "center")
```


We need to load the following packages.
```{r, message=FALSE}
library(ggplot2)
library(tidyverse)
```


### 1. Exercise 1 

(a) Import the CSV file as a tibble and remove unwanted columns
    ```{r}
titanic_1 <- read_csv("titanic.csv")

# remove unwanted columns
titanic_1 <- titanic_1[, c("class", "survived")]
titanic_1
    ```
```{r}
?read_csv
```


(b) Make a bar plot of passengers by class
    ```{r, out.width='60%', fig.height=3.5, fig.width=5}
  ggplot(titanic_1, aes(x = class)) +
      geom_bar() + labs(
        title = "Passengers in Titanic by class",
        caption = "Source: Encyclopedia Titanica",
        x = "Class",
        y = "Number of passengers"
        )
    ```


(c) Make a side-by-side bar plot to visualise the dependence of the number of 
survivors on their class
    ```{r, out.width='70%', fig.height=3.5, fig.width=6}
  ggplot(titanic_1, aes(fill = survived, x = class)) +
      geom_bar(position = "dodge") + # make the bar plot side-by-side
      labs(
        title = "Passengers in Titanic by class",
        caption = "Source: Encyclopedia Titanica",
        x = "Class",
        y = "Number of passengers"
        )
    ```


(d) Turn the built-in data set Titanic into a tibble
    ```{r}
  titanic_2 <-
      apply(Titanic, c(1, 4), sum) |>
      as_tibble(rownames = "class") |>
      rename(died = "No", survived = "Yes")
    ```


(e) Add a column to `titanic_2` with the total number of passengers in each class
    ```{r}
  titanic_2$total <- titanic_2$died + titanic_2$survived
    ```
We can see that for `titanic_1`, each passenger is a single data entry which
indicates their individual class and survival status. Meanwhile in `titanic_2`,
there are no entries for individual passengers, and only the total number of 
passengers, the number of survived passengers, and the number of passengers who 
died in each class is included. 


(f) Redo the plot in (b) with `titanic_2`
    ```{r, out.width='60%', fig.height=3.5, fig.width=5}
  ggplot(titanic_2, aes(x = class, y = total)) +
      geom_bar(stat = "identity") +
      labs(
        title = "Passengers in Titanic by class",
        caption = "Source: Encyclopedia Titanica",
        x = "Class",
        y = "Number of passengers"
        )
    ```


### 2. Exercise 2

(a) Change the column Expt to be a factor instead of an integer
    ```{r}
  # load the data
  data(morley)

  # turn the Expt column into a factor
  morley$Expt <- factor(morley$Expt)
    ```


(b) Make a scatter plot of Speed (y-axis) versus experiment number (x-axis)
    ```{r, out.width='60%', fig.width=5, fig.height=3.5}
  ggplot(morley, aes(x = Expt, y = Speed)) + 
      geom_point(
        aes(colour = Expt, 
            show.legend = FALSE)) +
      labs(
        title = "Measured speed in different experiments",
        x = "Experiment number",
        y = "Measured speed of light (km/s)")
    ```


(c) Make another scatter plot with geom_jitter()
    ```{r, out.width='60%', fig.width=5, fig.height=3.5}
  ggplot(morley, aes(Expt, Speed)) +
    geom_jitter(
      width = 0.2,
      height = 0,
      size = 0.8,
      aes(colour = Expt),
      show.legend = FALSE
    ) +
    labs(
      title = "Measured speed in different experiments",
      x = "Experiment number",
      y = "Measured speed of light (km/s)"
    )
    ```
The default width value needs to be changed for `geom_jitter` to make data that
belong one experiment clearly distinctive from data that belong to other
experiments. If this is not done, data points of different experiments would 
cluster together, making it difficult to tell which point belongs to which
experiment and draw insights from the plot.


(d) Briefly describe what a box plot shows in general

* A box plot shows five key statistics of continuous variables 
  grouped by a category. 
* The bottom and top of the box shows the 25th and 75th percentiles of some
  continuous data, while the line in the box shows the median.
* The end of the two whiskers show the minimum and maximum respectively. 
* Any point that lays outside the whiskers is considered an outlier. 


(e) Make a box plot of the measured speeds with one box for each experiment
```{r, out.width='60%', fig.width=5, fig.height=3.5}
  ggplot(morley, aes(Expt, Speed)) + 
  geom_boxplot() +
  labs(
    title = "Speed of light in different experiments",
    x = "Experiment number",
    y = "Measured speed of light (km/s)"
  )
```

(f) Make a violin plot with one violin for each experiment number
```{r, out.width='60%', fig.width=5, fig.height=3.5}
  ggplot(morley, aes(Expt, Speed)) + 
  geom_violin() +
  labs(
    title = "Speed of light in different experiments",
    x = "Experiment number",
    y = "Measured speed of light (km/s)"
  )
```


(g) What is the advantage of a violin plot compared to a box plot? What is the 
disadvantage?

* A violin plot is more advantageous than a box plot in showing the probability
  densities distributed over the entire data range. This is useful when we are
  interested in the detailed data distribution. If the data is bimodal, for
  instance, a violin plot woudl allow us to see the different modes while a
  box plot would not.

* One disadvantage of a violin plot is that it does not show the interquartile
  range or the median precisely. This would make it harder to compare those
  characteristics accross different data groups. 


(h) Make a faceted plot with one histogram for each experiment’s speed 
distribution
```{r, fig.width=4, fig.height=5, out.width='47%'}
  ggplot(morley, aes(x = Speed)) +
  geom_histogram(binwidth = 25) + # change to a suitable bin width
  facet_grid(Expt ~ .) + # make stacked histograms for each experiment number
  labs(
    title = "Speed distribution of each experiment",
    x = "Measured speed of light (km/s)",
    y = "Count"
  )
```

(i) Show the speed distribution of each experiment as a frequency polygon. Do 
not use faceting.

```{r out.width='60%', fig.width=5, fig.height=3.5}
  ggplot(morley, aes(Speed, colour = Expt)) + 
  geom_freqpoly(binwidth = 40) + 
  labs(
    title = "Speed distribution of each experiment",
    x = "Measured speed of light (km/s)",
    y = "Count",
    colour = "Experiment number"
  )
```

(j) What is the advantage of using non-faceted frequency polygons compared to 
faceted histograms? What is the disadvantage?

* The advantage of using frequency polygons instead of faceted histograms is
  that it saves space. Physically, the plot takes one set of axis and lables,
  and represents the information of different subgroups close to each other for
  ease of comparison.

* The disadvantage of using frequency polygons is readability. With the number of
  lines exceeding 3 or 4, is becomes increasingly difficult to 'follow' along each
  separate line as their slopes change. Plotting the lines on top of each other is
  also an area of concern which makes the plot less readable.



