---
title: "Solutions: Sea surface temperature anomaly"
subtitle: "YSC2210 - DAVis with R"
author: "Michael T. Gastner"
output:
  html_document
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, fig.align = "center")
```

I need the following packages for these solutions.

```{r}
library(grid)
library(stars)
library(tidyverse)
library(tmap)
```

The packages **envirem**, **knitr**, **raster** and **smoothr** are each only
needed once.
Therefore, I will not load the whole packages.
Instead, I will access the relevant functions in these packages with the
double-colon operator.

(1) In the next code chunk, specify the resolution of the temperature
    measurements that you want to use.
    I recommend setting the variable called `pixels` to `"360x180"` during
    code development and `"1440x720"` for a production run.

    ```{r}
    pixels <- "360x180"
    ```

```{r}
pixels
```

    Next, we import both GeoTIFFs.
    
    ```{r, cache=TRUE}
    read_sstan <- function(year) {
      str_c(
        "AMSRE_SSTAn_M_",
        year,
        "-11-01_rgb_",
        pixels,
        ".FLOAT.TIFF"
      ) |>
        raster::raster() |>
        envirem::pacificCentric(crop = FALSE) |>
        st_as_stars() |>
        set_names(str_c("y", year))
    }
    sstan <- exec(c, !!!map(c(2002, 2007), read_sstan))
    ```
    
    
```{r}
str(sstan)
```
    
    YOUR TASK: MAKE A PLOT
    
```{r}
tm_shape(sstan) + 
  tm_raster()
```
  
```{r}

tm_shape(sstan, projection = "+proj=moll +lon_0=180", raster.warp = FALSE) + 
  tm_raster()
```
    
```{r}
sstan <- sstan |>
  mutate(temp2002 = if_else(y2002 < 99998, y2002, NA_real_),
         temp2007 = if_else(y2007 < 99998, y2007, NA_real_),
         surface_2002 = case_when(
           y2002 == 99999 ~"Land",
           y2002 == 99998 ~"sea Ice",
           TRUE ~NA_character_ 
         ),
         surface_2007 = case_when(
           y2007 == 99999 ~"Land",
           y2007 == 99998 ~"sea Ice",
           TRUE ~NA_character_ )
  )
```
    
```{r}
p <- tm_shape(tm_raster(col = "temp2002"))
```
    
```{r}
tm_shape(sstan, projection = "+proj=moll +lon_0=180", raster.warp = FALSE) + 
  tm_raster(col = "temp2002") + 
  tm_shape(sstan, projection = "+proj=moll +lon_0=180", raster.warp = FALSE) +
  tm_raster(col = "surface_2002")
```
```{r}
?tm_raster
```

  