---
title: "Lesson 20"
author: "Team Red"
date: '2022-04-05'
output: 
  prettydoc::html_pretty:
  theme: cayman
  highlight: github
  toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```

I need the following packages for this exercise.

```{r, warning = FALSE}
library(countrycode)
library(readxl)
library(sf)
library(tidyverse)
library(tmap)
library(rmapshaper)
```

(1) Code from the exercise about relational data.

    ```{r}
    gdp_per_cap <- read_excel(
      "API_NY.GDP.PCAP.KD_DS2_en_excel_v2_3731742.xls",
      skip = 3
    )
    life_exp <- read_excel(
      "API_SP.DYN.LE00.IN_DS2_en_excel_v2_3731513.xls",
      skip = 3
    )

    # Next function can pivot any of the World Bank tibbles
    pivot_longer_wb <- function(tb, name) {
      tb |>
        select(-starts_with("Indicator")) |>
        pivot_longer(
          matches("^\\d{4}$"), # String with exactly 4 digits
          names_to = "year",
          values_to = {{ name }},
          names_transform = list(year = as.integer)
        )
    }
    wb <-
      Hmisc::llist(life_exp, gdp_per_cap) |>
      imap(pivot_longer_wb) |>
      reduce(left_join, by = c("Country Name", "Country Code", "year")) |>
      semi_join(codelist, by = c("Country Code" = "iso3c"))
    ```
    
(2) Import the country borders as `sf` objects. 

```{r}
import_and_simplify <- function(zip_file) {
  zip_file |>
    unzip() |>
    str_subset(".shp$") |>
    read_sf() |>
    rmapshaper::ms_simplify()
  }

country_borders <- import_and_simplify("wb_countries_admin0_10m.zip")
```






(3) Norway and France have invalid country codes in the column `ISO_A3`. Chnage them to NOR and FRA, respectively. 
```{r}
country_borders <- country_borders |>
  mutate(ISO_A3 = case_when(
    FORMAL_EN == "French Republic" ~ "FRA",
    FORMAL_EN == "Kingdom of Norway" ~ "NOR",
    TRUE ~ ISO_A3
  )
)
```

(4) Create an `sf` object called countries that contains the columns:

- name
- code 
- gdp_per_capita
- life_Exp
- geometry

```{r}
countries <- country_borders |>
  inner_join(wb, by = c("ISO_A3" = "Country Code")) |>
  filter(year == 2015) |>
  select(
    name = "Country Name",
    code = "ISO_A3",
    gdp_per_cap,
    life_exp,
    geometry
  )

all(st_is_empty(countries) == FALSE)
# no geometry is empty
```

##### 5. Are any countries or land amsses missing? 

```{r}
data(World)
tm_shape(World) + 
  tm_polygons(col = "red")+
  tm_shape(countries) +
  tm_polygons() 
```

Some regions/land masses are missing, such as Western Saharan, Antarctica, and Kerguelen Islands. Missing regions are shown in red on the map 


##### 6. Add the missing polygons 
```{r}
land_mass <- import_and_simplify("wb_land_10m.zip")

tm_shape(land_mass) + 
  tm_polygons("grey")+
  tm_shape(countries) +
  tm_polygons("grey") 
```

##### 7. What is the current projection of the map? Is it an equal-area projection? 

```{r}
st_crs(countries)$proj4string
```

longitude/latitude is an equidistant projection.


##### 8. 


```{r}
# https://cran.r-project.org/web/packages/tmap/tmap.pdf

tm_shape(land_mass, projection = "+proj=eck4") +
  tm_polygons() +
  tm_shape(countries, projection = "+proj=eck4") +
  tm_polygons(
    col = c("gdp_per_cap", "life_exp"),
    title =
      c(
        "GDP per capita \n(US$, PPP 2015)",
        "Life Expectancy at birth \n(years)"
      ),
    palette = list("OrRd", "BuGn"),
    breaks = list(
      c(
        -Inf,
        1000,
        2000,
        5000,
        10000,
        20000,
        50000,
        Inf
      ),
      c(
        -Inf,
        60,
        65,
        70,
        75,
        80,
        85,
        Inf
      )
    ),
    labels = list(
      c(
        "Less than 1,000",
        "1,000 to 2,000",
        "2,000 to 5,000",
        "5,000 to 10,000",
        "10,000 to 20,000",
        "20,000 to 50,000",
        "50,000 or more"
      ),
      c(
        "Less than 60",
        "60 to 65",
        "65 to 70",
        "70 to 75",
        "75 to 80",
        "80 to 85",
        "85 or more"
      )
    )
  ) +
  tm_layout(
    main.title = "Wealth and Health by Country in 2015",
    main.title.size = 1.5,
    legend.title.size = 1,
    legend.text.size = 0.8,
    legend.position = c("left", "top"),
    inner.margins = c(0.01, 0.3, 0.01, 0.01),
    bg.color = "lightblue",
    earth.boundary = TRUE,
    space.color = "white",
    frame = FALSE
  ) +
  tm_text(
    "code",
    size = "AREA",
    root = 5,
    remove.overlap = TRUE,
    size.lim = c(0.3, 0.5)
  ) +
  tm_credits(
    text = c("", "Source: World Bank"),
    size = 0.9,
    position = c(0.01, 0.02)
  )
```



