---
title: "plotly writing"
author: "shuna"
date: "4/8/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(plotly)
```


# 3D plots 

Interactivity comes in really useful when you have more variables to plot. Adding the third axis is very straightforward, you may simply add a `z` attribute to `plot_ly()`. In the plot below, you can clearly see three clusters of iris species. 

```{r}
p <- iris |>
  plot_ly(
    x = ~Sepal.Length,
    y = ~Petal.Length,
    z = ~Petal.Width,
    marker = list(size = 5)
  ) |>
  add_markers(color = ~Species)

p
```

**Plotly** enables us to rotate, zoom in/out, and check the coordinates of datapoints. This is something two dimensional non-interactive plots cannot do, and thus **Ploly** stands as one of the best options when it comes to three dimensional plots. 

For other types of 3d plots, you may refer to the Plotly Graphing library[^10]
and look for the appropriate trace functions. 

[^10]: <https://plotly.com/python/3d-charts/> Accessed on 8/4/2022

### Axes 

In 3d plots, axes are embedded in the `scene` object, but still inside the `layout` function. So directly putting `xaxis = list(title = "x")` does not yield the desired output. Below is an example of how to use the `scene` object. Note that the title and legend are outside of the `scene` object. 

```{r}
p |>
  layout(
    # main title 
    title = "Iris Species",
    # legend 
    legend = list(
      title = list(text = "<b> Species <b>"),
      itemsizing = "constant"
    ),
    # add titles for three axes
    scene = list(
      xaxis = list(title = "Sepal Length (cm)"),
      yaxis = list(title = "Petal Length (cm)"),
      zaxis = list(title = "Petal Width (cm)")
      )
  )
```


