---
title: "mock exercise"
author: "shuna"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```

```{r, include = FALSE}
library(tidyverse)
library(plotly)
lq <- read_csv("life_quality.csv")
```

Your goal is to create this plot. 
```{r, echo = FALSE, fig.width = 6, fig.height=4 , fig.align='center'}
colors <- c("#e41a1c", '#377eb8', '#4daf4a', '#984ea3', '#ff7f00')

lq |>
  plot_ly(
    x = ~gdp_per_capita,
    y = ~life_expectancy,
    mode = "markers",
    type = "scatter",
    color = ~continent,
    colors = colors,
    fill = ~"",
    marker = list(
      size = ~pop,
      sizemode = "diameter",
      opacity = 0.6,
      sizes = c(10, 50),
      line = list(
        width = 1,
        color = "white"
      )
    ),
    hoverinfo = "text",
    text = ~ paste(
      "Country:", country_name,
      "<br>GDP per capita:", paste0("$", round(gdp_per_capita, digit = 2)),
      "<br>Life Expectancy:", paste(round(life_expectancy, digits = 2), "years"),
      "<br>Population:", pop
    )
  ) |>
  layout(
    title = "Wealth and Health by Country in 2015",
    legend = list(
      title = list(text = "<b> Continent <b>"),
      itemsizing = "constant"
    ),
    xaxis = list(
      title = "GDP per capita (USD, PPP2015)",
      type = "log",
      showgrid = TRUE,
      tickmode = "linear"
    ),
    yaxis = list(
      title = "Life Expectancy (years)"
    ),
    annotations = list(
      x = 1.2, y = -0.1,
      text = "Source: World Bank",
      showarrow = F, xref = "paper", yref = "paper",
      xanchor = "right", yanchor = "auto", xshift = 0, yshift = 0,
      font = list(size = 12)
    ) 
  ) 
```
(0) Download https://michaelgastner.com/DAVisR_data/life_quality.csv.

(1) plot a scatterplot 

- x-axis: gdp_per_capita
- y-axis: life_expectancy 
- mode: markers


(2) set colors by continents

- make sure you have two separate arguments
- one to specify which column to use for coloring 
- the other to specify the colors 
- You may use this pallete for colouring continents. 
    ```{r}
    colors <- c("#e41a1c", '#377eb8', '#4daf4a', '#984ea3', '#ff7f00')
    ```

(3) set size 

- size by population 
- The default for the parameter `fill` is `None` and it seems to cause an error. 
- You will see the following warning "Warning: `line.width` does not currently support multiple values". 
- you may consider adding the following argument in plot_ly(); this code assigns an empty string to `fill`. 
    ```{r, eval=FALSE}
    fill = ~'',
    ```

- Unfortunately, as of today, plotly does not support the size legend. However, some people seem to have found a way around by creating a new plot just for legend and combine it with the original bubble chart using `subplot` function. You can try it once you are done with the rest of exercises. 

(4) assign texts to each point 

- show country name, gdp, life expectancy, and population 
- remember to put line breaks
- for gdp and life expectancy, round the numbers to 2 digits 

(5) add labels 

- add an appropriate title 
- add an appropriate legend 
- add an appropriate x and y axis 
- for adding the source, you may consider using the code below: 
    ```{r, eval = FALSE}
    annotations = list(
      x = 1.2, y = -0.1,
      text = "Source: World Bank",
      showarrow = F, xref = "paper", yref = "paper",
      xanchor = "right", yanchor = "auto", xshift = 0, yshift = 0,
      font = list(size = 12)
    )
    ```



(6) fine-tune, complete the plot 

- make sure the x-axis is log-transformed 
- make sure the sizes of the symbol in the legend are constant.
- improve the plot, if you would like. 




Thank you! 




