---
title: "plotly Exercise"
author: "Team Red"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```

```{r, include = FALSE}
library(tidyverse)
library(plotly)
lq <- read_csv("life_quality.csv")
```

Your goal is to create this plot using **plotly**. 
```{r, echo = FALSE, fig.width = 6, fig.height=4 , fig.align='center'}
colors <- c("#e41a1c", '#377eb8', '#4daf4a', '#984ea3', '#ff7f00')

lq |>
  plot_ly(
    x = ~gdp_per_capita,
    y = ~life_expectancy,
    mode = "markers",
    type = "scatter",
    color = ~continent,
    colors = colors,
    fill = ~"",
    size = ~pop,
    marker = list(
      sizemode = "diameter",
      opacity = 0.6,
      sizes = c(10, 50), #what does this do?
      line = list(
        width = 1,
        color = "white"
      )
    ),
    hoverinfo = "text",
    text = ~ paste(
      "Country:", country_name,
      "<br>GDP per capita:", paste0("$", round(gdp_per_capita, digit = 2)),
      "<br>Life Expectancy:", paste(round(life_expectancy, digits = 2), "years"),
      "<br>Population:", pop
    )
  ) |>
  layout(
    title = "Wealth and Health by Country in 2015",
    legend = list(
      title = list(text = "<b> Continent <b>"),
      itemsizing = "constant"
    ),
    xaxis = list(
      title = "GDP per capita (USD, PPP2015)",
      type = "log",
      showgrid = TRUE,
      tickmode = "linear"
    ),
    yaxis = list(
      title = "Life Expectancy (years)"
    ),
    annotations = list(
      x = 1.2, 
      y = -0.1,
      text = "Source: World Bank",
      showarrow = F, 
      xref = "paper", 
      yref = "paper",
      xanchor = "right", 
      yanchor = "auto", 
      xshift = 0, 
      yshift = 0,
      font = list(size = 12)
    ) 
  ) 
```
(0) Download https://michaelgastner.com/DAVisR_data/life_quality.csv.

(1) Make a scatterplot where: 

- The x-coordinate is the GDP per capita.
- The y-coordinate is the life expectancy. 
- `mode = "markers"`


(2) Set the colors of each point to indicate which continent it belongs to.

- Make sure you have two separate parameters `color` and `colors`:
    - `color` should specify which column or variable to map to the color of each point.
    - `colors` should specify which colors or palette to use.
- You may use this palette for coloring continents: 
    ```{r}
    colors <- c("#e41a1c", '#377eb8', '#4daf4a', '#984ea3', '#ff7f00')
    ```

(3) Set the size of each bubble to indicate the population.
- You will see the following warning `Warning: line.width does not currently support multiple values`.
- You may consider using `fill = ~''` in plot_ly(); this code assigns an empty string to `fill`. 

- Unfortunately, as of today, **plotly** does not support the size legend. However, some people seem to have found a way around by creating a new plot just for legend and combine it with the original bubble chart using `subplot` function. You can try it once you are done with the rest of exercises. 

(4) Assign texts to each point that will show when you hover over it.
- The text should contain the country name, GDP per capita, life expectancy, and population.
- Remember to add line breaks.
- For GDP per capita and life expectancy, round the numbers to 2 digits.
- Place a dollar sign in front of GDP per capita.

(5) Add labels to the plot using `layout()`
- Change the axis labels and give the plot a title. 
- Add an appropriate legend.
- Give credit to the World Bank as a source. To accomplish this, you may consider using the code below: 
    ```{r, eval = FALSE}
    annotations = list(
      x = 1.2, 
      y = -0.1,
      text = "Source: World Bank",
      showarrow = F, 
      xref = "paper", 
      yref = "paper",
      xanchor = "right", 
      yanchor = "auto", 
      xshift = 0, 
      yshift = 0,
      font = list(size = 12)
    ) 
    ```

(6) Fine-tune and complete the plot. 
- Make sure the x-axis is log-transformed.
- To transform into a logarithmic axis, you should use this code:

    ```{r, eval = FALSE}
    xaxis = list(
      type = "log", 
      showgrid = TRUE, 
      tickmode = "linear"
    )
    ```

- Make sure the sizes of the symbol in the legend are constant.
- Add a thin white border around the points to make it easier to distinguish overlapping points. 
- You should consider using the following code in `marker`. 
    ```{r}
    line = list(
            width = 1,
            color = "white"
    )
    ```

- Feel free to make more adjustments if you think they will improve the quality of the plot. 


Thank you! 




