---
title: "Plotly Tutorial"
author: "Team Red"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE,
                      fig.width = 5, fig.height = 4)
```



Packages needed for this tutorial: 
```{r, warning = FALSE}
library(tidyverse)
library(plotly)
```

## Introduction 
The R package, **plotly** enables us to creative beautiful graphs with interactive graphics. The interactivity aspect is useful in instances where the data set we are dealing with includes a lot of text information or when there is not enough space to plot everything (i.e., data looks really small and compressed). By including interactivity on the graph, users can access information easier, for example, hovering over a point in a bubble chart to see the exact values of the coordinates and name of which country, school, etc., the point represents. 

There are also lots of parallels between **plotly** and ggplot in the sense that most of the plots that can be created with ggplot can also be created with **plotly**, adding interactivity as a unique feature. 

The function used to plot graphs in **plotly** is `plot_ly()`. The first input to the function has to be the data set you are working with. Then, you can allocate variable names to "visual properties (e.g., `x`, `y`, `color`, etc)". We will look at how to plot different types of plots using this function shortly[^1]. 

[^1]: <https://plotly-r.com/overview.html> Accessed on 7/4/2022

## Layout 
The `layout()` function is an essential part in plotting graphs in **plotly** as it will be commonly used. The argument, `layout()` can add and modify different parts of the layout of your graph, including, setting the title, axis labels, axis range, plot size, etc. 

```{r}
layout(
  plot_ly(mtcars, x = ~mpg, y = ~hp),
  title = "Miles/(US) gallon and horsepower"
)
```


However, when we start to plot graphs that are more complex, it may be harder to read the code in the above example. Hence, we can use the pipe function `|>` to make it easier to read and create the same graph, just insert the title into the `layout()` function as it is the only addition to the layout of our graph[^1]. 

```{r}
mtcars |>
  plot_ly(x = ~mpg, y = ~hp) |>
  layout(
    title = "Miles/(US) gallon and horsepower",
    autosize = FALSE, width = 600, height = 500
  )
```

## Basic Charts 

### Scatterplots 

As seen in the previous section, `plot_ly()` will plot the data as a scatter plot if the data you input for both variables are both quantitative data. Just inputting the data and variables will you give a basic scatter plot (as seen under the section, 'Layout'). 

Another way to plot a scatter plot is by using `add_markers()`, which is for indicating what kind of plot you want to produce[^2]. However, this may not always be needed because `plot_ly()` will plot a scatter plot from two quantitative variables without any specification. 

[^2]: <https://plotly-r.com/scatter-traces.html#markers> Accessed on 7/4/2022

```{r}
mtcars |>
  plot_ly(x = ~mpg, y = ~hp) |>
  add_markers() |>
  layout(title = "Miles/(US) gallon and horsepower")
```

To customize your scatter plot even further, you'll have to add the argument `marker = `[^3]. 

[^3]: <https://plotly.com/r/line-and-scatter/> Accessed on 7/4/2022. 

```{r}
mtcars |>
  plot_ly(
    x = ~mpg, y = ~hp,
    marker = list(
      size = 10,
      color = "red",
      opacity = 0.5
    )
  ) |> # Note that plotly uses 'opacity' instead of 'alpha' to adjust the opacity of points!
  layout(title = "Miles/(US) gallon and horsepower") 
```

### Line Charts

Line charts also requires two variables of quantitative data for input. In order to change it to a line chart, we have to add the argument `add_lines()` to the plot. 

```{r}
mtcars |>
  plot_ly(x = ~mpg, y = ~hp) |>
  add_lines() |>
  layout(title = "Miles/(US) gallon and horsepower")
```

### Bubble Chart 

Bubble charts in **plotly** are based off the same arguments for creating a scatter plot. The only distinguishing detail is that we have to assign a variable to the argument `size = ` within the `marker = ` list[^4].

[^4]: <https://plotly.com/r/bubble-charts/> Accessed on 7/4/2022. 

```{r}
mtcars |>
  plot_ly(
    x = ~mpg, y = ~hp,
    marker = list(size = ~wt, opacity = 0.5, color = "red")
  ) |>
  add_markers() |>
  layout(title = "Miles/(US) gallon and horsepower")
```

## `ggplotly()` 

To get a better understanding of how **plotly** works, we can first compare it with **ggplot**. Let us first construct a **ggplot** object. In this example, we make use of the `mtcars` data and form a scatterplot, plotting `mpg` on the x-axis and `wt` on the y-axis.
```{r}
mtcars <- mtcars
ggplot_example <- ggplot(mtcars, aes(x = mpg, y = wt)) +
  geom_point()

ggplot_example
```

We can convert `ggplot_example` into a **plotly** object by passing it as an argument in `ggplotly()`. When we hover over the different points, we see the corresponding `mpg` and `wt` value for each point. By setting the argument of `tooltip` to `"all"`, all the aesthetic mappings of a point can be seen when we hover the cursor over the point. 
```{r}
ggplotly(p = ggplot_example, tooltip = "all")
```

The main difference between `ggplotly()` and `plot_ly()` is that we can use **ggplot** functions when creating the plot that we pass to `ggplotly()` as it simply converts **ggplot** objects into **plotly** objects. On the other hand, if we use `plot_ly()` in creating our plot, we can only use  **plotly** functions in doing so. 

## Adding and Updating Traces
A trace is a set of data with specifications for their locations on the plot. We can always add and update **plotly** objects by using `add_trace()`. With this method, we can add new features to our plots. For instance, we can add a trace to `example_2`.
```{r}
plot_ly(mtcars, x = ~mpg, y = ~wt, type = "scatter", mode = "markers") |>
  add_trace(
    x = c(10, 15, 20), y = c(2, 3, 4), type = "scatter",
    mode = "markers"
  )
```

We can also start with an empty figure then add traces to this figure. For instance, we can add three bars to an empty figure.
```{r}
plot_ly() |>
  add_trace(x = c(1, 2, 3), y = c(5, 10, 3), type = "bar")
```


## Subplots with Plotly 

Using the function `subplot()` allows you to style mutliple plots and display them
in a singular plotting space. 

`Subplot()` possesses several advantages over the `par(mfrow)` parameter: 

- Setting `par(mfrow)` requires you to reset the parameter again after the desired
 subplots have been arrange, while `subplpot()`does not.
- `par(mfrow)` does not allow you to create plots of different sizes, while
 `subplot()` offers the option for you to create multiple custom-sized subplots.
 
### Creating simple subplots 

Using the `mtcars` dataset again, let's stack two plots on each other. Here, both plots represent the relationship between Miles/(US) gallon and horsepower as a scatterplot and line graph. 
 
```{r}
mt_scatter <- mtcars |>
  plot_ly(x = ~mpg, y = ~hp) |>
  add_markers() |>
  layout(title = "Miles/(US) gallon and horsepower")

mt_line <- mtcars |>
  plot_ly(x = ~mpg, y = ~hp) |>
  add_lines() |>
  layout(title = "Miles/(US) gallon and horsepower")

# Specify the name of the plots to display together 
subplot(
  mt_scatter, 
  mt_line,
  # use the nrows argument to arrange the subplots 
  nrows = 2,
  # When TRUE, states that both graphs should share X-axis labels
  shareX = TRUE, 
  margin = 0.05
)
```

Any subsequent plots placed after the above `subplot()` chunk will not be affected by 
`subplot()`.

- The argument `nrows` allows us to specify the number of rows the plots will be arranged
accordingly to. If we would like to place the two plots side-by-side instead of stacking them,
we can specify `nrows = 1` instead.
- As we have two plots containing the same set of information, the x-axis can be standardized
and displayed on the bottom graph. We set `shareX = TRUE` for this. 
You may search `?subplot` in the console for the full documentation guide of arguments. 
 
### Using subplot() with add_trace()
 
`Subplot()` also works with plots with traces created using the `add_trace()`
function. 
 
```{r}
plot_1 <- plot_ly() |>
  add_trace(x = c(1, 2, 3), y = c(5, 10, 3), type = "bar") 

plot_2 <- plot_ly() |>
  add_trace(x = c(3, 2, 1), y = c(10, 3, 5), type = "bar") 

subplot(plot_1, plot_2, 
  nrows = 1,
  # creates a small margin between the two subplots
  margin = 0.05
) |> 
   layout(
     # this provides the main overarching title shared by the two plots 
     title = "two example plots", 
     # this allows us to add a list of annotations,in this case the titles for the subplots 
     annotations = list( 
    list(x = 0.15 , y = 1, text = "Plot 1", showarrow = F, xref='paper', yref='paper'), 
    list(x = 0.85 , y = 1, text = "Plot 2", showarrow = F, xref='paper', yref='paper')
    ) 
  ) 
```

One issue to note is that using `subplot()` with subplots that each contain a title
often results in some of the subplot titles going missing in the final plotting space.
You can see this issue with the `mtcars` Miles/(US) gallon and horsepower figure we 
constructed earlier. 

A suggested remedy is included in the above graph.[^1]

[^1]: <https://plotly.com/r/subplots/> Accessed on 7/4/2022

It uses the `annotation` argument under `layout()` to manually place annotations using x and y coordinates.

- `xref` and `yref` are crucial arguments that set the container `x` and `y` refer to, where "container" spans the entire `width` of the plot. "paper" refers to the width of the plotting area only.[^2]

[^2]: <https://plotly.com/r/reference/#Layout_and_layout_style_object> Accessed on 7/4/2022 

### Creating basic custom-sized suplots 

One powerful feature of `subplot()` is the ability to create custom-sized subplots. 
This can be useful in informing viewers on which particular subplots act to substantiate 
more crucial subplots. 

Here is an example below. We want to create 2 smaller stacked subplots of equal combined height to that of a larger subplot on the right. To do this, we create multiple vectors that will contain our subplots.

Here are the subplots we will use:

- The same Miles/(US) gallon and horsepower scatterplot and line graph as the two smaller subplots
(which we have titled "mt_scatter" and "mt_line" from the `mt_cars` dataset).
- The add_trace plot we created titled "plot_1" as the larger plot on the right. 


```{r}
s1 <- subplot(mt_scatter, mt_line, nrows = 2, shareX = TRUE)
s2 <- subplot(s1, plot_1, margin = 0.05) |>
  # remove the title that automatically takes the title from one of the subplots
  layout(title = NA)
s2
```

In summary, creating basic custom-sized subplots can be done by placing "smaller" subplots
within a vector that nests itself within another vector containing the "larger" subplot. 

## 3D plots 

### Scatterplots
Interactivity comes in really useful when you have more variables to plot. Adding the third axis is very straightforward, you may simply add a `z` attribute to `plot_ly()`. In the plot below, you can clearly see three clusters of iris species. 

```{r}
p <- iris |>
  plot_ly(
    x = ~Sepal.Length,
    y = ~Petal.Length,
    z = ~Petal.Width,
    marker = list(size = 5)
  ) |>
  add_markers(color = ~Species)

p
```

**Plotly** enables us to rotate, zoom in/out, and check the coordinates of datapoints. This is something two dimensional non-interactive plots cannot do, and thus **Ploly** stands as one of the best options when it comes to three dimensional plots. 

For other types of 3d plots, you may refer to the Plotly Graphing library[^10]
and look for the appropriate trace functions. 

[^10]: <https://plotly.com/python/3d-charts/> Accessed on 8/4/2022

### Axes 

In 3d plots, axes are embedded in the `scene` object, but still inside the `layout`. So directly putting `xaxis = list(title = "x")` does not yield the desired output. Below is an example of how to use the `scene` object. Note that the title and legend are outside of the `scene` object. 

```{r}
p |>
  layout(
    # main title 
    title = "Iris Species",
    # legend 
    legend = list(
      title = list(text = "<b> Species <b>"),
      itemsizing = "constant"
    ),
    # add titles for three axes
    scene = list(
      xaxis = list(title = "Sepal Length (cm)"),
      yaxis = list(title = "Petal Length (cm)"),
      zaxis = list(title = "Petal Width (cm)")
      )
  )
```
