---
title: "plotly using subplot function"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load the pacakges: 
```{r, message = FALSE}
library(tidyverse)
library(plotly)
```

# Subplots with Plotly 

Using the function `subplot()` allows you to style mutliple plots and display them
in a singular plotting space. 

`Subplot()` possesses several advantages over the `par(mfrow)` parameter: 

- Setting `par(mfrow)` requires you to reset the parameter again after the desired
 subplots have been arrange, while `subplpot()`does not.
- `par(mfrow)` does not allow you to create plots of different sizes, while
 `subplot()` offers the option for you to create multiple custom-sized subplots.
 
### Creating simple subplots 

Using the `mtcars` dataset again, let's stack two plots on each other. Here, both plots represent the relationship between Miles/(US) gallon and horsepower as a scatterplot and line graph. 
 
```{r}
mt_scatter <- mtcars |>
  plot_ly(x = ~mpg, y = ~hp) |>
  add_markers() |>
  layout(title = "Miles/(US) gallon and horsepower")

mt_line <- mtcars |>
  plot_ly(x = ~mpg, y = ~hp) |>
  add_lines() |>
  layout(title = "Miles/(US) gallon and horsepower")

# Specify the name of the plots to display together 
subplot(
  mt_scatter, 
  mt_line,
  # use the nrows argument to arrange the subplots 
  nrows = 2,
  # When TRUE, states that both graphs should share X-axis labels
  shareX = TRUE, 
  margin = 0.05
)
```

Any subsequent plots placed after the above `subplot()` chunk will not be affected by 
`subplot()`.

- The argument `nrows` allows us to specify the number of rows the plots will be arranged
accordingly to. If we would like to place the two plots side-by-side instead of stacking them,
we can specify `nrows = 1` instead.
- As we have two plots containing the same set of information, the x-axis can be standardized
and displayed on the bottom graph. We set `shareX = TRUE` for this. 
You may search `?subplot` in the console for the full documentation guide of arguments. 
 
### Using subplot() with add_trace()
 
`Subplot()` also works with plots with traces created using the `add_trace()`
function. 
 
```{r}
plot_1 <- plot_ly() |>
  add_trace(x = c(1, 2, 3), y = c(5, 10, 3), type = "bar") 

plot_2 <- plot_ly() |>
  add_trace(x = c(3, 2, 1), y = c(10, 3, 5), type = "bar") 

subplot(plot_1, plot_2, 
  nrows = 1,
  # creates a small margin between the two subplots
  margin = 0.05
) |> 
   layout(
     # this provides the main overarching title shared by the two plots 
     title = "two example plots", 
     # this allows us to add a list of annotations,in this case the titles for the subplots 
     annotations = list( 
    list(x = 0.15 , y = 1, text = "Plot 1", showarrow = F, xref='paper', yref='paper'), 
    list(x = 0.85 , y = 1, text = "Plot 2", showarrow = F, xref='paper', yref='paper')
    ) 
  ) 
```

One issue to note is that using `subplot()` with subplots that each contain a title
often results in some of the subplot titles going missing in the final plotting space.
You can see this issue with the `mtcars` Miles/(US) gallon and horsepower figure we 
constructed earlier. 

A suggested remedy is included in the above graph.[^1]

[^1]: <https://plotly.com/r/subplots/> Accessed on 7/4/2022

It uses the `annotation` argument under `layout()` to manually place annotations using x and y coordinates.

- `xref` and `yref` are crucial arguments that set the container `x` and `y` refer to, where "container" spans the entire `width` of the plot. "paper" refers to the width of the plotting area only.[^2]

[^2]: <https://plotly.com/r/reference/#Layout_and_layout_style_object> Accessed on 7/4/2022 

### Creating basic custom-sized suplots 

One powerful feature of `subplot()` is the ability to create custom-sized subplots. 
This can be useful in informing viewers on which particular subplots act to substantiate 
more crucial subplots. 

Here is an example below. We want to create 2 smaller stacked subplots of equal combined height to that of a larger subplot on the right. To do this, we create multiple vectors that will contain our subplots.

Here are the subplots we will use:

- The same Miles/(US) gallon and horsepower scatterplot and line graph as the two smaller subplots
(which we have titled "mt_scatter" and "mt_line" from the `mt_cars` dataset).
- The add_trace plot we created titled "plot_1" as the larger plot on the right. 


```{r}
s1 <- subplot(mt_scatter, mt_line, nrows = 2, shareX = TRUE)
s2 <- subplot(s1, plot_1, margin = 0.05) |>
  # remove the title that automatically takes the title from one of the subplots
  layout(title = NA)
s2
```

In summary, creating basic custom-sized subplots can be done by placing "smaller" subplots
within a vector that nests itself within another vector containing the "larger" subplot. 


```{r}
library(plotly)

data <- read.csv("https://raw.githubusercontent.com/plotly/datasets/master/school_earnings.csv")

fig <- plot_ly(data, x = ~Women, y = ~Men, text = ~School, type = 'scatter', mode = 'markers',
        marker = list(size = ~Gap, opacity = 0.5))
fig <- fig %>% layout(title = 'Gender Gap in Earnings per University',
         xaxis = list(showgrid = FALSE),
         yaxis = list(showgrid = FALSE))

fig

```

 