---
title: "test_analysis"
author: "shuna"
date: "2023-03-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readr)
library(hms)
library(nlme)
library(nortest)
library(lme4)
library(DHARMa)
library(MASS)
```

do we treat distance as a factor, or continuous vairable? 

## function

```{r}
public_holidays <- as.Date(c("2022-12-31", "2023-01-01","2023-01-02", "2023-01-03", "2023-01-27", "2023-01-28", "2023-01-29", "2023-01-30", "2023-01-31", "2023-02-10", "2023-02-13", "2023-03-01"))

prep_data <- function(df){
  df |>
    mutate(
      transect = as.double(substr(FOLDER, 2,2)),
      distance = case_when(
        substr(FOLDER, 4, 4) == "5" ~ 50,
        substr(FOLDER, 4, 4) == "1" ~ 100,
        TRUE ~ 200
        ),
      time_sec = as.numeric(TIME) + OFFSET,
      time_day = as_hms(time_sec),
      to_classify = weekdays(DATE),
      Date = as.Date(DATE, format = "%m/%d/%Y"),
      days = case_when(
        Date %in% public_holidays ~ "off_day",
        to_classify == "Saturday" ~ "off_day",
        to_classify == "Sunday" ~ "off_day",
        TRUE ~ "weekday")
    ) |>
    dplyr::select(
      DATE, top_match = `TOP1MATCH*`, top_dist = TOP1DIST,
      transect, distance, time_day, days, DURATION
    )
}
#RJ <- prep_data(read_csv("gallus_clusters.csv"))
BB <- prep_data(read_csv("data/BB_clusters.csv"))
#DR <- prep_data(read_csv("drongo_clusters.csv"))
```

# BB

```{r}
BB_analysis <- BB |>
  ## filter for the top_match
  filter(top_match == "Straw-headed Bulbul") |>
  ## filter for the distance from the cluster 
  filter(top_dist <= 0.917)
```

## duration analysis
```{r}
BB_analysis <- BB |>
  mutate(dist = as.factor(distance))

BB_duration_model <- lme(DURATION ~ distance * days, random = ~1|transect, data = BB_analysis)
BB_duration_model <- lme(DURATION ~ dist * days, random = ~1|transect, data = BB_analysis)
summary(BB_duration_model)
## diagnostic plot

### variance for the dataset 
plot(BB_duration_model, resid(.,scaled = T) ~ fitted(.) | transect, abline = 0, lty = 2)
plot(BB_duration_model)
# a pattern of residuals decreasing as fitted values increase is observed

### normality 
qqnorm(BB_duration_model, abline = c(0, 1), lty = 2)
qqnorm(BB_duration_model, ~resid(., type = "p") | transect , abline = c(0,1) , lty = 2)
# distribution is not normal

### large enough sample size always reject the null hypothesis? 
ad.test(BB_analysis$DURATION)
hist(BB_analysis$DURATION)
```

```{r}
summary(BB_duration_model)
BB_analysis |>
  ggplot(aes(x = days, y = DURATION, color = as.factor(distance))) +
  geom_boxplot()

BB_analysis |>
  ggplot(aes(x = days, y = DURATION)) +
  geom_boxplot() +
  coord_cartesian(ylim = c(2, 4))

BB_analysis |>
  ggplot(aes(x = distance, y = DURATION, color = days)) +
  geom_boxplot() +
  geom_smooth(method = "lm")
```

```{r}
BB_duration_model.1 <- lme(DURATION ~ distance + days, random = ~1|transect, data = BB_analysis)

bp <- 100
# 50 - 100
b1 <- function(x, bp) ifelse(x <= bp, x, 0)
b1(BB_analysis$distance, 100)
# 100 - 200
b2 <- function(x, bp) ifelse(x < bp, 0, x)
b2(BB_analysis$distance, 100)

# interaction
BB_duration_model.2 <- lmer(DURATION ~ b1(distance, 100) * days + b2(distance, 100) * days + (1|transect), data = BB_analysis)
summary(BB_duration_model.2)
plot(BB_duration_model.2)
# the interaction is changing the relationship between days and duration of call. the effect size is small
# base on p-values, shouldnt remove. base on AIC, should remove
# add in point about the effect size being small when intepreting the results
# go ahead with biological intuition
```

```{r}
BB_duration_model.2 <- glmer.nb(DURATION ~ b1(distance, 100) * days + b2(distance, 100) * days + (1|transect), data = BB_analysis)
summary(BB_duration_model.2)
```

```{r}
plot(BB_duration_model.2)

# parallel
BB_duration_model.3 <- lmer(DURATION ~ b1(distance, 100) + b2(distance, 100) + days + (1|transect), data = BB_analysis)
summary(BB_duration_model.3)
plot(BB_duration_model.3)

AIC(BB_duration_model.2, BB_duration_model.3)

### variance for the dataset 
plot(BB_duration_model.2, resid(.,scaled = T) ~ fitted(.) | transect, abline = 0, lty = 2)
plot(BB_duration_model.2)
# a pattern of residuals decreasing as fitted values increase is observed

### normality 
qqnorm(BB_duration_model.2, abline = c(0, 1), lty = 2)
qqnorm(BB_duration_model.2, ~resid(., type = "p") | transect , abline = c(0,1) , lty = 2)
# distribution is not normal
```

```{r}

BB_duration_model.4 <- glmmPQL(DURATION ~ distance * days, random = ~1|transect, data = BB_analysis, family = quasi)
summary(BB_duration_model.4)

BB_duration_model.5 <- glmmPQL(DURATION ~ distance + days, random = ~1|transect, data = BB_analysis, family = quasi(link = "identity", variance = "constant"))
summary(BB_duration_model.3)

# BB_duration_model.2<- lme(DURATION ~ distance, random = ~1|transect, data = BB_analysis)
# summary(BB_duration_model.2)

# duration decreases as distance increases
```


## dawn chorus 


https://link-springer-com.libproxy1.nus.edu.sg/chapter/10.1007/978-3-030-39200-0_3 

```{r}
BB_chorus <- BB_analysis |>
  filter(time_day <= as.hms("10:00:00", format = "%h%m:%s"))  |>
  group_by(transect, distance, days) |>
  summarise(n = n()) 
```

```{r}
# there isnt a clear cut off for relationship 
# split the data set to 50-100, and 200
BB_chorus_model.1 <- glmer(n ~ b1(distance, 100) * days + b2(distance, 100) * days + (1|transect), 
                         data = BB_chorus, family = poisson)
## rescale variables 
BB_chorus_model.2 <- glmer(n ~ scale(b1(distance, 100)) * days + scale(b2(distance, 100)) * days + (1|transect), data = BB_chorus, family = poisson)

plot(simulateResiduals(BB_chorus_model.2))

# no problems detected

BB_chorus_model.3 <- glmer(n ~ scale(distance) + days + (1|transect), 
                         data = BB_chorus, family = poisson)

plot(simulateResiduals(BB_chorus_model.3))

# parallel model - 
BB_chorus_model.4 <- glmer(n ~ scale(b1(distance, 100)) + scale(b2(distance, 100)) + days + (1|transect), 
                         data = BB_chorus, family = poisson)

plot(simulateResiduals(BB_chorus_model.4))

## problem with residuals 

# BB_chorus_model.3 <- glmer.nb(n ~ scale(distance) * days + (1|transect), data = BB_chorus)
# 
# plot(simulateResiduals(BB_chorus_model.3))
# 
summary(BB_chorus_model.2)
summary(BB_chorus_model.3)
summary(BB_chorus_model.4)
# As you increase the distance by one, the expected number of calls increases by a factor of exp(-0.15199) = 0.8589%.

# number of calls made decreases as distances increases
exp(0.49426)
# On weekdays, the number of calls is 1.64 greater than the number of calls on weekends
exp(3.1032-0.21103)

exp(3.1032-0.35943)

# we observe a decrease in the number of calls from distance 200 - 100m and distance 100m - 50m.
exp(3.1032)
```

# Drongo
  
```{r}
DR_analysis <- DR |>
  filter(top_match == "drongo") |>
  filter(top_dist <= 0.5529) 
```


## duration 

```{r}
DR_duration_model <- lme(DURATION ~ distance * days, random = ~1|transect, data = DR_analysis)

## diagnostic plot

### variance for the dataset 
plot(DR_duration_model, resid(.,scaled = T) ~ fitted(.) | transect, abline = 0, lty = 2)
plot(DR_duration_model)

### normality 
qqnorm(DR_duration_model, abline = c(0, 1), lty = 2)
qqnorm(DR_duration_model, ~resid(., type = "p") | transect , abline = c(0,1) , lty = 2)

### large enough sample size always reject the null hypothesis? 
ad.test(log(DR_analysis$DURATION))
hist(log(DR_analysis$DURATION))

DR_duration_model.2 <- lme(log(DURATION) ~ distance * days, random = ~1|transect, data = DR_analysis)
summary(DR_duration_model.2)

DR_duration_model.3 <- glmmPQL(DURATION ~ distance * days, random = ~1|transect, data = DR_analysis, family = quasi)
summary(DR_duration_model.3)

# 
# DR_duration_model.3 <- lme(log(DURATION) ~ distance + days, random = ~1|transect, data = DR_analysis)
# summary(DR_duration_model.3)
```

## dawn chorus 

```{r}
DR_chorus <- DR_analysis |>
  filter(time_day <= as.hms("10:00:00", format = "%h%m:%s"))  |>
  group_by(transect, distance, days) |>
  summarise(n = n()) 
```

```{r}
DR_chorus_model.1 <- glmer(n ~ distance * days + (1|transect), 
                         data = DR_chorus, family = poisson)
## rescale variables 
DR_chorus_model.2 <- glmer(n ~ scale(distance) * days + (1|transect), 
                         data = DR_chorus, family = poisson)

plot(simulateResiduals(DR_chorus_model.2))
## problem with residuals and outliers

DR_chorus_model.3 <- glmer.nb(n ~ scale(distance) * days + (1|transect), data = DR_chorus)

plot(simulateResiduals(DR_chorus_model.3))
# problems solved 
```

```{r}
summary(DR_chorus_model.3)
## interaction not significant 

DR_chorus_model.4 <-  glmer.nb(n ~ scale(distance) + days + (1|transect), data = DR_chorus)
summary(DR_chorus_model.4)
# weekday not significant

DR_chorus_model.5 <-  glmer.nb(n ~ scale(distance) + (1|transect), data = DR_chorus)
summary(DR_chorus_model.5)

# DR_chorus_model.5 <- glmmPQL(n ~ scale(distance), random = ~1|transect, 
#                              data = DR_chorus, family = quasipoisson)
# summary(DR_chorus_model.5)
```

# Summary 

## Junglefowl 
distance is significantly related to duration and dawn chorus 

## Asian Koel 
distance and type of day are significantly associated with the duration of the call, 
no association found for the dawn chorus. 

## Bulbul
distance and type of day are significantly related to both duration and dawn chorus

## Drongo
distance and type of day are significantly related to duration of the call.
distance is significantly related to dawn chorus

```{r}
library(segmented)

```

