---
title: "test_analysis"
author: "shuna"
date: "2023-03-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readr)
library(hms)
library(nlme)
library(nortest)
library(lme4)
library(DHARMa)
library(MASS)
```

do we treat distance as a factor, or continuous vairable? 

## function

```{r}
public_holidays <- as.Date(c("2022-12-31", "2023-01-01","2023-01-02", "2023-01-03", "2023-01-27", "2023-01-28", "2023-01-29", "2023-01-30", "2023-01-31", "2023-02-10", "2023-02-13", "2023-03-01"))

prep_data <- function(df){
  df |>
    mutate(
      transect = as.double(substr(FOLDER, 2,2)),
      distance = case_when(
        substr(FOLDER, 4, 4) == "5" ~ 50,
        substr(FOLDER, 4, 4) == "1" ~ 100,
        TRUE ~ 200
        ),
      time_sec = as.numeric(TIME) + OFFSET,
      time_day = as_hms(time_sec),
      to_classify = weekdays(DATE),
      Date = as.Date(DATE, format = "%m/%d/%Y"),
      days = case_when(
        Date %in% public_holidays ~ "off_day",
        to_classify == "Saturday" ~ "off_day",
        to_classify == "Sunday" ~ "off_day",
        TRUE ~ "weekday")
    ) |>
    dplyr::select(
      DATE, top_match = `TOP1MATCH*`, top_dist = TOP1DIST,
      transect, distance, time_day, days, DURATION
    )
}

RJ <- prep_data(read_csv("data/gallus_clusters.csv"))
AK <- prep_data(read_csv("data/koel_clusters.csv"))
```

# Red junglefowl

```{r}
RJ_analysis <- RJ |>
  ## filter for the top_match
  filter(top_match == "chicken") |>
  ## filter for the disrtance from the cluster 
  filter(top_dist <= 0.8646)
```

## duration analysis
```{r}
RJ_duration_model <- lme(DURATION ~ distance * days, random = ~1|transect, data = RJ_analysis)

## diagnostic plot

### variance for the dataset 
plot(RJ_duration_model, resid(.,scaled = T) ~ fitted(.) | transect, abline = 0, lty = 2)
plot(RJ_duration_model)

### normality 
qqnorm(RJ_duration_model, abline = c(0, 1), lty = 2)
qqnorm(RJ_duration_model, ~resid(., type = "p") | transect , abline = c(0,1) , lty = 2)

### large enough sample size always reject the null hypothesis? 
ad.test(RJ_analysis$DURATION)
hist(RJ_analysis$DURATION)
```

```{r}
summary(RJ_duration_model)
```
```{r}
RJ_duration_model.1 <- lme(DURATION ~ distance + days, random = ~1|transect, data = RJ_analysis)
summary(RJ_duration_model.1)
RJ_duration_model.2<- lme(DURATION ~ distance, random = ~1|transect, data = RJ_analysis)
summary(RJ_duration_model.2)
```


## dawn chorus 


https://link-springer-com.libproxy1.nus.edu.sg/chapter/10.1007/978-3-030-39200-0_3 

```{r}
RJ_chorus <- RJ_analysis |>
  filter(time_day <= as.hms("10:00:00", format = "%h%m:%s"))  |>
  group_by(transect, distance, days) |>
  summarise(n = n()) 
```

```{r}
RJ_chorus_model.1 <- glmer(n ~ distance * days + (1|transect), 
                         data = RJ_chorus, family = poisson)
## rescale variables 
RJ_chorus_model.2 <- glmer(n ~ scale(distance) * days + (1|transect), 
                         data = RJ_chorus, family = poisson)

plot(simulateResiduals(RJ_chorus_model.2))
## problem with residuals 

RJ_chorus_model.3 <- glmer.nb(n ~ scale(distance) * days + (1|transect), data = RJ_chorus)

plot(simulateResiduals(RJ_chorus_model.3))

summary(RJ_chorus_model.3)
```
### simplification

```{r}
RJ_chorus_model.4 <- glmer.nb(n ~ scale(distance) + days + (1|transect), data = RJ_chorus)
summary(RJ_chorus_model.4)
```

# Asian Koel 
  
```{r}
AK_analysis <- AK |>
  filter(top_match == "Koel") |>
  filter(top_dist <= 1) 
```


## duration 

```{r}
AK_duration_model <- lme(DURATION ~ distance * days, random = ~1|transect, data = AK_analysis)

## diagnostic plot

### variance for the dataset 
plot(AK_duration_model, resid(.,scaled = T) ~ fitted(.) | transect, abline = 0, lty = 2)
plot(RJ_duration_model)

### normality 
qqnorm(AK_duration_model, abline = c(0, 1), lty = 2)
qqnorm(AK_duration_model, ~resid(., type = "p") | transect , abline = c(0,1) , lty = 2)

### large enough sample size always reject the null hypothesis? 
ad.test(log(AK_analysis$DURATION))
hist(log(AK_analysis$DURATION))

AK_duration_model.2 <- lme(log(DURATION) ~ distance * days, random = ~1|transect, data = AK_analysis)
summary(AK_duration_model.2)

AK_duration_model.3 <- lme(log(DURATION) ~ distance + days, random = ~1|transect, data = AK_analysis)
summary(AK_duration_model.3)

```

## dawn chorus 

```{r}
AK_chorus <- AK_analysis |>
  filter(time_day <= as.hms("10:00:00", format = "%h%m:%s"))  |>
  group_by(transect, distance, days) |>
  summarise(n = n()) 
```

```{r}
AK_chorus_model.1 <- glmer(n ~ distance * days + (1|transect), 
                         data = AK_chorus, family = poisson)
## rescale variables 
AK_chorus_model.2 <- glmer(n ~ scale(distance) * days + (1|transect), 
                         data = AK_chorus, family = poisson)

plot(simulateResiduals(AK_chorus_model.2))
## problem with residuals 

AK_chorus_model.3 <- glmer.nb(n ~ scale(distance) * days + (1|transect), data = AK_chorus)

plot(simulateResiduals(AK_chorus_model.3))
# problem not solved 

AK_chorus_model.4 <- glmmPQL(n ~ scale(distance) * days, random = ~1|transect, 
                             data = AK_chorus, family = quasipoisson)
```

```{r}
summary(AK_chorus_model.4)
## interaction not significant 

AK_chorus_model.5 <- glmmPQL(n ~ scale(distance) + days, random = ~1|transect, 
                             data = AK_chorus, family = quasipoisson)
summary(AK_chorus_model.5)

AK_chorus_model.6 <- glmmPQL(n ~ days, random = ~1|transect, 
                             data = AK_chorus, family = quasipoisson)
summary(AK_chorus_model.6)
```

# Summary 

## Junglefowl 
distance is significantly related to duration and dwn chorus 

## Asian Koel 
distance and type of day are significantly associated with the duration of the call, 
no association found for the dawn chorus. 



