---
title: "Lecture7data"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


t-test
```{r}
old = c(20.8, 2.8, 50.0, 33.3, 29.4, 38.9, 29.4, 52.6, 14.3)
young = c(45.5, 55.0, 60.7, 61.5, 61.1, 65.5, 42.9, 37.5)
```


```{r}
mean(young)
mean(old)
```

```{r}
?t.test
```

```{r}
t.test(young, old, var.equal = TRUE)
# see that the condifence interval does not contain 0 
```

```{r}
# test if the two samples ha ve equal standard deviation
var.test(young, old)
```

```{r}
(mean(young) - mean(old)) / ( sd(young) * (sqrt(1/length(young) + 1/length(old))))
```

paired t-test
```{r}
data = data.frame(crossfertilized = 
                    c(23.5, 12, 21, 22, 19.125, 21.5, 22.125,
                      20.375, 18.250, 21.625, 23.25, 21.0, 
                      22.125, 23.0, 12.0),
           selffertilized = 
             c(17.375, 20.375, 20.0, 20.0, 18.375, 18.625, 
               18.625, 15.25, 16.5, 18.0, 16.25, 18.0, 12.75, 15.5, 18.0))
```

```{r}
var.test(data$crossfertilized, data$selffertilized)
```

```{r}
t.test(data$crossfertilized, data$selffertilized, paired = TRUE, var.equal = FALSE)
```

lots of data in biology is non-normal 
small sample sizes (problem whtehr you use parametric or non-parametric)
non-parametric makes no assumption about the underlying distribution 

- we dont test the number as data themselves. 
- we calculate the rank 
- get the sum of the ranks by class 
- calculates the difference in sums 

non-parametric
```{r}
old = c(3, 1, 11, 6, 4.5, 8, 4.5, 12, 2)
young = c(10, 13, 14, 6, 15, 17, 9, 7)
```

```{r}
?wilcox.test
```

```{r}
wilcox.test(young, old, alternative = 'greater')
```

```{r}
foo <- data.frame(condition = c(rep("Old", length(old)),
                                rep("Young", length(young))),
                                data = c(old,young))
foo$rank <- rank(foo$data)
foo
```

```{r}
stripchart(rank ~ condition, data = foo, vertical = TRUE)
```

```{r}
sum(subset(foo, condition == "Old")$rank)
sum(subset(foo, condition == "Young")$rank)
```
```{r}
# test statistic 
sum(subset(foo, condition == "Young")$rank) - sum(subset(foo, condition == "Old")$rank)
```
# if we wanna see if the test statistics are significant, we can do the permutation by mixing the labels 

```{r}
foo
```

```{r}
test_statistics = c()
for (i in 1:1000){
  foo_sample = sample(1:17, 17, replace = FALSE)
  old_sample = foo$rank[foo_sample[which(foo$condition == "Old")]]
  young_sample = foo$rank[foo_sample[which(foo$condition == "Young")]]
  old_sum = sum(old_sample)
  young_sum = sum(young_sample)
  test_statistics[i] <- young_sum - old_sum
}
```

```{r}
hist(test_statistics)
```

```{r}
sum(test_statistics >= 42) / length(test_statistics)
```
# lets try with the mean 
```{r}
ts = mean(foo$data[foo$condition == "Young"]) - mean(foo$data[foo$condition == "Old"])
ts
```

```{r}
test_stats_mean = c()
for (i in 1:1000){
  foo_sample = sample(1:17, 17, replace = FALSE)
  old_sample = foo$data[foo_sample[which(foo$condition == "Old")]]
  young_sample = foo$data[foo_sample[which(foo$condition == "Young")]]
  old_mean = mean(old_sample)
  young_mean = mean(young_sample)
  test_stats_mean[i] <- young_mean - old_mean
}
```

```{r}
hist(test_stats_mean)
```
```{r}
sum(test_stats_mean >=  ts) / length(test_stats_mean)
```
```{r}
40/120 * 55/120 * 120
```

```{r}
insects = data.frame(
  spray = c(15,35,5),
  non_spray = c(25,25,15)
)

insects
```
```{r}
qchisq(0.95, 2)
```

```{r}
chisq.test(insects)
```
```{r}
0.8^2 + 0.32 + 0.04
```

```{r}
chisq.test(c(77,38,10), p = c(0.64, 0.32, 0.04))
```

```{r}
mut <- data.frame(
  normal = c(7,1),
  weird = c(2,5)
)
mut
```

```{r}
chisq.test(mut)
```

```{r}
chisq.test(mut)$expected
```

the warning comes when the expected frequencies at least one entry is less than 5

```{r}
fisher.test(mut)
```

