---
title: "Lecture3"
output: html_document
---

```{r}
library(dplyr)
library(ggplot2)
```

### Part 1

Body temperature of 12 individuals in degrees centigrade 
```{r}
table7_1 = c(37.0, 36.0, 37.1, 37.1, 36.2, 37.3, 36.8, 37.0, 36.3, 36.9, 36.9, 36.7, 36.8)
```

What is the mean of this dataset? 

```{r}
mean(table7_1)
```

What is the median of this dataset?

```{r}
median(table7_1)
```

Trim out the bottom and top 10% of values and calculate the mean?

```{r}
mean(table7_1, trim = 0.1)
```

What is the mode of this dataset?

<https://statisticsglobe.com/mode-in-r-programming-example>
```{r}
# 
mode <- function(x) {                      
  unique_x <- unique(x)
  tabulate_x <- tabulate(match(x, unique_x))
  unique_x[tabulate_x == max(tabulate_x)]
}
```

```{r}
mode(table7_1)
```
```{r}
names(table(table7_1)[table(table7_1) == max(table(table7_1))])
```


# plot a histogram of table7_1 and add lines to show the measures of central tendency

```{r}
ggplot() + 
  aes(table7_1)+ 
  geom_histogram(binwidth=0.1, colour="black", fill="white")+
  labs(x = "Body temperarue")+
  geom_vline(xintercept = mean(table7_1), colour = "green")+
  geom_vline(xintercept = median(table7_1), colour = "blue")
```

what is the range of this data? 

```{r}
range(table7_1)
max(table7_1) - mean(table7_1)
```

what is the IQR of this dataset? 

```{r}
IQR(table7_1)
```

what are the quartiles of this dataset?

```{r}
quantile(table7_1)
```


what is the standard deviation of this dataset? 

```{r}
sd(table7_1)
```

what is the variance of this dataset

```{r}
var(table7_1)
```

##### World bank

the file _worldbank.tsv_ contains lots of information on different statistics associated with different countries ... lets load this into R

```{r}
worldbank = read.delim("worldbank.tsv", sep='\t')
```

lets make a histogram of the population growth .... 

```{r}
ggplot(data = worldbank, aes(x = Population_growth))+
  geom_histogram(fill = 'white',
                 color = 'black')+
  labs(x = "Population Growth")
```

OK - how about making a boxplot of Population_growth - separating countries by their Climate Region?

```{r}
ggplot(data = worldbank, aes(x = Region, y = Population_growth))+
  geom_boxplot()+
  labs(x = "Climate Region",
       y = "Population Growth")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

what about a scatter plot of  Population_growth - separating countries by their Climate Region?

```{r}
# ?stripchart
```

```{r}
ggplot(data = worldbank, aes(x=Region, y=Population_growth)) + 
  geom_jitter(position=position_jitter(0.1), size = 2, alpha = 0.5)+
  labs(x = "Climate Region",
       y = "Population Growth")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


what about a violin plot of Population_growth - separating countries by Climate Region? 

```{r, messages=FALSE, warnings=FALSE}
library(vioplot)
vioplot(worldbank$Population_growth ~ worldbank$Region,
        xlab = "", ylab = "Population Growth", las = 2)
```

Plot the cumulative frequency distribution of Population_growth 

```{r}
# ?ecdf
```

```{r}
ggplot(data = worldbank, aes(x = Population_growth)) +
  stat_ecdf(geom = "step")+
  labs(x = "Population Growth",
       y = "Frequency",
       title = "Cumulative Frequency Distribution")+
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
sessionInfo()
```