---
title: "New York Mayoral Election 2017"
author: "Team 5B"
date: "1/26/2022"
output: 
    html_document:
        toc: true
        theme: united
---

``` {r setup, include=FALSE, echo=FALSE}
library(kableExtra)
library(tidyverse)
library(scales)
ny_mayor <- read.csv("ny_mayor.csv")
```

# Introduction

The wikipedia article indicates that Bill de Blaiso won[^1].

    ‘Incumbent Democrat Bill de Blasio won reelection to a second term with 66.2% of the vote against Republican Nicole Malliotakis.’

Thus, it is our clear belief, as the people of wikipedia speak the truth, that Bill really did win.

# Plotting Against the Table

``` {r setup, include=FALSE}
kbl(ny_mayor,
    caption = "New York Mayoral Election 2017")
```
## Data presented as a table

``` {r}

```

| Borough | de Biasio | Malliotakis | Other |
|:--------|:----------|:------------|:------|
|         |           |             |       |
|         |           |             |       |
|         |           |             |       |
|         |           |             |       |
|         |           |             |       |
|         |           |             |       |


## Data presented as a bar plot

``` {r include=FALSE, echo=FALSE}
# Put calls to library() at the start of your RMD file
ny_mayor_long <- read_csv("ny_mayor.csv") |>
mutate(borough = str_replace(borough, " ", "\n")) |> pivot_longer(-borough) |>
mutate(name = fct_rev(name))
ggplot(ny_mayor_long, aes(borough, value, fill = name)) +
  geom_col(position = "dodge") +
  scale_x_discrete(limits = rev) +
  coord_flip() +
  scale_y_continuous(labels = label_comma()) +
  scale_fill_manual(
    labels = c("Other", "Malliotakis", "de Blasio"),
values = c("palegreen4", "firebrick2", "steelblue") )+
  labs(
    title = "New York Mayoral Election 2017",
    caption = "Source: CNN",
    x = "Borough",
    y = "Votes",
    fill = "Candidate" 
) +
guides(fill = guide_legend(reverse = TRUE)) +
theme(legend.key.size = unit(1, "lines"))
```

## What is better: table or bar plot?

[^1]:Accessed [here](https://en.wikipedia.org/wiki/2017_New_York_City_mayoral_election) on 2022-01-26.
