---
title: "New York Mayoral Election 2017"
author: "Shuna"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
---

```{r, include=FALSE, echo=FALSE, message=FALSE}
library(tidyverse)
library(dplyr)
library(scales)
library(kableExtra)
library(readr)
library(janitor)

ny <- read_csv("ny_mayor.csv")
```

### Introduction 

An election for Mayor of New York City was held on November 7, 2017.

>"Incumbent Democrat Bill de Blasio won reelection to a second term with 66.2% of the vote against Republican Nicole Malliotakis." ^[Quote from <https://en.wikipedia.org/wiki/2017_New_York_City_mayoral_election>, accessed on 28 January 2022]

Wow that's great. 

### Data presented as a table 
```{r}
# add a row for sum of each column
ny_with_total<- ny %>%
  # reference :<https://stackoverflow.com/questions/4946873/add-row-to-a-data-frame-with-total-sum-for-each-column>
  adorn_totals("row")
# convert it to a data frame 
ny <- as.data.frame(ny_with_total)
```


```{r}
# making a table
ny %>%
  # add a caption
  kbl(
    caption = "New York Mayoral Election 2017",
    # add commas for numbers
    format.args = list(big.mark = ",")
  ) %>%
  # group columns
  add_header_above(c(
    " " = 1,
    "Candidate" = 3
  )) %>%
  # add footnote
  footnote(
    general = "<small>Source: CNN</small>",
    # remove title
    general_title = "",
    # keep the footnote small
    fixed_small_size = TRUE,
    # title will not be excaped against HTML
    escape = FALSE
  ) %>%
  # add alternative row colors
  kable_paper("striped",
    # reduce the size
    full_width = F
  ) %>%
  # add a solid line for total
  row_spec(5,
    extra_css = c("border-bottom: 1px solid black")
  )
  
```

### Data presented as a bar plot 

```{r}
ny_mayor_long <- read_csv("ny_mayor.csv") |>
  mutate(borough = str_replace(borough, " ", "\n")) |>
  pivot_longer(-borough) |>
  mutate(name = fct_rev(name))

ggplot(
  ny_mayor_long,
  aes(borough, value, fill = name)
) +
  # make barplot
  geom_col(position = "dodge") +
  scale_x_discrete(limits = rev) +
  # flip the plot
  coord_flip() +
  # add commas to the numbers
  scale_y_continuous(labels = label_comma()) +
  # edit the legend and colors
  scale_fill_manual(
    labels = c(
      "Other",
      "Malliotakis",
      "de Blasio"
    ),
    values = c(
      "palegreen4",
      "firebrick2",
      "steelblue"
    )
  ) +
  # add axis labels, titles, and footnote
  labs(
    title = "New York Mayoral Election 2017",
    caption = "Source: CNN",
    x = "Borough",
    y = "Votes",
    fill = "Candidate"
  ) +
  guides(fill = guide_legend(reverse = TRUE)) +
  theme(legend.key.size = unit(1, "lines"))

```
### What is better: table or bar plot? 

advantages 
- can get accurate representation of the data 
- can see the overall trend by total row (though we can simply add the same for the bar plot) 



disadvantages 
- harder to make comparisons 




