---
title: "New York Mayoral Election 2017"
author: "Shuna"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
---

```{r, include=FALSE, echo=FALSE, message=FALSE}
library(tidyverse)
library(dplyr)
library(scales)
library(kableExtra)
library(readr)
library(janitor)

ny <- read_csv("ny_mayor.csv")
```

### Introduction 

An election for Mayor of New York City was held on November 7, 2017.

>"Incumbent Democrat Bill de Blasio won reelection to a second term with 66.2% of the vote against Republican Nicole Malliotakis." ^[Quote from <https://en.wikipedia.org/wiki/2017_New_York_City_mayoral_election>, accessed on 28 January 2022]

This quote is extracted from the Wikipedia page. 

### Data presented as a table 
```{r, echo=FALSE}
# add a row for sum of each column
ny_with_total<- ny |>
  # reference :<https://stackoverflow.com/questions/4946873/add-row-to-a-data-frame-with-total-sum-for-each-column>
  adorn_totals("row")
# convert it to a data frame 
ny <- as.data.frame(ny_with_total)
# rename columns 
ny <- rename(ny, 
       "Borough" = "borough",
       "de Blasio" = "de_blasio",
       "Malliotakis" = "malliotakis",
       "Other" = "other")

```


```{r, echo=FALSE}
# making a table
ny |>
  # add a caption
  kbl(caption = "New York Mayoral Election 2017",
      # add commas for numbers
      format.args = list(big.mark = ",")) |>
  # group columns 
  add_header_above(c(" " = 1, 
                     "Candidate" = 3)) |>
  # add footnote
  footnote(general = "<small>Source: CNN</small>",
           # remove title 
           general_title = "",
           # keep the footnote small
           fixed_small_size = TRUE,
           # title will not be excaped against HTML
           escape = FALSE) |>
  # add alternative row colors
  kable_paper("striped", 
              # reduce the size
              full_width = F) |>
  # add a solid line for total 
  row_spec(5,
    extra_css = c("border-bottom: 1px solid black"))
  
```


### Data presented as a bar plot 

```{r, echo=FALSE, message=FALSE}
ny_mayor_long <- read_csv("ny_mayor.csv") |>
  mutate(borough = str_replace(borough, " ", "\n")) |>
  pivot_longer(-borough) |>
  mutate(name = fct_rev(name)) 

ggplot(ny_mayor_long, 
       aes(borough, value, fill = name)) +
  # make barplot
  geom_col(position = "dodge") +
  scale_x_discrete(limits = rev) +
  # flip the plot
  coord_flip() + 
  # add commas to the numbers
  scale_y_continuous(labels = label_comma()) + 
  # edit the legend and colors 
  scale_fill_manual(
    labels = c("Other", 
               "Malliotakis", 
               "de Blasio"),
    values = c("palegreen4", 
               "firebrick2", 
               "steelblue")) + 
  # add axis labels, titles, and footnote
  labs(
    title = "New York Mayoral Election 2017",
    caption = "Source: CNN",
    x = "Borough",
    y = "Votes",
    fill = "Candidate") + 
  guides(fill = guide_legend(reverse = TRUE)) + 
  theme(legend.key.size = unit(1,"lines"))

```

### What is better: table or bar plot? 

With a table, we can easily obtain the exact numerical values and quantities without having to go an extra step of calculating it from the height of each bar.
The total number of votes is much easier to represent in a table than a bar plot
too. It also provides a more "objective" sense of comparison between the different values, as compared to bar plots. The latter may be more susceptible to "visual manipulation" to mislead viewers into making comparisons of inaccurate magnitudes (e.g. when the y axis does not start at 0). 

However, if we wish to infer a trend or make other comparisons, a table does not 
provide as much visual convenience to aid us like a bar plot would. For example, 
we would not be able to visualize the exact magnitude to which de Blasio outcompeted his opponents, but a bar graph allows us to easily see that he
received a landslide victory in most boroughs. If we wanted to understand the 
rough proportion of votes taken by other candidates at first glance, the proportions of each bar helps us reach a rough conclusion faster. 




