---
title: "New York Mayoral Election 2017"
author: "Group A"
date: "1/26/2022"
output: 
  html_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(kableExtra)
ny_mayor <- read.csv("ny_mayor.csv")
library(tidyverse)
library(scales)
```

## Introduction

An election for Mayor of New York City was held on November 7, 2017. 

> ‘Incumbent Democrat Bill de Blasio won reelection to a second term with 66.2% 
of the vote against Republican Nicole Malliotakis.’ [^1]

This following exercise will show the results of the election in tables. 

[^1]: <https://en.wikipedia.org/wiki/2017_New_York_City_mayoral_election> Accessed on 26/1/2022

## Data presented as a table


```{r table, echo = FALSE}
ny_mayor[nrow(ny_mayor) + 1, ] <- c("Total", sum(ny_mayor$de_blasio), sum(ny_mayor$malliotakis), sum(ny_mayor$other))
# from: https://statisticsglobe.com/add-new-row-to-data-frame-in-r accessed on 29/1/2022

kbl(
  ny_mayor,
  caption = "New York Mayoral Election 2017",
  format.args = list(big.mark = ","),
  col.names = c("Borough", "de Balsio", "Malliotakis", "Other")
) |>
  add_header_above(c("", "Candidate" = 3)) |>
  kable_paper(full_width = F, html_font = "Arial") |>
  kable_styling("striped", full_width = F) |>
  row_spec(5, extra_css = "border-bottom: 1px solid;") |>
  # from: https://stackoverflow.com/questions/68010981/table-with-package-kableextra-how-to-add-lines-to-the-tables
  footnote(
    general_title = " ",
    general = c("<small>Source: CNN </small>"),
    footnote_as_chunk = T,
    escape = F
  )
# from: https://stackoverflow.com/questions/60176526/kableextra-change-font-size-of-table-footnote accessed on 29/1/2022
# Could not find how to modify "CNN" to the colour blue and add the hyperlink to it.
```

## Data presented as a bar plot
```{r bar plot, echo = FALSE}
ny_mayor_long <- read_csv("ny_mayor.csv") |>
  mutate(borough = str_replace(borough, " ", "\n")) |>
  pivot_longer(-borough) |>
  mutate(name = fct_rev(name))
ggplot(ny_mayor_long, aes(borough, value, fill = name)) +
  geom_col(position = "dodge") +
  scale_x_discrete(limits = rev) +
  coord_flip() +
  scale_y_continuous(labels = label_comma()) +
  scale_fill_manual(
    labels = c("Other", "Malliotakis", "de Blasio"),
    values = c("palegreen4", "firebrick2", "steelblue")
  ) +
  labs(
    title = "New York Mayoral Election 2017",
    caption = "Source: CNN",
    x = "Borough",
    y = "Votes",
    fill = "Candidate"
  ) +
  guides(fill = guide_legend(reverse = TRUE)) +
  theme(legend.key.size = unit(1, "lines"))
```


## What is better: table or bar plot?
Advantages of a table:

- Ability to know the exact values for each data point. 
- The data will be organised such that the data for each candidate is grouped
together in one column, making it easy to locate which data point you 
want to look at according to the borough. 

Disadvantages of a table: 

- Inability to instantly read trends for the candidates over different boroughs. 
- Readers may not be able to know which candidate has the most votes 
at a quick glance. 

Advantages of a bar plot:

- Readers are able to visualize the data for each candidate at a glance, thus, 
conveniently knowing each candidate has the most or least votes for each Borough. 
- This also leads to the data being easily understood among a wider audience. 
- Each bar can also be filled with color, making the trends bold and clear for 
every candidate. 
- In the bar chart shown above, the colors, blue, red, and green are 
distinguished easily from each other, enabling easier interpretations of the 
data. 

Disadvantages of a bar plot:

- Not being able to know the exact number of votes each candidate has, only a 
rough gauge on where the number lies along the x-axis. 
- Additionally, having different data sets plotted on the same bar plot means 
that some data sets might get 'squished' (e.g. differences in the 'Other' borough 
cannot be seen clearly because the scale is dominated by the larger 'de Blasio' 
borough data). 
