---
title: "Practising R Programming"
author: "Team D"
date: "1/28/2022"
output: html_document
---

``` {r setup, include=FALSE, echo=FALSE}
library(dplyr)
library(bsts)
library(microbenchmark)
```

# Days of the week in Shire calendar

```{r}
# shire_day_from_georgian <- function(georgian) {
#   dplyr::if_else()
# }
```

```{r}
shire_day_from_georgian <- function(g_day) {
  dplyr::case_when(
    g_day == "Monday" ~ "Sterday",
    g_day == "Tuesday" ~ "Sunday",
    g_day == "Wednesday" ~ "Monday",
    g_day == "Thursday" ~ "Trewsday",
    g_day == "Friday" ~ "Hevensday",
    g_day == "Saturday" ~ "Mersday",
    g_day == "Sunday" ~ "Highday",
    TRUE ~ NA_character_
  ) 
}

shire_day_from_georgian("Monday")
shire_day_from_georgian("Tuesday")
shire_day_from_georgian("Wednesday")
shire_day_from_georgian("Thursday")
shire_day_from_georgian("Friday")
shire_day_from_georgian("Saturday")
shire_day_from_georgian("Sunday")
shire_day_from_georgian("Nonday")
```

```{r}
georgian <- c(
  "Monday", 
  "Tuesday", 
  "Wednesday", 
  "Thursday", 
  "Friday", 
  "Saturday", 
  "Sunday" 
)

shire <- c(
  "Sterday",
  "Sunday",
  "Monday",
  "Trewsday",
  "Hevensday",
  "Mersday",
  "Highday"
)

shire_day_from_georgian <- function(g_day) {
  dplyr::recode(g_day, georgian = shire, .default = NA_character_)
}

shire_day_from_georgian("Monday")
```

# Measuring run-times with the microbenchmark package

```{r}
abs_with_if_else <- function(x) { dplyr::if_else(x >= 0, x, -x) }

abs_with_subsetting <- function(x) { 
  neg <- (x < 0)
       x[neg] <- -x[neg]
  x }

abs_with_data_type_conversion <- function(x)  { ((x > 0) - (x < 0)) * x }

abs_with_for_loop <- function(x) { 
  for (i in seq_along(x)) {
    if (x[i] < 0) {
      x[i] <- -x[i]
    }
  }
  x 
}

nums <- rnorm(1e6)

microbenchmark::microbenchmark(
  abs(nums),
  abs_with_if_else(nums),
  abs_with_subsetting(nums),
  abs_with_data_type_conversion(nums),
  abs_with_for_loop(nums)
)

```

# Comparing two functions for calulating Pythagorean sums

```{r}
pythag1 <- function(a, b) {
  sqrt(a^2 + b^2)
}

pythag2 <- function(a, b) {
  absa <- abs(a)
  absb <- abs(b)
  p <- pmax(absa, absb)
  q <- pmin(absa, absb)
  ratio <- q / p
  ratio[is.nan(ratio)] <- 1
  p * sqrt(1.0 + ratio^2)
}
```

## What is the purpose of is.nan(ratio) in the second-to-last line of pythag2()’s function body?

## Compare the run-times of pythag1() and pythag2() with the microbenchmark package. Use identical input consisting of long numeric vectors. Summarise your observation as a comment in the R script.

```{r}
a <- rnorm(1e5)
b <- rnorm(1e5)

microbenchmark::microbenchmark(
  pythag1(a, b),
  pythag2(a, b)
)
```

## By performing numerical tests, find out under which conditions the functions numerically overflow. When do the functions underflow? Comment on the observed differences between pythag1() and pythag2().

This will work:

```{r}
x_large <- 3e306
y_large <- 4e306

pythag1(x_large, y_large)
pythag2(x_large, y_large)
```

But this will not:

```{r}
x_large <- 3e307
y_large <- 4e307

pythag1(x_large, y_large)
pythag2(x_large, y_large)
```

Similarly, this will work:

```{r}
x_small <- 3e-324
y_small <- 4e-324

pythag1(x_small, y_small)
pythag2(x_small, y_small)
```

But this will not:

```{r}
x_small <- 3e-325
y_small <- 4e-325

pythag1(x_small, y_small)
pythag2(x_small, y_small)
```

## Is pythag1() or pythag2() better as a general-purpose method? There is no simple right-or-wrong answer. I am interested in your reasoning.



# Floating-point accuracy

```{r}
0.1 + 0.2
0.1 + 0.2 == 0.3
round(0.1 + 0.2) == 0.3
all.equal(0.1 + 0.2, 0.3)
all.equal(0.1 + 0.2, 0.3, tolerance = .Machine$double.eps)
all.equal(0.2, 0.3, tolerance = 0.2)
all.equal(1.1, 10.1, tolerance = 20)
```

The default value of tolerance is close to 1.5e-8 (according to the help page). By modifying this value, we can have any two numbers considered equal. Thus, since (0.1 + 0.2) cannot be represented accurately in binary, all that all.equal does is compare the two values, and ensuring they are within the tolerance.